---
title: "profiling veo data"
output: html_document
author: Joanna Schroeder
---

Veteran Employment Outcomes (VEO) are new experimental Census Bureau statistics on labor market outcomes for recent Army veterans. These statistics are tabulated by occupation, employer industry (if employed), and various veteran characteristics. They are generated by matching service member information with a national database of jobs, using state-of-the-art confidentiality protection mechanisms to protect the underlying data. The VEO are made possible through data sharing partnerships between the U.S. Army, State Labor Market Information offices, and the U.S. Census Bureau. VEO data are currently only available at the national level.

The VEO data covers all Army service members who have completed their initial term of service - meaning they have served the time they signed up for when they enlisted and were not discharged early - and were discharged between 2000 and 2015.

Types of variables: Earnings, employment, separation cohort, age, sex, race, ethnicity, education at enlistment, AFQT test score, pay grade, years of service, military occupation, employer geography (not available yet), employer industry.

```{r setup, include=FALSE, }
library(readxl)
library(ggplot2)
library(reshape2)
library(tidyverse)
library(igraph)
library(scales)

get_db_conn <-
  function(db_name = "sdad",
           db_host = "postgis1",
           db_port = "5432",
           db_user = Sys.getenv("db_usr"),
           db_pass = Sys.getenv("db_pwd")) {
    RPostgreSQL::dbConnect(
      drv = RPostgreSQL::PostgreSQL(),
      dbname = db_name,
      host = db_host,
      port = db_port,
      user = db_user,
      password = db_pass
    )
  }

con <- get_db_conn()

job <- DBI::dbGetQuery(con, "
                       SELECT * FROM bgt_res.job
                       WHERE job.id in (
                       SELECT ID FROM bgt_res.pers
                       WHERE pers.msa like '%47900%'
                       )"
)

DBI::dbDisconnect(con)

veo_occ <- read_excel("~/git/DSPG2020/career/src/lehd/veopu.xlsx", sheet = 3, range = "B5:AE117")

veo_ind <- read_excel("~/git/DSPG2020/career/src/lehd/veopu.xlsx", sheet = 5, range = "B5:Y173")

veo_age <- read_excel("~/git/DSPG2020/career/src/lehd/veopu.xlsx", sheet = 6, range = "B5:AD21")

veo_sex <- read_excel("~/git/DSPG2020/career/src/lehd/veopu.xlsx", sheet = 7, range = "B5:AD21")

veo_race <- read_excel("~/git/DSPG2020/career/src/lehd/veopu.xlsx", sheet = 8, range = "B5:AF101")

veo_edu <- read_excel("~/git/DSPG2020/career/src/lehd/veopu.xlsx", sheet = 10, range = "B5:AD29")
```

# Age and Employment Counts
```{r warning = FALSE}
# Age at enlistment
veo_age_long <- veo_age %>% select(label_cohort, label_agegrp, y1_emp, y1_nonemp) %>% melt(id.vars = c("label_cohort", "label_agegrp"))

veo_age_long$value <- as.numeric(veo_age_long$value)

ggplot(veo_age_long, aes(x = variable, y = value, fill = label_agegrp)) +
  geom_col(position = "dodge") +
  labs(title = "VEO Age and Employment Counts")
```

# Sex and Employment Counts
```{r warning = FALSE}
veo_sex_long <- veo_sex %>% select(label_cohort, label_sex, y1_emp, y1_nonemp) %>% melt(id.vars = c("label_cohort", "label_sex"))

veo_sex_long$value <- as.numeric(veo_sex_long$value)

ggplot(veo_sex_long, aes(x = variable, y = value, fill = label_sex)) +
  geom_col(position = "dodge") +
  labs(title = "VEO Sex and Employment Counts")
```

# Education and Employment Counts
```{r warning = FALSE}
#Education at enlistment
veo_edu_long <- veo_edu %>% select(label_cohort, label_education, y1_emp, y1_nonemp) %>% melt(id.vars = c("label_cohort", "label_education"))

veo_edu_long$value <- as.numeric(veo_edu_long$value)

ggplot(veo_edu_long, aes(x = variable, y = value, fill = label_education)) +
  geom_col(position = "dodge") +
  labs(title = "VEO Education and Employment Counts")
```

# Race and Employment Counts
```{r warning = FALSE}
veo_race_long <- veo_race %>% select(label_cohort, label_race, label_ethnicity, y1_emp, y1_nonemp) %>% melt(id.vars = c("label_cohort", "label_race", "label_ethnicity"))

veo_race_long$value <- as.numeric(veo_race_long$value)

ggplot(veo_race_long, aes(x = variable, y = value, fill = label_race)) +
  geom_col(position = "dodge") +
  labs(title = "VEO Race and Employment Counts")

ggplot(veo_race_long, aes(x = variable, y = value, fill = label_ethnicity)) +
  geom_col(position = "dodge") +
  labs(title = "VEO Ethnicity and Employment Counts")
```

# VEO Occupation and Employment Counts
```{r warning = FALSE}
# Read in crosswalk between DOD and SOC codes
crosswalk <- read_csv("~/git/DSPG2020/career/src/lehd/milx0120.csv")
military <- read_csv("~/git/DSPG2020/career/src/lehd/Military_Specific.csv")

mil_crosswalk <- crosswalk %>% select(DODOCC, ONET1) %>% filter(ONET1 %in% military$Code) %>% transmute(DODOCC = str_extract(DODOCC, "[0-9][0-9][0-9]"), ONET1) %>% distinct() %>% filter(DODOCC <= 186)

mil_crosswalk_nodes <- gather(mil_crosswalk) %>% select(value, key) %>% distinct()

net <- graph_from_data_frame(d=mil_crosswalk, vertices = mil_crosswalk_nodes, directed = F) 

library(RColorBrewer)
coul  <- brewer.pal(2, "Set1")
my_color <- coul[as.numeric(as.factor(V(net)$key))]

plot(net, vertex.color=my_color)

veo_occ_long <- veo_occ %>% select(label_cohort, dod_occ_code, y1_emp, y1_nonemp, y5_emp, y5_nonemp, y10_emp, y10_nonemp) %>% melt(id.vars = c("label_cohort", "dod_occ_code"))

veo_occ_long$value <- as.numeric(veo_occ_long$value)

ggplot(veo_occ_long, aes(x = dod_occ_code, y = value, fill = dod_occ_code)) +
  geom_col(position = "dodge") +
  facet_grid(rows = vars(variable)) +
  labs(title = "VEO Occupation and Employment Counts")
```

# VEO Industry and Employment Counts
```{r warning = FALSE}
# Create long df of veo industries
veo_ind_long <- veo_ind %>% select(label_cohort, label_industry, industry, y1_emp, y5_emp, y10_emp) %>% melt(id.vars = c("label_cohort", "label_industry", "industry"))

veo_ind_long$value <- as.integer(veo_ind_long$value)

# Remove created federal government industry code for comparison
veo_ind_long<- veo_ind_long %>% mutate(industry_compare = ifelse(industry == 99, 92, industry)) %>% mutate(label_industry_compare = ifelse(industry == 99, "Public Administration", label_industry))
  
# Calculate percentage in industries
veo_ind_long_perc <- veo_ind_long %>% drop_na() %>% group_by(variable) %>% mutate(perc = value/sum(value))

# Plot
ggplot(veo_ind_long_perc, aes(x = label_industry_compare, y = perc, fill = label_industry_compare)) +
  geom_col(show.legend = FALSE) + 
  labs(title = "VEO Industry and Employment Counts") +
  #facet_grid(rows = vars(variable)) +
  coord_flip() +
  theme_classic()
```

```{r warning = FALSE}
# Create pattern for military ids and gather ids
military <- "55-[0-9][0-9][0-9][0-9].[0-9][0-9]"
military_ids <- job %>% filter(str_detect(job$onet, military)) %>% distinct(id)

# Create df of resumes for military ids, get counts in each sector
res_mil <- job %>% filter(id %in% military_ids$id) %>% group_by(sector) %>% summarise(n = n()) %>% drop_na()

# Join industry labels by NAICS codes
res_mil_labels <- res_mil %>% left_join(select(veo_ind_long, label_industry, industry), by = c("sector" = "industry"))

# Calculate percent in each sector, prepare dfs for rbind
res_mil_labels <- res_mil_labels %>% distinct() %>% mutate(perc = n/sum(n)) %>% mutate(variable = "bgt") %>% rename(value = n, label_industry_compare = label_industry, industry_compare = sector)

veo_ind_long_perc <- veo_ind_long_perc %>% select(variable, value, industry_compare, label_industry_compare, perc)

#Rbind
veo_bgt_industry_profile <- rbind(veo_ind_long_perc, res_mil_labels)

#Sum industry percent
veo_bgt_industry_profile <- veo_bgt_industry_profile %>% group_by(variable, industry_compare) %>% mutate(percen = sum(perc))

ggplot(veo_bgt_industry_profile, aes(x = label_industry_compare, y = percen, fill = variable)) +
  geom_col(position = "dodge") +
  labs(title = "Combined Industry and Employment Percents") +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_classic()

```

```{r}
lodes <- read_excel("~/git/DSPG2020/career/src/lehd/otm.xls", sheet = 1, range = "A24:AG45")

lodes <- lodes %>% rename("label_industry" = "...1", "count_2017" = "2017", "share_2017" = "...3", "count_2016" = "2016", "share_2016" = "...5", "count_2015" = "2015", "share_2015" = "...7", "count_2014" = "2014", "share_2014" = "...9", "count_2013" = "2013", "share_2013" = "...11", "count_2012" = "2012", "share_2012" = "...13", "count_2011" = "2011", "share_2011" = "...15","count_2010" = "2010", "share_2010" = "...17", "count_2009" = "2009", "share_2009" = "...19", "count_2008" = "2008", "share_2008" = "...21", "count_2007" = "2007", "share_2007" = "...23", "count_2006" = "2006", "share_2006" = "...25", "count_2005" = "2005", "share_2005" = "...27", "count_2004" = "2004", "share_2004" = "...29", "count_2003" = "2003", "share_2003" = "...31", "count_2002" = "2002", "share_2002" = "...33")

# Remove first row
lodes <- lodes[-c(1), ]

# Change to long
lodes <- lodes %>% select("label_industry", contains("share")) %>% melt(id.vars = c("label_industry")) 

lodes$value <- as.numeric(lodes$value)

#Summarise percent by industry, add lodes variable for rbind
lodes_sum <- lodes %>% group_by(label_industry) %>% summarise(percen = mean(value)) %>% mutate(variable = "lodes")

# Check for differences in labels
unique(lodes_sum$label_industry)
unique(veo_bgt_industry_profile$label_industry_compare)

# Add compared labels

labels <- unique(veo_bgt_industry_profile$label_industry_compare) %>% sort()

lodes_sum <- lodes_sum %>% mutate(label_industry_compare = labels)
# Check for matches, looks good!
# Remove old industry labels for rbind
lodes_sum <- lodes_sum %>% select(-label_industry)

# Deselect perc column
veo_bgt_industry_profile <- veo_bgt_industry_profile %>% select(-c(perc, value))

# Rbind
all_industry_profile <- rbind(veo_bgt_industry_profile, lodes_sum) 

# Add bgt percent reference column, calculate bgt percent difference
all_industry_profile <- all_industry_profile %>% left_join(select(res_mil_labels, perc, label_industry_compare), by = "label_industry_compare") %>% rename(bgt_percent = perc) %>% mutate(diff = percen - bgt_percent) %>% mutate(percent_diff = abs(diff)/percen)


ggplot(all_industry_profile, aes(filter(variable != bgt), x = label_industry_compare, y = diff, fill = variable)) +
  geom_col(position = "dodge") +
  labs(title = "Combined Industry and Employment Percents Difference from BGT") +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_classic()


ggplot(lodes, aes(x = label_industry, y = value, fill = label_industry)) +
  geom_col(position = "dodge", show.legend = FALSE) +
  labs(title = "LODES Industry and Employment Percents") +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_classic()

uvapal1 = c("#232D4B","#0E879C","#E6CE3A","#E57200")
uvapal2 = c("#2C4F6B", "#E57200")

plot <- all_industry_profile %>%
  filter(variable != "lodes") %>% 
  mutate(diff_lab = ifelse(diff == 0, NA, (paste0(round(percent_diff*100, digits = 1),"%")))) %>%
  mutate(position = ifelse(variable == "y1_emp", 3,
                           ifelse(variable == "y5_emp", 2,
                                  ifelse(variable == "y10_emp", 1, 4)))) %>%
  arrange(position) %>%
  ggplot(aes(x = reorder(label_industry_compare, percen), y = percen, 
             fill = reorder(variable, position))) +
  geom_col(position = "dodge") +
  labs(title = "Benchmarking BGT: Estimates of Veteran Employment by Industry",
       subtitle = "Comparison of BGT subset and VEO one, five, and ten years after discharge",
       fill = "Estimate source",
       caption = "Percent labels indicate percent difference between estimate source and BGT
Note: BGT is subset to DC MSA and military specific *ONET occupations while VEO data is nationally representative of all Army service members
       Data from BGT (pulled 2015-2018) and VEO (2000-2015)") +
  xlab("NAICS Industry") +
  ylab("") +
  scale_y_continuous(labels = scales::percent) +
  geom_text(aes(x = label_industry_compare,
                y = percen,
                label = diff_lab, 
                fill = reorder(variable, position)),
            position = position_dodge(width = 1),
            hjust = -0.5, size = 2.3, fontface = "plain") +
  scale_fill_manual(values = uvapal1, labels = c("VE0 10 year", "VEO 5 year", "VEO 1 year", "BGT"), guide = guide_legend(reverse = TRUE)) +
  coord_flip() +
  theme_classic()


plot + theme(legend.justification = c(1,0), legend.position = c(1,0))


plot2 <- all_industry_profile %>%
  filter(!(variable %in% c("y1_emp", "y5_emp", "y10_emp"))) %>% 
  mutate(diff_lab = ifelse(diff == 0, NA, (paste0(round(percent_diff*100, digits = 1),"%")))) %>%
  mutate(position = ifelse(variable == "bgt", 2, 1)) %>%
  arrange(position) %>%
  ggplot(aes(x = reorder(label_industry_compare, percen), y = percen, 
             fill = reorder(variable, position))) +
  geom_col(position = position_dodge2(width = 1, preserve = "single")) +
  labs(title = "Benchmarking BGT: Estimates of DC MSA Employment by Industry",
       subtitle = "Comparison of BGT subset and LODES data for all jobs in DC MSA",
       fill = "Estimate source",
       caption = "Percent labels indicate percent difference between estimate source and BGT
Note: BGT is subset to DC MSA and military specific *ONET occupations while LODES data represents all workers in DC MSA
       Data from BGT (pulled 2015-2018) and LODES (2002-2017)") +
  xlab("NAICS Industry") +
  ylab("") +
  scale_y_continuous(labels = scales::percent) +
  geom_text(aes(x = label_industry_compare,
                y = percen,
                label = diff_lab, 
                fill = reorder(variable, position)),
            position = position_dodge2(width = 1, preserve = "single"),
            hjust = -0.5, size = 2, fontface = "plain") +
  scale_fill_manual(values = uvapal2, labels = c("LODES", "BGT"), 
                    guide = guide_legend(reverse = TRUE)) +
  coord_flip() +
  theme_classic()  

plot2 + theme(legend.justification = c(1,0), legend.position = c(1,0))
```

