---
title: "Lodes OD Accquistion"
author: "Owen Hart"
date: "7/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading in libraries
```{r}
library(R.utils)
library(data.table)
library(tidyr)
library(dplyr)
```

Downloading main lodes od
```{r}
#2017
download.file(
  url = "https://lehd.ces.census.gov/data/lodes/LODES7/or/od/or_od_main_JT00_2017.csv.gz", 
  destfile = "../data/lodes/or_od_main_JT00_2017.csv.gz")

gunzip("../data/lodes/or_od_main_JT00_2017.csv.gz")

#2016
download.file(
  url = "https://lehd.ces.census.gov/data/lodes/LODES7/or/od/or_od_main_JT00_2016.csv.gz", 
  destfile = "../data/lodes/or_od_main_JT00_2016.csv.gz")

gunzip("../data/lodes/or_od_main_JT00_2016.csv.gz")

#2015
download.file(
  url = "https://lehd.ces.census.gov/data/lodes/LODES7/or/od/or_od_main_JT00_2015.csv.gz", 
  destfile = "../data/lodes/or_od_main_JT00_2015.csv.gz")

gunzip("../data/lodes/or_od_main_JT00_2015.csv.gz")
```

Downloading aux lodes od
```{r}
download.file(
  url = "https://lehd.ces.census.gov/data/lodes/LODES7/or/od/or_od_aux_JT00_2017.csv.gz", 
  destfile = "../data/lodes/or_od_aux_JT00_2017.csv.gz")

gunzip("../data/lodes/or_od_aux_JT00_2017.csv.gz")


download.file(
  url = "https://lehd.ces.census.gov/data/lodes/LODES7/or/od/or_od_aux_JT00_2016.csv.gz", 
  destfile = "../data/lodes/or_od_aux_JT00_2016.csv.gz")

gunzip("../data/lodes/or_od_aux_JT00_2016.csv.gz")


download.file(
  url = "https://lehd.ces.census.gov/data/lodes/LODES7/or/od/or_od_aux_JT00_2015.csv.gz", 
  destfile = "../data/lodes/or_od_aux_JT00_2015.csv.gz")

gunzip("../data/lodes/or_od_aux_JT00_2015.csv.gz")
```

Reading in files
```{r}
main17 <- fread("../data/lodes/or_od_main_JT00_2017.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))
aux17 <- fread("../data/lodes/or_od_aux_JT00_2017.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))
full_data17 <- rbind(main17, aux17)

main16 <- fread("../data/lodes/or_od_main_JT00_2016.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))
aux16 <- fread("../data/lodes/or_od_aux_JT00_2016.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))
full_data16 <- rbind(main16, aux16)

main15 <- fread("../data/lodes/or_od_main_JT00_2015.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))
aux15 <- fread("../data/lodes/or_od_aux_JT00_2015.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))
full_data15 <- rbind(main15, aux15)
```

Filtering so that it only includes Wasco County 
----------
ADD NEIGHBORHING COUNTIES!!!
```{r}
wasco17_od <- full_data17 %>% 
  filter(w_geocode %in% full_data17$w_geocode[substr(full_data17$w_geocode, 1, 5) == "41065"]|h_geocode %in% full_data17$h_geocode[substr(full_data17$h_geocode, 1, 5) == "41065"])


wasco16_od <- full_data16 %>% 
  filter(w_geocode %in% full_data16$w_geocode[substr(full_data16$w_geocode, 1, 5) == "41065"]|h_geocode %in% full_data16$h_geocode[substr(full_data16$h_geocode, 1, 5) == "41065"])

wasco15_od <- full_data15 %>% 
  filter(w_geocode %in% full_data15$w_geocode[substr(full_data15$w_geocode, 1, 5) == "41065"]|h_geocode %in% full_data15$h_geocode[substr(full_data15$h_geocode, 1, 5) == "41065"])
```

##stop here and save, then agg differently based on usecase

```{r}
saveRDS(wasco17_od, "../data/lodes/filtered/wasco17_od.Rds")
saveRDS(wasco16_od, "../data/lodes/filtered/wasco16_od.Rds")
saveRDS(wasco15_od, "../data/lodes/filtered/wasco15_od.Rds")
```

Loading in WA Data for Outflows plot (neighboring counties in WA)

```{r}
download.file(
  url = "https://lehd.ces.census.gov/data/lodes/LODES7/wa/od/wa_od_main_JT00_2017.csv.gz", 
  destfile = "../data/lodes/wa_od_main_JT00_2017.csv.gz")

gunzip("../data/lodes/wa_od_main_JT00_2017.csv.gz")

#2016
download.file(
  url = "https://lehd.ces.census.gov/data/lodes/LODES7/wa/od/wa_od_main_JT00_2016.csv.gz", 
  destfile = "../data/lodes/wa_od_main_JT00_2016.csv.gz")

gunzip("../data/lodes/wa_od_main_JT00_2016.csv.gz")

#2015
download.file(
  url = "https://lehd.ces.census.gov/data/lodes/LODES7/wa/od/wa_od_main_JT00_2015.csv.gz", 
  destfile = "../data/lodes/wa_od_main_JT00_2015.csv.gz")

gunzip("../data/lodes/wa_od_main_JT00_2015.csv.gz")

###Aux

download.file(
  url = "https://lehd.ces.census.gov/data/lodes/LODES7/wa/od/wa_od_aux_JT00_2017.csv.gz", 
  destfile = "../data/lodes/wa_od_aux_JT00_2017.csv.gz")

gunzip("../data/lodes/wa_od_aux_JT00_2017.csv.gz")


download.file(
  url = "https://lehd.ces.census.gov/data/lodes/LODES7/wa/od/wa_od_aux_JT00_2016.csv.gz", 
  destfile = "../data/lodes/wa_od_aux_JT00_2016.csv.gz")

gunzip("../data/lodes/wa_od_aux_JT00_2016.csv.gz")


download.file(
  url = "https://lehd.ces.census.gov/data/lodes/LODES7/wa/od/wa_od_aux_JT00_2015.csv.gz", 
  destfile = "../data/lodes/wa_od_aux_JT00_2015.csv.gz")

gunzip("../data/lodes/wa_od_aux_JT00_2015.csv.gz")
```

```{r}
main17 <- fread("../data/lodes/wa_od_main_JT00_2017.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))
aux17 <- fread("../data/lodes/wa_od_aux_JT00_2017.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))
full_data17 <- rbind(main17, aux17)

main16 <- fread("../data/lodes/wa_od_main_JT00_2016.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))
aux16 <- fread("../data/lodes/wa_od_aux_JT00_2016.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))
full_data16 <- rbind(main16, aux16)

main15 <- fread("../data/lodes/wa_od_main_JT00_2015.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))
aux15 <- fread("../data/lodes/wa_od_aux_JT00_2015.csv", 
              colClasses = c("w_geocode" = "character",
                             "h_geocode" = "character"))
full_data15 <- rbind(main15, aux15)
```



filtered so workplace is in wasco
```{r}
wa17_od <- full_data17 %>% 
  filter(h_geocode %in% full_data17$h_geocode[substr(full_data17$h_geocode, 1, 5) == "41065"])

wa16_od <- full_data16 %>% 
  filter(h_geocode %in% full_data16$h_geocode[substr(full_data16$h_geocode, 1, 5) == "41065"])

wa15_od <- full_data15 %>% 
  filter(h_geocode %in% full_data15$h_geocode[substr(full_data15$h_geocode, 1, 5) == "41065"])
```



save to filtered to load into lodes_od_plots.Rmd when doing outflows
```{r}
saveRDS(wa17_od, "../data/lodes/filtered/wa17_od.Rds")
saveRDS(wa16_od, "../data/lodes/filtered/wa16_od.Rds")
saveRDS(wa15_od, "../data/lodes/filtered/wa15_od.Rds")
```

