---
title: "Lodes OD Plots"
author: "Owen Hart"
date: "7/15/2020"
output: html_document
---

```{r}
library(R.utils)
library(data.table)
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(sf)
library(leaflet)
library(reshape2)
library(raster)
library(tigris)
library(viridis)
library(plotly)
```

reading in filtered lodes data with geometries
```{r}
wasco17_od <- readRDS("~/git/dspg20wasco/data/lodes/filtered/wasco17_od.Rds")
wasco16_od <- readRDS("~/git/dspg20wasco/data/lodes/filtered/wasco16_od.Rds")
wasco15_od <- readRDS("~/git/dspg20wasco/data/lodes/filtered/wasco17_od.Rds")

wa17_od <- readRDS("~/git/dspg20wasco/data/lodes/filtered/wa17_od.Rds")
wa16_od <- readRDS("~/git/dspg20wasco/data/lodes/filtered/wa16_od.Rds")
wa15_od <- readRDS("~/git/dspg20wasco/data/lodes/filtered/wa15_od.Rds")

pnw_wasco17_od <- rbind(wasco17_od, wa17_od)
pnw_wasco16_od <- rbind(wasco16_od, wa16_od)
pnw_wasco15_od <- rbind(wasco15_od, wa15_od)

```

```{r}
# select tracts from geocode strings
pnw_wasco17_od$w_geocode_tract <- substr(pnw_wasco17_od$w_geocode, start = 1, stop = 11)
pnw_wasco17_od$h_geocode_tract <- substr(pnw_wasco17_od$h_geocode, start = 1, stop = 11)

pnw_wasco16_od$w_geocode_tract <- substr(pnw_wasco16_od$w_geocode, start = 1, stop = 11)
pnw_wasco16_od$h_geocode_tract <- substr(pnw_wasco16_od$h_geocode, start = 1, stop = 11)

pnw_wasco15_od$w_geocode_tract <- substr(pnw_wasco15_od$w_geocode, start = 1, stop = 11)
pnw_wasco15_od$h_geocode_tract <- substr(pnw_wasco15_od$h_geocode, start = 1, stop = 11)
```

aggregate by tract level
```{r}
jobs_all_2017 <- pnw_wasco17_od %>%
  group_by(createdate, w_geocode_tract, h_geocode_tract) %>% 
  summarise(S000 = sum(S000), 
            SA01 = sum(SA01), 
            SA02 = sum(SA02), 
            SA03 = sum(SA03), 
            SE01 = sum(SE01), 
            SE02 = sum(SE02), 
            SE03 = sum(SE03), 
            SI01 = sum(SI01), 
            SI02 = sum(SI02), 
            SI03 = sum(SI03))
jobs_all_2016 <- pnw_wasco16_od %>%
  group_by(createdate, w_geocode_tract, h_geocode_tract) %>% 
  summarise(S000 = sum(S000), 
            SA01 = sum(SA01), 
            SA02 = sum(SA02), 
            SA03 = sum(SA03), 
            SE01 = sum(SE01), 
            SE02 = sum(SE02), 
            SE03 = sum(SE03), 
            SI01 = sum(SI01), 
            SI02 = sum(SI02), 
            SI03 = sum(SI03))
jobs_all_2015 <- pnw_wasco15_od %>%
  group_by(createdate, w_geocode_tract, h_geocode_tract) %>% 
  summarise(S000 = sum(S000), 
            SA01 = sum(SA01), 
            SA02 = sum(SA02), 
            SA03 = sum(SA03), 
            SE01 = sum(SE01), 
            SE02 = sum(SE02), 
            SE03 = sum(SE03), 
            SI01 = sum(SI01), 
            SI02 = sum(SI02), 
            SI03 = sum(SI03))
```

```{r}
jobs_all_2017$h_state <- substr(jobs_all_2017$h_geocode_tract, 1, 2)
jobs_all_2017$h_county <- substr(jobs_all_2017$h_geocode_tract, 3, 5)
jobs_all_2017$h_tract <- substr(jobs_all_2017$h_geocode_tract, 6, 11)
jobs_all_2017$h_stcty <- substr(jobs_all_2017$h_geocode_tract, 1, 5)

jobs_all_2016$h_state <- substr(jobs_all_2016$h_geocode_tract, 1, 2)
jobs_all_2016$h_county <- substr(jobs_all_2016$h_geocode_tract, 3, 5)
jobs_all_2016$h_tract <- substr(jobs_all_2016$h_geocode_tract, 6, 11)
jobs_all_2016$h_stcty <- substr(jobs_all_2016$h_geocode_tract, 1, 5)

jobs_all_2015$h_state <- substr(jobs_all_2015$h_geocode_tract, 1, 2)
jobs_all_2015$h_county <- substr(jobs_all_2015$h_geocode_tract, 3, 5)
jobs_all_2015$h_tract <- substr(jobs_all_2015$h_geocode_tract, 6, 11)
jobs_all_2015$h_stcty <- substr(jobs_all_2015$h_geocode_tract, 1, 5)

#now work geoid
jobs_all_2017$w_state <- substr(jobs_all_2017$w_geocode_tract, 1, 2)
jobs_all_2017$w_county <- substr(jobs_all_2017$w_geocode_tract, 3, 5)
jobs_all_2017$w_tract <- substr(jobs_all_2017$w_geocode_tract, 6, 11)
jobs_all_2017$w_stcty <- substr(jobs_all_2017$w_geocode_tract, 1, 5)


jobs_all_2016$w_state <- substr(jobs_all_2016$w_geocode_tract, 1, 2)
jobs_all_2016$w_county <- substr(jobs_all_2016$w_geocode_tract, 3, 5)
jobs_all_2016$w_tract <- substr(jobs_all_2016$w_geocode_tract, 6, 11)
jobs_all_2016$w_stcty <- substr(jobs_all_2016$w_geocode_tract, 1, 5)

jobs_all_2015$w_state <- substr(jobs_all_2015$w_geocode_tract, 1, 2)
jobs_all_2015$w_county <- substr(jobs_all_2015$w_geocode_tract, 3, 5)
jobs_all_2015$w_tract <- substr(jobs_all_2015$w_geocode_tract, 6, 11)
jobs_all_2015$w_stcty <- substr(jobs_all_2015$w_geocode_tract, 1, 5)
```

Must load in Oregon XWalk csv
```{r}
#download.file(
#  url = "https://lehd.ces.census.gov/data/lodes/LODES7/or/or_xwalk.csv.gz", 
#  destfile = "../data/lodes/or_xwalk.csv.gz")
#gunzip("../data/lodes/or_xwalk.csv.gz")
or_xw <- read_csv("~/git/dspg20wasco/data/lodes/or_xwalk.csv", col_types = cols(cty = col_character()))
wa_xw <- read_csv("~/git/dspg20wasco/data/lodes/wa_xwalk.csv", col_types = cols(cty = col_character()))
wa_xw <- wa_xw[(wa_xw$ctyname %like% "Klickitat") | (wa_xw$ctyname %like% "Skamania"), ]
or_wa_xw <- rbind(or_xw, wa_xw)
```
What % of jobs flow from outside of Fairfax County?
```{r}
bool_cond17 <-jobs_all_2017$h_stcty != "41065" & (jobs_all_2017$w_stcty == "41065")
bool_cond16 <-jobs_all_2016$h_stcty != "41065" & (jobs_all_2016$w_stcty == "41065")
bool_cond15 <-jobs_all_2015$h_stcty!= "41065" & (jobs_all_2015$w_stcty == "41065")
flows_in_17 <-jobs_all_2017[bool_cond17, ]
flows_in_16 <-jobs_all_2016[bool_cond16, ]
flows_in_15 <-jobs_all_2015[bool_cond15, ]
```

Of the flows into Wasco County from outside, what percent are from other counties in Oregon/Washington?
```{r}
agg_cnty_2017 <- flows_in_17[flows_in_17$h_state == "41" | flows_in_17$h_state == "53", ] %>%
  group_by(h_stcty) %>%
  summarise(S000 = sum(S000))
agg_cnty_with_names_2017 <- inner_join(x = agg_cnty_2017, y = unique(or_wa_xw %>% dplyr::select(cty, ctyname)), by = c("h_stcty" = "cty"))
trimmed_2017 <- agg_cnty_with_names_2017[order(-agg_cnty_with_names_2017$S000), ][1:12, ]

agg_cnty_2016 <- flows_in_16[flows_in_16$h_state == "41" | flows_in_16$h_state == "53", ] %>%
  group_by(h_stcty) %>%
  summarise(S000 = sum(S000))
agg_cnty_with_names_2016 <- inner_join(x = agg_cnty_2016, y = unique(or_wa_xw %>% dplyr::select(cty, ctyname)), by = c("h_stcty" = "cty"))
trimmed_2016 <- agg_cnty_with_names_2016[order(-agg_cnty_with_names_2016$S000), ][1:12, ]

agg_cnty_2015 <- flows_in_15[flows_in_15$h_state == "41" | flows_in_15$h_state == "53", ] %>%
  group_by(h_stcty) %>%
  summarise(S000 = sum(S000))
agg_cnty_with_names_2015 <- inner_join(x = agg_cnty_2015, y = unique(or_wa_xw %>% dplyr::select(cty, ctyname)), by = c("h_stcty" = "cty"))
trimmed_2015 <- agg_cnty_with_names_2015[order(-agg_cnty_with_names_2015$S000), ][1:12, ]
```
Aggregating by county over time (line plot)
```{r}
top_12_in_over_time <- inner_join(trimmed_2017, agg_cnty_with_names_2016, by = "ctyname")
top_12_in_over_time <- top_12_in_over_time %>%
  dplyr::select(h_stcty.x, ctyname, S000.x, S000.y)
top_12_in_over_time$'2017' <- top_12_in_over_time$S000.x
top_12_in_over_time$'2016' <- top_12_in_over_time$S000.y
top_12_in_over_time <- inner_join(top_12_in_over_time, agg_cnty_with_names_2015, by = "ctyname") %>%
  dplyr::select(h_stcty.x, ctyname, '2017', '2016', S000)

top_12_in_over_time$'2015' <- top_12_in_over_time$S000
top_12_in_over_time$h_stcty <- top_12_in_over_time$h_stcty.x
top_12_in_over_time <- top_12_in_over_time %>% dplyr::select(-S000, -h_stcty.x)
top_12_in_over_time <- top_12_in_over_time %>% dplyr::select(ctyname, '2017', '2016', '2015')
top_12_in_trans <- data.frame(t(top_12_in_over_time[, -1]))
colnames(top_12_in_trans) <- top_12_in_over_time$ctyname
top_12_in_trans$year <- c(2017, 2016, 2015) #SAVE NOW
write_csv(top_12_in_trans, "../data/app_12_inflows_wasco.csv")

#line plot for all jobs 2015-2017 by county
ggplotly(ggplot(top_12_in_trans, aes(x = year)) +
  ggtitle("Number of jobs flowing into Wasco County (2015-2017)") +
  labs(x = "Year", y = "Number of Jobs", colour = "County") + 
  geom_line(aes(y = `Hood River County, OR`, color = "Hood River County")) + 
  geom_line(aes(y = `Klickitat County, WA`, color = "Klickitat County, WA")) + 
  geom_line(aes(y = `Multnomah County, OR`, color = "Multnomah County")) + 
  geom_line(aes(y = `Clackamas County, OR`, color = "Clackamas County")) + 
  geom_line(aes(y = `Marion County, OR`, color = "Marion County")) + 
  geom_line(aes(y = `Washington County, OR`, color = "Washington County")) + 
  geom_line(aes(y = `Deschutes County, OR`, color = "Deschutes County")) + 
  geom_line(aes(y = `Jefferson County, OR`, color = "Jefferson County")) + 
  geom_line(aes(y = `Lane County, OR`, color = "Lane County")) + 
  geom_line(aes(y = `Umatilla County, OR`, color = "Umatilla County")) + 
  geom_line(aes(y = `Sherman County, OR`, color = "Sherman County")) +
  geom_line(aes(y = `Skamania County, WA`, color = "Skamania County, WA")) +
  scale_x_continuous(breaks = 0:2100))
```




Outflows (those who live in wasco county but do not work there, open to other states)
```{r}
#want those who work in washington to live in wasco, and those who live in wasco to not work in wasco
bool_cond17 <-(jobs_all_2017$h_stcty == "41065" & jobs_all_2017$w_state == "53") | (jobs_all_2017$h_stcty == "41065" & jobs_all_2017$w_stcty != "41065")
bool_cond16 <-(jobs_all_2016$h_stcty == "41065" & jobs_all_2016$w_state == "53") | (jobs_all_2016$h_stcty == "41065" & jobs_all_2016$w_stcty != "41065")
bool_cond15 <-(jobs_all_2015$h_stcty == "41065" & jobs_all_2015$w_state == "53") | (jobs_all_2015$h_stcty == "41065" & jobs_all_2015$w_stcty != "41065")

flows_out_17 <-jobs_all_2017[bool_cond17, ]
flows_out_16 <-jobs_all_2016[bool_cond16, ]
flows_out_15 <-jobs_all_2015[bool_cond15, ]
```

Most popular counties that those who live in Wasco County but not work in, go to work
```{r}
out_agg_2017 <- flows_out_17 %>%
  group_by(w_stcty) %>%
  summarise(S000 = sum(S000))
essential <- or_wa_xw %>%
  dplyr::select(cty, ctyname)
agg_out_names_2017 <- inner_join(x = out_agg_2017, y = unique(essential), by = c("w_stcty" = "cty"))
trimmed_out_2017 <- rbind(agg_out_names_2017[order(-agg_out_names_2017$S000), ][1:11, ],
                          agg_out_names_2017[order(-agg_out_names_2017$S000), ][24, ])
out_agg_2016 <- flows_out_16 %>%
  group_by(w_stcty) %>%
  summarise(S000 = sum(S000))
essential <- or_wa_xw %>%
  dplyr::select(cty, ctyname)
agg_out_names_2016 <- inner_join(x = out_agg_2016, y = unique(essential), by = c("w_stcty" = "cty"))
trimmed_out_2016 <- rbind(agg_out_names_2016[order(-agg_out_names_2016$S000), ][1:11, ],
                          agg_out_names_2016[order(-agg_out_names_2016$S000), ][23, ])
out_agg_2015 <- flows_out_15 %>%
  group_by(w_stcty) %>%
  summarise(S000 = sum(S000))
essential <- or_wa_xw %>%
  dplyr::select(cty, ctyname)
agg_out_names_2015 <- inner_join(x = out_agg_2015, y = unique(essential), by = c("w_stcty" = "cty"))
trimmed_out_2015 <- rbind(agg_out_names_2015[order(-agg_out_names_2015$S000), ][1:11, ],
                          agg_out_names_2015[order(-agg_out_names_2015$S000), ][15, ])
```

outflows over time!
```{r}
top_12_out_over_time <- inner_join(trimmed_out_2017, agg_out_names_2016, by = "ctyname")
top_12_out_over_time <- top_12_out_over_time %>%
  dplyr::select(w_stcty.x, ctyname, S000.x, S000.y)
top_12_out_over_time$'2017' <- top_12_out_over_time$S000.x
top_12_out_over_time$'2016' <- top_12_out_over_time$S000.y
top_12_out_over_time <- inner_join(top_12_out_over_time, agg_out_names_2015, by = "ctyname") %>%
  dplyr::select(w_stcty.x, ctyname, '2017', '2016', S000)
top_12_out_over_time$'2015' <- top_12_out_over_time$S000
top_12_out_over_time$h_county <- top_12_out_over_time$w_stcty.x
top_12_out_over_time <- top_12_out_over_time %>% dplyr::select(-S000, -w_stcty.x)
top_12_out_over_time <- top_12_out_over_time %>% dplyr::select(ctyname, '2017', '2016', '2015')
top_12_out_trans <- data.frame(t(top_12_out_over_time[, -1]))
colnames(top_12_out_trans) <- top_12_out_over_time$ctyname
top_12_out_trans$year <- c(2017, 2016, 2015)
write_csv(top_12_out_trans, "../data/app_12_outflows_wasco.csv")
top_12_out_trans

ggplotly(ggplot(top_12_out_trans, aes(x = year)) +
  ggtitle("Number of jobs flowing out of Wasco County (2015-2017)") +
  labs(x = "Year", y = "Number of Jobs", colour = "County") + 
  geom_line(aes(y = `Hood River County, OR`, color = "Hood River County")) + 
  geom_line(aes(y = `Multnomah County, OR`, color = "Multnomah County")) + 
  geom_line(aes(y = `Clackamas County, OR`, color = "Clackamas County")) + 
  geom_line(aes(y = `Klickitat County, WA`, color = "Klickitat County, WA")) + 
  geom_line(aes(y = `Deschutes County, OR`, color = "Deschutes County")) + 
  geom_line(aes(y = `Washington County, OR`, color = "Washington County")) + 
  geom_line(aes(y = `Marion County, OR`, color = "Marion County")) + 
  geom_line(aes(y = `Jefferson County, OR`, color = "Jefferson County")) + 
  geom_line(aes(y = `Umatilla County, OR`, color = "Umatilla County")) + 
  geom_line(aes(y = `Lane County, OR`, color = "Lane County")) + 
  geom_line(aes(y = `Sherman County, OR`, color = "Sherman County")) +
  geom_line(aes(y = `Skamania County, WA`, color = "Skamania County, WA")) +
  scale_x_continuous(breaks = 0:2100))

#ggsave(path = "../output/lodes_maps/flows/out_flows", device = "png", filename = "out_county_flows_all_years.png", plot = last_plot())
```


Mapping S. Wasco
```{r}
wasco_points <- blocks("OR", county = "Wasco")
wasco_lines <- data.frame(wasco_points)
south_wasco_points <- st_read("../data/shps/swsd")
wasco_geo_points <- st_read("../data/shps/county")
```

Reading in filtered file again
```{r}
wasco17_od2 <- readRDS("../data/lodes/filtered/wasco17_od.Rds")
wasco16_od2 <- readRDS("../data/lodes/filtered/wasco16_od.Rds")
wasco15_od2 <- readRDS("../data/lodes/filtered/wasco17_od.Rds")
```

Joining with Wasco Lines
```{r}
wascolines_17 <- inner_join(wasco17_od2, wasco_lines, by = c("w_geocode" = "GEOID10"))
wascolines_16 <- inner_join(wasco16_od2, wasco_lines, by = c("w_geocode" = "GEOID10"))
wascolines_15 <- inner_join(wasco15_od2, wasco_lines, by = c("w_geocode" = "GEOID10"))
```

Grouping by work geocode (because that is going to be Wasco County)
```{r}
agg_17 <- st_as_sf(wascolines_17) %>%
  group_by(w_geocode, INTPTLAT10, INTPTLON10) %>%
  summarise(S000 = sum(S000), 
            SA01 = sum(SA01), 
            SA02 = sum(SA02), 
            SA03 = sum(SA03), 
            SE01 = sum(SE01), 
            SE02 = sum(SE02), 
            SE03 = sum(SE03), 
            SI01 = sum(SI01), 
            SI02 = sum(SI02), 
            SI03 = sum(SI03))
saveRDS(agg_17, "../data/app_lodes_od_agg_2017.Rds")
agg_16 <- st_as_sf(wascolines_16) %>%
  group_by(w_geocode, INTPTLAT10, INTPTLON10) %>%
  summarise(S000 = sum(S000), 
            SA01 = sum(SA01), 
            SA02 = sum(SA02), 
            SA03 = sum(SA03), 
            SE01 = sum(SE01), 
            SE02 = sum(SE02), 
            SE03 = sum(SE03), 
            SI01 = sum(SI01), 
            SI02 = sum(SI02), 
            SI03 = sum(SI03))
saveRDS(agg_16, "../data/app_lodes_od_agg_2016.Rds")
agg_15 <- st_as_sf(wascolines_15) %>%
  group_by(w_geocode, INTPTLAT10, INTPTLON10) %>%
  summarise(S000 = sum(S000), 
            SA01 = sum(SA01), 
            SA02 = sum(SA02), 
            SA03 = sum(SA03), 
            SE01 = sum(SE01), 
            SE02 = sum(SE02), 
            SE03 = sum(SE03), 
            SI01 = sum(SI01), 
            SI02 = sum(SI02), 
            SI03 = sum(SI03))
saveRDS(agg_15, "../data/app_lodes_od_agg_2015.Rds")
```

interactive maps
```{r}
#S000 (all jobs) by year
od_S000leaf <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolylines(
    data = wasco_geo_points,
    color = "black",
    weight = 1,
    group = "Basemap",
    label = "Wasco County") %>%
  addPolylines(
    data = south_wasco_points,
    color = "purple",
    weight = 2,
    opacity = 1,
    group = "Basemap",
    label = "South Wasco Region") %>%
  addPolygons(
    data = st_as_sf(agg_17),
    weight = 1,
    opacity = 0,
    fillOpacity = .7,
    group = "2017",
    fillColor = ~colorQuantile(viridis_pal(option = "D")(5), domain = agg_17$S000)(agg_17$S000),
    label = ~lapply(paste(sep = "", 
                          "<strong> Number of All Jobs: <strong>", 
                          agg_17$S000), htmltools::HTML)) %>%
  addLegend(
    data = rbind(agg_17, agg_16, agg_15),
    "bottomright",
    pal = colorQuantile(viridis_pal(option = "D")(5), domain = rbind(agg_17, agg_16, agg_15)$S000),
    values = ~ S000,
    labFormat = function(type, cuts, p) {
      n = length(cuts)
      p = paste0(round(p * 100), '%')
      cuts = paste0(formatC(cuts[-n]), " - ", formatC(cuts[-1]))},
    title = "Number of All Jobs<br>by Census Tract<br>in Wasco County",
    na.label = "NA") %>%
  addPolygons(
    data = st_as_sf(agg_16),
    weight = 1,
    opacity = 0,
    fillOpacity = .7,
    group = "2016",
    fillColor = ~colorQuantile(viridis_pal(option = "D")(5), domain = agg_16$S000)(agg_16$S000),
    label = ~lapply(paste(sep = "", 
                          "<strong> Number of All Jobs: <strong>", 
                          agg_16$S000), htmltools::HTML)) %>%
  addPolygons(
    data = st_as_sf(agg_15),
    weight = 1,
    opacity = 0,
    fillOpacity = .7,
    group = "2015",
    fillColor = ~colorQuantile(viridis_pal(option = "D")(5), domain = agg_15$S000)(agg_15$S000),
    label = ~lapply(paste(sep = "", 
                          "<strong> Number of All Jobs: <strong>", 
                          agg_15$S000), htmltools::HTML)) %>%
  addLayersControl(
    baseGroups = c("2017", "2016", "2015"),
    options = layersControlOptions(collapsed = FALSE)) %>%
  hideGroup(c("2016", "2015"))
```

```{r}
#SI01 (Goods Producing industry sectors) by year
colors_SI01 <- colorQuantile(viridis_pal(option = "D")(3), domain = unique(rbind(agg_17, agg_16, agg_15)$SI01))

od_SI01leaf <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolylines(
    data = wasco_geo_points,
    color = "black",
    weight = 1,
    group = "Basemap",
    label = "Wasco County") %>%
  addPolylines(
    data = south_wasco_points,
    color = "purple",
    weight = 2,
    opacity = 1,
    group = "Basemap",
    label = "South Wasco Region") %>%
  addPolygons(
    data = st_as_sf(agg_17),
    weight = 1,
    opacity = 0,
    fillOpacity = .7,
    group = "2017",
    fillColor = ~colors_SI01((agg_17$SI01)),
    label = ~lapply(paste(sep = "", 
                          "<strong> Number of Goods Producing Jobs: <strong>", 
                          agg_17$SI01), htmltools::HTML)) %>%
  addLegend(
    data = rbind(agg_17, agg_16, agg_15),
    "bottomright",
    pal = colors_SI01,
    values = ~ unique(SI01),
    labFormat = function(type, cuts, p) {
      n = length(cuts)
      p = paste0(round(p * 100), '%')
      cuts = paste0(formatC(cuts[-n]), " - ", formatC(cuts[-1]))},
    title = "Number of Goods Producing<br> Jobs by Census Tract<br>in Wasco County",
    na.label = "NA") %>%
  addPolygons(
    data = st_as_sf(agg_16),
    weight = 1,
    opacity = 0,
    fillOpacity = .7,
    group = "2016",
    fillColor = ~colors_SI01((agg_16$SI01)),
    label = ~lapply(paste(sep = "", 
                          "<strong> Number of Goods Producing Jobs: <strong>", 
                          agg_16$SI01), htmltools::HTML)) %>%
  addPolygons(
    data = st_as_sf(agg_15),
    weight = 1,
    opacity = 0,
    fillOpacity = .7,
    group = "2015",
    fillColor = ~colors_SI01((agg_15$SI01)),
    label = ~lapply(paste(sep = "", 
                          "<strong> Number of Goods Producing Jobs: <strong>", 
                          agg_15$SI01), htmltools::HTML)) %>%
  addLayersControl(
    baseGroups = c("2017", "2016", "2015"),
    options = layersControlOptions(collapsed = FALSE)) %>%
  hideGroup(c("2016", "2015"))
```


```{r}
#SI02 (Trade, Transportation, and Utilities industry sectors) by year
colors_SI02 <- colorQuantile(viridis_pal(option = "D")(3), domain = unique(rbind(agg_17, agg_16, agg_15)$SI02))

od_SI02leaf <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolylines(
    data = wasco_geo_points,
    color = "black",
    weight = 1,
    group = "Basemap",
    label = "Wasco County") %>%
  addPolylines(
    data = south_wasco_points,
    color = "purple",
    weight = 2,
    opacity = 1,
    group = "Basemap",
    label = "South Wasco Region") %>%
  addPolygons(
    data = st_as_sf(agg_17),
    weight = 1,
    opacity = 0,
    fillOpacity = .7,
    group = "2017",
    fillColor = ~colors_SI02((agg_17$SI02)),
    label = ~lapply(paste(sep = "", 
                          "<strong> Number of Utility Jobs: <strong>", 
                          agg_17$SI02), htmltools::HTML)) %>%
  addLegend(
    data = rbind(agg_17, agg_16, agg_15),
    "bottomright",
    pal = colors_SI02,
    values = ~ unique(SI02),
    labFormat = function(type, cuts, p) {
      n = length(cuts)
      p = paste0(round(p * 100), '%')
      cuts = paste0(formatC(cuts[-n]), " - ", formatC(cuts[-1]))},
    title = "Number of Trade, Transportation,<br>and Utilities Jobs<br>by Census Tract in<br>
    Wasco County",
    na.label = "NA") %>%
  addPolygons(
    data = st_as_sf(agg_16),
    weight = 1,
    opacity = 0,
    fillOpacity = .7,
    group = "2016",
    fillColor = ~colors_SI02((agg_16$SI02)),
    label = ~lapply(paste(sep = "", 
                          "<strong> Number of Utility Jobs: <strong>", 
                          agg_16$SI02), htmltools::HTML)) %>%
  addPolygons(
    data = st_as_sf(agg_15),
    weight = 1,
    opacity = 0,
    fillOpacity = .7,
    group = "2015",
    fillColor = ~colors_SI02((agg_15$SI02)),
    label = ~lapply(paste(sep = "", 
                          "<strong> Number of Utility Jobs: <strong>", 
                          agg_15$SI02), htmltools::HTML)) %>%
  addLayersControl(
    baseGroups = c("2017", "2016", "2015"),
    options = layersControlOptions(collapsed = FALSE)) %>%
  hideGroup(c("2016", "2015"))
```

```{r}
#SI03 (All Other Services industry sectors) by year
colors_SI03 <- colorQuantile(viridis_pal(option = "D")(3), domain = unique(rbind(agg_17, agg_16, agg_15)$SI03))

od_SI03leaf <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolylines(
    data = wasco_geo_points,
    color = "black",
    weight = 1,
    group = "Basemap",
    label = "Wasco County") %>%
  addPolylines(
    data = south_wasco_points,
    color = "purple",
    weight = 2,
    opacity = 1,
    group = "Basemap",
    label = "South Wasco Region") %>%
  addPolygons(
    data = st_as_sf(agg_17),
    weight = 1,
    opacity = 0,
    fillOpacity = .7,
    group = "2017",
    fillColor = ~colors_SI03((agg_17$SI03)),
    label = ~lapply(paste(sep = "", 
                          "<strong> Number of Other Service Jobs: <strong>", 
                          agg_17$SI03), htmltools::HTML)) %>%
  addLegend(
    data = rbind(agg_17, agg_16, agg_15),
    "bottomright",
    pal = colors_SI03,
    values = ~ unique(SI03),
    labFormat = function(type, cuts, p) {
      n = length(cuts)
      p = paste0(round(p * 100), '%')
      cuts = paste0(formatC(cuts[-n]), " - ", formatC(cuts[-1]))},
    title = "Number of All Other Services<br>Sector Jobs by Census Tract<br>in Wasco County",
    na.label = "NA") %>%
  addPolygons(
    data = st_as_sf(agg_16),
    weight = 1,
    opacity = 0,
    fillOpacity = .7,
    group = "2016",
    fillColor = ~colors_SI03((agg_16$SI03)),
    label = ~lapply(paste(sep = "", 
                          "<strong> Number of Other Service Jobs: <strong>", 
                          agg_16$SI03), htmltools::HTML)) %>%
  addPolygons(
    data = st_as_sf(agg_15),
    weight = 1,
    opacity = 0,
    fillOpacity = .7,
    group = "2015",
    fillColor = ~colors_SI03((agg_15$SI03)),
    label = ~lapply(paste(sep = "", 
                          "<strong> Number of Other Service Jobs: <strong>", 
                          agg_15$SI03), htmltools::HTML)) %>%
  addLayersControl(
    baseGroups = c("2017", "2016", "2015"),
    options = layersControlOptions(collapsed = FALSE)) %>%
  hideGroup(c("2016", "2015"))
```


Showing all maps
```{r}
od_S000leaf
od_SI01leaf
od_SI02leaf
od_SI03leaf
```
