---
title: "14_nonveteran_sequences"
author: "Joanna Schroeder"
date: "9/1/2020"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
for (pkg in c("tidyverse",  "data.table", "R.utils", "maditr", "stringr", "stringi", "dplyr", "ggplot2", "lubridate", "gt", "DataExplorer", "TraMineR", "TraMineRextras", "cluster", "tools", "WeightedCluster", "grid", "gridExtra")) {
  library(pkg, character.only = TRUE)
}

uva_color_palette <- 
c("#232D4B", #space cadet
  "#2C4F6B", #indigo dye
  "#0E879C", #blue munsell
  "#60999A", #cadet blue
  "#D1E0BF", #tea green
  "#D9E12B", #pear
  "#E6CE3A", #citrine
  "#E6A01D", #marigold
  "#E57200" #princeton orange
)
```

## Purpose

We are interested in comparing the results of the veteran sequence analysis and clustering to nonvet
erans. For this analysis we will replicate methods used to subset the veteran population, though a few differences are necessary to account for differences in the datasets.

* The nonveteran data includes many more resumes than the veterans. WeightedCluster must be used for most subsets to aggregate to n < 46,431 (limit of seqdist).
* We are comparing civilian career paths. Veterans have a career state of, "Military transition unemployment," to describe a period of missing resume data between the date of military exit and the first civilian job. For nonveterans, all unemployment falls in the state of, "Civilian unemployment." 

Does it make

```{r prepare-data, include = FALSE}
## Import all resume data -----------------------
bg_all_job <- fread("~/git/DSPG2020/career/data/03_bg_all_job.csv")
#bg_vet_job <- read_csv("~/git/DSPG2020/career/data/04_bg_vet_job.csv")

## Create a variable for year entering the job market ---------------
bg_all_job <- bg_all_job %>%
  mutate(year_enter_job_market = year(date_enter_job_market))%>%
  select(-noofjobs, -sector, -tenure) 

## Create the df for input into seqformat(), create vairbale for year of job start, create a variable for years in job market (used for subsetting), arrange by job zone and take distinct to manually remove overlapping jobs by taking the highest job zone for a year, remove veterans, create variables to align dates for sequence analysis  ----
bg_all_job_seq <- bg_all_job %>%
  select(id, end_year, onet_job_zone, startdate, enddate, job_duration_day) %>%
  mutate(start_year = year(startdate))
bg_all_job_seq <- bg_all_job_seq %>%
  group_by(id) %>%
  mutate(years_in_job_market = max(end_year) - min(start_year)) %>%
  select(id, start_year, end_year, onet_job_zone, years_in_job_market)%>%
  group_by(id)%>%
  arrange(desc(onet_job_zone))%>%
  group_by(id)%>%
  distinct(id, start_year, end_year, .keep_all = TRUE)
sts_all <- bg_all_job_seq %>%
#  filter(years_in_job_market >= 10) %>%
  select(-years_in_job_market) %>%
  group_by(id) %>% mutate(veteran = ifelse(any(onet_job_zone == 55), 1, 0)) %>% ungroup() %>%
  filter(veteran == 0) %>% group_by(id) %>%
  mutate(start_year_a = start_year - min(start_year) + 1) %>%
  mutate(end_year_a = end_year - min(start_year) + 1)

## Used for creating fixed-time subsets, which are written in the data folder ------
# sts_all %>% group_by(id) %>%
#  spread(start_year, onet_job_zone)

#sts_all <- as.matrix(sts_all)
#sts_all <- as.data.frame(sts_all)
  
#length(unique(sts_all$id))

# manual conversion to sequence format
#dcast(sts_all, id ~ start_year_a, value.var = "onet_job_zone")

#process <- sts_all %>% group_by(id) %>% mutate(enter = min(start_year)) %>% select(id, enter) %>% distinct()

#sts_all <- seqformat(sts_all, from = "SPELL", to = "STS",
#            id = "id",  begin = "start_year_a", end = "end_year_a", 
#            status = "onet_job_zone", process = FALSE)

#all.seq <- seqdef(sts_all, left="DEL", gaps="Civilian unemployment", right="DEL", cpal = c("#CBBEB5",  uva_color_palette[1], uva_color_palette[2],uva_color_palette[4],  uva_color_palette[6], uva_color_palette[9], "#CBBEB5"))

# Filter out veterans who do not have a complete career in time period of interest
#sts_all <- all.seq[, 1:10]

#write.csv(sts_all, "~/git/DSPG2020/career/data/sts_all_ten.csv")

## Function to create sequence object with cohort year filters ------------------
career_sequence_exit_year_non <- function(name = "sts_vet_ten", enter_year, exit_year) {
sts_vet <- sts_all %>%
  filter(start_year >= exit_year & start_year <= enter_year) #%>%
  #select(-years_after_onet55)

sts_vet <- as.matrix(sts_vet)
sts_vet <- as.data.frame(sts_vet)
  
sts_vet <- seqformat(sts_vet, from = "SPELL", to = "STS",
            id = "id",  begin = "start_year", end = "end_year", 
            status = "onet_job_zone", process = FALSE)

sts_vet <- seqdef(sts_vet, left="DEL", gaps="Civilian unemployment", right="DEL", cpal = c("#CBBEB5",  uva_color_palette[1], uva_color_palette[2],uva_color_palette[4],  uva_color_palette[6], uva_color_palette[9], "#CBBEB5"))

names(sts_vet) <- paste0(1:ncol(sts_vet))

assign(name, sts_vet, envir = .GlobalEnv)
}

## Creating sequence objects for 1988 and 1964 cohorts (a computational time decision) ------
career_sequence_exit_year_non("sts_non_1988", 1997, 1988)
career_sequence_exit_year_non("sts_non_1964", 1987, 1964)

## Read in fixed-time sequences, need to redefine as sequence objects -------------
sts_nonvet_ten <- read.csv("~/git/DSPG2020/career/data/sts_all_ten.csv", row.names = 1)
sts_nonvet_twenty <- read.csv("~/git/DSPG2020/career/data/sts_all_twenty.csv", row.names = 1)
sts_nonvet_thirty <- read.csv("~/git/DSPG2020/career/data/sts_all_thirty.csv", row.names = 1)

sts_nonvet_ten <- seqdef(sts_nonvet_ten, left="DEL", gaps="Civilian unemployment", right="DEL", cpal = c("#CBBEB5",  uva_color_palette[1], uva_color_palette[2],uva_color_palette[4],  uva_color_palette[6], uva_color_palette[9], "#CBBEB5"))
sts_nonvet_ten <- sts_nonvet_ten[, 1:10]

sts_nonvet_twenty <- seqdef(sts_nonvet_twenty, left="DEL", gaps="DEL", right="DEL", cpal = c("#CBBEB5",  uva_color_palette[1], uva_color_palette[2],uva_color_palette[4],  uva_color_palette[6], uva_color_palette[9], "#CBBEB5"))

sts_nonvet_thirty <- seqdef(sts_nonvet_thirty, left="DEL", gaps="Civilian unemployment", right="DEL", cpal = c("#CBBEB5",  uva_color_palette[1], uva_color_palette[2],uva_color_palette[4],  uva_color_palette[6], uva_color_palette[9], "#CBBEB5"))
sts_nonvet_thirty <- sts_nonvet_thirty[, 1:30]

## Aggregate cases for later clustering --------------
agg_1988 <- wcAggregateCases(sts_non_1988)
agg_1964 <- wcAggregateCases(sts_non_1964)

agg_ten <- wcAggregateCases(sts_nonvet_ten)
agg_twenty <- wcAggregateCases(sts_nonvet_twenty)
agg_thirty <- wcAggregateCases(sts_nonvet_thirty)

## Veteran data
# Import veteran job data ------
bg_vet_job <-read_csv("~/git/DSPG2020/career/data/04_bg_vet_job.csv")%>%
  mutate(year_enter_job_market = year(date_enter_job_market))%>%
  select(-noofjobs, -sector, -tenure) %>%
  group_by(id) %>%
  mutate(years_in_job_market = max(end_year) - (min(start_year)))

# What date did each veterans end their last ONET 55 job? (Date of military exit) ----
vet_endmilitary <- bg_vet_job%>%
  mutate(date_end_onet55 = if_else(is_onet55==T, enddate, as.Date(NA)))%>%
  filter(!is.na(date_end_onet55))%>%  #exluce people who don't have valid onet55 code
  select(id, date_end_onet55) %>%
  #keep the latest onet55 job
  group_by(id)%>%   
  arrange(desc(date_end_onet55))%>%
  group_by(id)%>%
  distinct(id, .keep_all = TRUE) 

# Join data together ------
bg_vet_job  <- inner_join(bg_vet_job, vet_endmilitary, by = "id")

# Prepare the data for sequence format ------
bg_vet_job_seq <- bg_vet_job %>%
  select(id, end_year, onet_job_zone, startdate, enddate, job_duration_day, date_end_onet55, year_enter_job_market, years_in_job_market) %>%
  mutate(year_end_onet55 = year(date_end_onet55)) %>%
  select(-year_enter_job_market) %>%
# Find jobs that came after the date ended onet55 job (Veteran career)
  filter(startdate >= date_end_onet55)%>% 
# Filter out 55 jobs that have the same start and end date  
  filter(onet_job_zone != 55)  %>%  
  mutate(start_year = year(startdate))

# Add variables, align sequences -----------
bg_vet_job_seq <- bg_vet_job_seq %>%
  mutate(years_after_onet55 = max(end_year) - year_end_onet55) %>%
  mutate(start_year = start_year - year_end_onet55 + 1)%>%  #transform from calender year to year start sequence analysis
  mutate(end_year = end_year - year_end_onet55 + 1) %>% #transform from calender year to year start sequence analysis
  select(id, start_year, end_year, onet_job_zone, year_end_onet55, years_after_onet55)%>%
  group_by(id)%>%
  arrange(desc(onet_job_zone))%>%
  group_by(id)%>%
  distinct(id, start_year, end_year, .keep_all = TRUE)

career_sequence_exit_year_vet <- function(name = "sts_vet_ten", enter_year, exit_year) {
sts_vet <- bg_vet_job_seq %>%
  filter(year_end_onet55 >= exit_year & year_end_onet55 <= enter_year) #%>%
  #select(-years_after_onet55)

sts_vet <- as.matrix(sts_vet)
sts_vet <- as.data.frame(sts_vet)
  
sts_vet <- seqformat(sts_vet, from = "SPELL", to = "STS",
            id = "id",  begin = "start_year", end = "end_year", 
            status = "onet_job_zone", process = FALSE)

# DEL left gaps for comparison
sts_vet <- seqdef(sts_vet, left="DEL", gaps="Civilian unemployment", right="DEL", cpal = c("#CBBEB5",  uva_color_palette[1], uva_color_palette[2],uva_color_palette[4],  uva_color_palette[6], uva_color_palette[9], "#CBBEB5"))

names(sts_vet) <- paste0(1:ncol(sts_vet))

assign(name, sts_vet, envir = .GlobalEnv)
}


career_sequence_exit_year_vet("sts_vet_1988", 1997, 1988)
career_sequence_exit_year_vet("sts_vet_1964", 1987, 1964)

# Object for cluster analysis
career_sequence <- function(name = "sts_vet_ten", years) {
sts_vet <- bg_vet_job_seq %>%
  filter(years_after_onet55 >= years) %>%
  select(-years_after_onet55)

sts_vet <- as.matrix(sts_vet)
sts_vet <- as.data.frame(sts_vet)
  
sts_vet <- seqformat(sts_vet, from = "SPELL", to = "STS",
            id = "id",  begin = "start_year", end = "end_year", 
            status = "onet_job_zone", process = FALSE)

# DEL left gaps for comparison
vet.seq <- seqdef(sts_vet, left="DEL", gaps="Civilian unemployment", right="DEL", cpal = c("#CBBEB5",  uva_color_palette[1], uva_color_palette[2],uva_color_palette[4],  uva_color_palette[6], uva_color_palette[9], "#CBBEB5"))

# Filter out veterans who do not have a complete career in time period of interest
sts_vet <- vet.seq[, 1:years]

assign(name, sts_vet, envir = .GlobalEnv)
}

career_sequence("sts_vet_ten", 10)
career_sequence("sts_vet_twenty", 20)
career_sequence("sts_vet_thirty", 30)

```


```{r exploratory}
cohorts_rect <- data.frame("max_exit" = c(2017, 2007, 1997, 1987),
                           "min_exit" = c(2007, 1997, 1987, 1965),
                           "fill" = c("a", "b", "c", "d")) 

bg_all_job_seq %>%
  group_by(id) %>% mutate(year_enter_job_market = min(start_year)) %>% select(id, year_enter_job_market) %>%
  distinct() %>%
  group_by(year_enter_job_market) %>% summarise(n = n()) %>%
  ggplot(aes(x = year_enter_job_market, y = n)) +
  geom_area() +
  geom_rect(data=cohorts_rect, inherit.aes=FALSE, 
            aes(xmin=max_exit, 
                xmax=min_exit, ymin=0,
                ymax=30000, fill = fill),
            color="transparent", alpha=0.3) +
  xlim(1960, 2017) +
  labs(main = "Distribution of years entering the civilian workforce for nonveterans",
       y = "n = 628,219")
  theme_bw()
  
bg_vet_job_seq %>%
  group_by(id) %>% mutate(year_enter_job_market = (year_end_onet55 + min(start_year) - 1)) %>% select(id, year_enter_job_market) %>%
  distinct() %>%
  group_by(year_enter_job_market) %>% summarise(n = n()) %>%
ggplot(aes(x = year_enter_job_market, y = n)) +
  geom_area() +
  geom_rect(data=cohorts_rect, inherit.aes=FALSE, 
            aes(xmin=max_exit, 
                xmax=min_exit, ymin=0,
                ymax=300, fill = fill),
            color="transparent", alpha=0.3) +
  xlim(1960, 2017) +
  labs(main = "Distribution of years entering the civilian workforce for veterans",
       y = "n = 5,185")
  theme_bw()
```
The 1988 cohorts include individuals in the blue rectangle. 
The 1964 cohorts include individuals in the purple rectangle.

## Comparison of 1988 cohort
```{r 1988-cohorts, fig.height=8}
## Exploratory plots ------
exploratory_ggplots <- function(name, sts_vet_ten) {
statd <- seqstatd(sts_vet_ten)
states <- statd$Frequencies
states <- as.data.frame(states)
names(states) <- paste0(1:ncol(sts_vet_ten))
states <- states %>% rownames_to_column(var = "state")
states <- melt(states, id = "state")

v_states <- statd$ValidStates
v_states <- as.data.frame(v_states)
v_states <- v_states %>% rownames_to_column(var = "variable")

meant <- seqmeant(sts_vet_ten)
meant <- as.data.frame(meant)
meant <- meant %>% rownames_to_column(var = "state")

library(scales)
d <- ggplot() + 
  geom_area(data = states, aes(x = as.numeric(variable), y = value, fill = state, position = "fill")) +
  geom_line(data = v_states, aes(x = as.numeric(variable), y = v_states/nrow(sts_vet_ten)),
            size = 2) +
  scale_y_continuous(name = paste0("Share of careers in a job state, (n = ", nrow(sts_vet_ten), ")"),
                     sec.axis = sec_axis(trans = ~.*(nrow(sts_vet_ten)), name = "Number of careers in year of career")) +
  xlim(1,ncol(sts_vet_ten)) +
  scale_fill_manual(values = c("#CBBEB5",  uva_color_palette[1], uva_color_palette[2], uva_color_palette[4],  uva_color_palette[6], uva_color_palette[9], "#CBBEB5"), labels = c("Missing zone", "Zone 1", "Zone 2", "Zone 3", "Zone 4", "Zone 5", "Civilian unemployed")) +
  xlab("Year of career") +
  labs(title = paste0(name, ": Veterans post-military career path"),
       fill = "Job state") +
  theme_classic() +
  theme(legend.position = "bottom")

mt <- ggplot(data = meant, aes(x = state, y = Mean, fill = state)) +
  geom_col() +
  scale_fill_manual(values = c("#CBBEB5",  uva_color_palette[1], uva_color_palette[2], uva_color_palette[4],  uva_color_palette[6], uva_color_palette[9], "#CBBEB5"), labels = c("Missing zone", "Zone 1", "Zone 2", "Zone 3", "Zone 4", "Zone 5", "Civilian unemployed")) +
  scale_x_discrete(name = "Job state", labels = c("Missing zone", "Zone 1", "Zone 2", "Zone 3", "Zone 4", "Zone 5", "Civilian unemployed")) +
  labs(title = paste0(name, ": Mean time spent in each state for veteran's whole career sequence"),
       fill = "Job state",
       caption = "Data: BGT Resumes") +
  ylim(0,12) +
  ylab("Mean time (years)") +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 40, hjust = 1))

assign(paste0(name, "_dplot"), d, envir = .GlobalEnv)
assign(paste0(name, "_mtplot"), mt, envir = .GlobalEnv)
}

exploratory_ggplots("non_1988", sts_non_1988)
exploratory_ggplots("vet_1988", sts_vet_1988)

nonvetids <- sts_non_1988 %>% rownames_to_column("id") %>% select(id)
vetids <- sts_vet_1964 %>% rownames_to_column("id") %>% select(id)
nonvetids <- as.data.frame(nonvetids)
vetids <- as.data.frame(vetids)
ids <- rbind(nonvetids, vetids)

# Filtering out NAs, which are about 50%
bg_all_demographic %>% filter(id %in% ids$id) %>%
  filter(!is.na(degree_highest)) %>%
  group_by(veteran) %>% mutate(total = n()) %>% group_by(veteran, degree_highest) %>%
  summarise(n = n(), percent = round((n/total) * 100, 2)) %>% distinct() %>%
  arrange(veteran, -percent)
# Veterans have higher masters and bachelors degrees, phd is about the same, high school and associates are lower


grid.arrange(non_1988_dplot, vet_1988_dplot, ncol = 2)
grid.arrange(non_1988_mtplot, vet_1988_mtplot, ncol = 2)
```

## 1964 cohort comparison
```{r 1964-cohorts, fig.height=8}
exploratory_ggplots("non_1964", sts_non_1964)
exploratory_ggplots("vet_1964", sts_vet_1964)

nonvetids <- sts_non_1964 %>% rownames_to_column("id") %>% select(id)
vetids <- sts_vet_1964 %>% rownames_to_column("id") %>% select(id)
nonvetids <- as.data.frame(nonvetids)
vetids <- as.data.frame(vetids)
ids <- rbind(nonvetids, vetids)

# Filtering out NAs, which are about 50%
bg_all_demographic %>% filter(id %in% ids$id) %>%
  filter(!is.na(degree_highest)) %>%
  group_by(veteran) %>% mutate(total = n()) %>% group_by(veteran, degree_highest) %>%
  summarise(n = n(), percent = round((n/total) * 100, 2)) %>% distinct() %>%
  arrange(veteran, -percent)

# Veterans have a higher % with masters degrees, bachelors, lower with phd, high school, associates

grid.arrange(non_1964_dplot, vet_1964_dplot, ncol = 2)
grid.arrange(non_1964_mtplot, vet_1964_mtplot, ncol = 2)
```

## Fixed time cohorts
```{r fixed-time}
## Function for creating a transition matrix
career_sequence_transition <- function(transition_name = "transition_ten", cost_name = "cost_ten", sequence_object) {
## Computing transition rates
#counts
transition_matrix <- seqtrate(sequence_object, weighted=FALSE, count=TRUE)
round(transition_matrix,2)
#standardize counts

#make the diagnal 0
diag(transition_matrix) = 0
round(transition_matrix,2)

#calculate rowsum
rowsum <- apply(transition_matrix, 1, sum)

#calculate percent 
for(i in 1:nrow(transition_matrix)) {
  for (j in 1:ncol(transition_matrix)){
    transition_matrix[i,j] = transition_matrix[i,j]/rowsum[i]
  }
}

#code NAs as 0
for(j in 1:ncol(transition_matrix)){
      transition_matrix[7,j] = transition_matrix[7,j] = 0
}
transition_matrix <- round(transition_matrix,2)


#convert transition matrix to cost matrix
cost_matrix <- 2-transition_matrix
round(cost_matrix,2)

#make the diagnal 0
diag(cost_matrix) = 0
cost_matrix <- round(cost_matrix,2)

assign(transition_name, transition_matrix, envir = .GlobalEnv)
assign(cost_name, cost_matrix, envir = .GlobalEnv)
}

gg_transition_matrix <- function(tr, title) {
round(tr, 2)
transition.matrix.df <- as.data.frame(tr)
transition.matrix.df$startState <- row.names(transition.matrix.df)
transition.matrix.df <- melt(transition.matrix.df)
colnames(transition.matrix.df)[colnames(transition.matrix.df) == "variable"] <- "endState"
                                        
ggplot(data = transition.matrix.df, aes(x = endState, y = factor(startState,        levels = rev(levels(factor(startState)))), fill = value)) + 
  geom_tile() +    
  geom_text(aes(label=round(value,3))) +  
  scale_x_discrete(position = "top") +  
  scale_fill_gradient(low = "white", high = "#E57200") +
  theme(legend.position="bottom", panel.background = element_blank(),         plot.title = element_text(face = "bold")) +  
  labs(fill = "Transition Probability",        
       title = title,     
       y = "Current Job",       
       x = "Next Job")
}

career_sequence_transition("transition_ten_non", "cost_ten_non", sts_nonvet_ten)
career_sequence_transition("transition_twenty_non", "cost_twenty_non", sts_nonvet_twenty)
career_sequence_transition("transition_thirty_non", "cost_thirty_non", sts_nonvet_thirty)

## Function for creating sequence graphs
career_sequence_graphs <- function(sequence_object) {
seqdplot(sequence_object)
seqmtplot(sequence_object)
}

career_sequence_graphs(sts_nonvet_ten)
career_sequence_graphs(sts_vet_ten)

career_sequence_graphs(sts_nonvet_twenty)
career_sequence_graphs(sts_vet_twenty)

career_sequence_graphs(sts_nonvet_thirty)
career_sequence_graphs(sts_vet_thirty)
```

## Comparing clustering

```{r clustering, fig.height=5, fig.width=15}
# Compute transition rates -------
career_sequence_transition("transition_ten_vet", "cost_ten_vet", sts_vet_ten)
career_sequence_transition("transition_twenty_vet", "cost_twenty_vet", sts_vet_twenty)
career_sequence_transition("transition_thirty_vet", "cost_thirty_vet", sts_vet_thirty)

career_sequence_transition("transition_ten_non", "cost_ten_non", sts_nonvet_ten)
career_sequence_transition("transition_twenty_non", "cost_twenty_non", sts_nonvet_twenty)
career_sequence_transition("transition_thirty_non", "cost_thirty_non", sts_nonvet_thirty)

# Compute clusters ------------------
## We pick PAM 8 as a baseline, this could be changed ------------------
# vet ten
diss <- seqdist(sts_vet_ten, method = "OM", indel = "auto", sm = cost_ten_vet, with.missing = TRUE, full.matrix=FALSE)
clusterpam <- wcKMedoids(diss, k = 8)

clusterpam_8_v10 <- clusterpam$clustering
clusterpam_8_v10fac <- factor(clusterpam_8_v10, labels = paste("Type", 1:8))

clusterpam_8_v10fac <- factor(clusterpam_8_v10fac, levels = c("Type 2", "Type 8", "Type 3", "Type 7", "Type 4", "Type 5", "Type 1", "Type 6"))

# vet twenty
#diss <- seqdist(sts_vet_twenty, method = "OM", indel = "auto", sm = cost_twenty_vet, with.missing = TRUE, full.matrix=FALSE)
#clusterpam <- wcKMedoids(diss, k = 8)

#clusterpam_8_v20 <- clusterpam$clustering
#clusterpam_8_v20fac <- factor(clusterpam_8_v20, labels = paste("Type", 1:8))

#clusterpam_8_v20fac <- factor(clusterpam_8_v20fac, levels = c("Type 2", "Type 8", "Type 3", "Type 7", "Type 4", "Type 5", "Type 1", "Type 6"))

# vet thirty
diss <- seqdist(sts_vet_thirty, method = "OM", indel = "auto", sm = cost_thirty_vet, with.missing = TRUE, full.matrix=FALSE)
clusterpam <- wcKMedoids(diss, k = 8)

clusterpam_8_v30 <- clusterpam$clustering
clusterpam_8_v30fac <- factor(clusterpam_8_v30, labels = paste("Type", 1:8))

clusterpam_8_v30fac <- factor(clusterpam_8_v30fac, levels = c("Type 2", "Type 8", "Type 3", "Type 7", "Type 4", "Type 5", "Type 1", "Type 6"))

# nonvet ten
unique_sts_non_ten <- sts_nonvet_ten[agg_ten$aggIndex,]

nonvet.seq <- seqdef(unique_sts_non_ten, weights = agg_ten$aggWeights)
diss <- seqdist(nonvet.seq, method = "OM", indel = "auto", sm = cost_ten_non, with.missing = TRUE, full.matrix=FALSE)
clusterpam <- wcKMedoids(diss, k = 8)

clusterpam_8_n10 <- clusterpam$clustering
clusterpam_8_n10fac <- factor(clusterpam_8_n10[agg_ten$disaggIndex], labels = paste("Type", 1:8))

# nonvet twenty
#unique_sts_non_twenty <- sts_nonvet_twenty[agg_twenty$aggIndex,]

#nonvet.seq <- seqdef(unique_sts_non_twenty, weights = agg_twenty$aggWeights)
#diss <- seqdist(nonvet.seq, method = "OM", indel = "auto", sm = cost_twenty_non, with.missing = TRUE, full.matrix=FALSE)
#clusterpam <- wcKMedoids(diss, k = 8)

#clusterpam_8_n20 <- clusterpam$clustering
#clusterpam_8_n20fac <- factor(clusterpam_8_n20[agg_twenty$disaggIndex], labels = paste("Type", 1:8))

# nonvet thirty
unique_sts_non_thirty <- sts_nonvet_thirty[agg_thirty$aggIndex,]

nonvet.seq <- seqdef(unique_sts_non_thirty, weights = agg_thirty$aggWeights)
diss <- seqdist(nonvet.seq, method = "OM", indel = "auto", sm = cost_thirty_non, with.missing = TRUE, full.matrix=FALSE)
clusterpam <- wcKMedoids(diss, k = 8)

clusterpam_8_n30 <- clusterpam$clustering
clusterpam_8_n30fac <- factor(clusterpam_8_n30[agg_thirty$disaggIndex], labels = paste("Type", 1:8))


seqdplot(sts_vet_ten[,1:10], group = clusterpam_8_v10fac, border = NA, use.layout = TRUE, cols = 8, withlegend = F, main = "v10")
seqlegend(sts_vet_ten)
seqdplot(sts_nonvet_ten[,1:10], group = clusterpam_8_n10fac, border = NA, use.layout = TRUE, cols = 8, withlegend = F, main = "n10")

seqdplot(sts_vet_thirty[,1:30], group = clusterpam_8_v30fac, border = NA, use.layout = TRUE, cols = 8, withlegend = F, main = "v30")
seqlegend(sts_vet_ten)
seqdplot(sts_nonvet_thirty[,1:30], group = clusterpam_8_n30fac, border = NA, use.layout = TRUE, cols = 8, withlegend = F, main = "n30")

```



## Appendix
```{r, include = FALSE, eval = FALSE}
# DO NOT RUN
# How many identical sequences are there?
library(WeightedCluster)

career_sequence_exit_year <- function(name = "sts_vet_ten", enter_year, exit_year) {
sts_vet <- sts_all %>%
  filter(start_year >= exit_year & start_year <= enter_year) #%>%
  #select(-years_after_onet55)

sts_vet <- as.matrix(sts_vet)
sts_vet <- as.data.frame(sts_vet)
  
sts_vet <- seqformat(sts_vet, from = "SPELL", to = "STS",
            id = "id",  begin = "start_year", end = "end_year", 
            status = "onet_job_zone", process = FALSE)

sts_vet <- seqdef(sts_vet, left="DEL", gaps="Civilian unemployment", right="DEL", cpal = c("#CBBEB5",  uva_color_palette[1], uva_color_palette[2],uva_color_palette[4],  uva_color_palette[6], uva_color_palette[9], "#CBBEB5"))

names(sts_vet) <- paste0(1:ncol(sts_vet))

assign(name, sts_vet, envir = .GlobalEnv)
}

career_sequence_exit_year("sts_non_2008", 2017, 2008)
career_sequence_exit_year("sts_non_1998", 2007, 1998)
career_sequence_exit_year("sts_non_1988", 1997, 1988)
career_sequence_exit_year("sts_non_1964", 1987, 1964)

agg_2008 <- wcAggregateCases(sts_non_2008)
print(agg_2008)
agg_1998 <- wcAggregateCases(sts_non_1998)
print(agg_1998)
agg_1988 <- wcAggregateCases(sts_non_1988)
print(agg_1988)
#Number of disaggregated cases:  181889 
#Number of aggregated cases:  17765 
#Average aggregated cases:  10.23862 
#Average (weighted) aggregation:  10.23862 
agg_1964 <- wcAggregateCases(sts_non_1964)
print(agg_1964)
#Number of disaggregated cases:  62534 
#Number of aggregated cases:  8866 
#Average aggregated cases:  7.053237 
#Average (weighted) aggregation:  7.053237 

agg_ten <- wcAggregateCases(sts_nonvet_ten)
print(agg_ten)
#Number of disaggregated cases:  387383 
#Number of aggregated cases:  41151 
#Average aggregated cases:  9.413696 
#Average (weighted) aggregation:  9.413696 
agg_twenty <- wcAggregateCases(sts_nonvet_twenty)
print(agg_twenty)
#Number of disaggregated cases:  159609 
#Number of aggregated cases:  63552 
#Average aggregated cases:  2.511471 
#Average (weighted) aggregation:  2.511471 
agg_thirty <- wcAggregateCases(sts_nonvet_thirty)
print(agg_thirty)
#Number of disaggregated cases:  46718 
#Number of aggregated cases:  29633 
#Average aggregated cases:  1.576553 
#Average (weighted) aggregation:  1.576553 
```



```{r fig.height=8, fig.width=10, eval = FALSE}
## Function for creating a transition matrix
career_sequence_transition <- function(transition_name = "transition_ten", cost_name = "cost_ten", sequence_object) {
## Computing transition rates
#counts
transition_matrix <- seqtrate(sequence_object, weighted=FALSE, count=TRUE)
round(transition_matrix,2)
#standardize counts

#make the diagnal 0
diag(transition_matrix) = 0
round(transition_matrix,2)

#calculate rowsum
rowsum <- apply(transition_matrix, 1, sum)

#calculate percent 
for(i in 1:nrow(transition_matrix)) {
  for (j in 1:ncol(transition_matrix)){
    transition_matrix[i,j] = transition_matrix[i,j]/rowsum[i]
  }
}

#code NAs as 0
for(j in 1:ncol(transition_matrix)){
      transition_matrix[7,j] = transition_matrix[7,j] = 0
}
transition_matrix <- round(transition_matrix,2)


#convert transition matrix to cost matrix
cost_matrix <- 2-transition_matrix
round(cost_matrix,2)

#make the diagnal 0
diag(cost_matrix) = 0
cost_matrix <- round(cost_matrix,2)

assign(transition_name, transition_matrix, envir = .GlobalEnv)
assign(cost_name, cost_matrix, envir = .GlobalEnv)
}

gg_transition_matrix <- function(tr, title) {
round(tr, 2)
transition.matrix.df <- as.data.frame(tr)
transition.matrix.df$startState <- row.names(transition.matrix.df)
transition.matrix.df <- melt(transition.matrix.df)
colnames(transition.matrix.df)[colnames(transition.matrix.df) == "variable"] <- "endState"
                                        
ggplot(data = transition.matrix.df, aes(x = endState, y = factor(startState,        levels = rev(levels(factor(startState)))), fill = value)) + 
  geom_tile() +    
  geom_text(aes(label=round(value,3))) +  
  scale_x_discrete(position = "top") +  
  scale_fill_gradient(low = "white", high = "#E57200") +
  theme(legend.position="bottom", panel.background = element_blank(),         plot.title = element_text(face = "bold")) +  
  labs(fill = "Transition Probability",        
       title = title,     
       y = "Current Job",       
       x = "Next Job")
}

career_sequence_transition("transition_ten_non", "cost_ten_non", sts_nonvet_ten)
career_sequence_transition("transition_twenty_non", "cost_twenty_non", sts_nonvet_twenty)
career_sequence_transition("transition_thirty_non", "cost_thirty_non", sts_nonvet_thirty)

## Function for creating sequence graphs
career_sequence_graphs <- function(sequence_object) {
seqdplot(sequence_object)
seqmtplot(sequence_object)
}

career_sequence_graphs(sts_nonvet_ten)
tr <- seqtrate(sts_nonvet_ten)
gg_transition_matrix(tr, "Ten Years Nonveteran Year to Year Transition Matrix")
gg_transition_matrix(transition_ten_non, "Ten Years Nonveteran Job to Job Transition Matrix")

career_sequence_graphs(sts_nonvet_twenty)
tr <- seqtrate(sts_nonvet_twenty)
gg_transition_matrix(tr, "Twenty Years Nonveteran Year to Year Transition Matrix")
gg_transition_matrix(transition_twenty_non, "Twenty Years Nonveteran Job to Job Transition Matrix")

career_sequence_graphs(sts_nonvet_thirty)
tr <- seqtrate(sts_nonvet_thirty)
gg_transition_matrix(tr, "Thirty Years Nonveteran Year to Year Transition Matrix")
gg_transition_matrix(transition_thirty_non, "Thirty Years Nonveteran Job to Job Transition Matrix")
```

# Clustering

```{r, eval = FALSE}
sts_all_thirty <- read.csv("~/git/DSPG2020/career/data/sts_all_thirty.csv", row.names = 1)

all.seq_thirty <- seqdef(sts_all_thirty, left="DEL", gaps="Civilian unemployment", right="DEL", cpal = c("#CBBEB5",  uva_color_palette[1], uva_color_palette[2],uva_color_palette[4],  uva_color_palette[6], uva_color_palette[9], "#CBBEB5"))

# Filter out veterans who do not have a complete career in time period of interest
sts_all_thirty <- all.seq_thirty[, 1:30]

library(WeightedCluster)

agg_sts_all_thirty <- wcAggregateCases(sts_nonvet_thirty)
print(agg_sts_all_thirty)

unique_sts_all_thirty <- sts_nonvet_thirty[agg_sts_all_thirty$aggIndex,]

nonvet.seq <- seqdef(unique_sts_all_thirty, weights = agg_sts_all_thirty$aggWeights)
diss <- seqdist(nonvet.seq, method = "OM", indel = "auto", sm = cost_thirty_non, with.missing = TRUE, full.matrix=FALSE)
clusterward <- agnes(diss, diss = TRUE, method = "ward")
clusterward <- hclust(as.dist(diss), method = "ward.D2", members = agg_sts_all_thirty$aggWeights)

saveRDS(clusterward, file = "~/git/DSPG2020/career/data/clusterward_30yrs_noretirement.rds")

```

## 1988 cohort Comparison


```{r fig.height = 10, eval = FALSE}
career_sequence_exit_year <- function(name = "sts_vet_ten", enter_year, exit_year) {
sts_vet <- sts_all %>%
  filter(start_year >= exit_year & start_year <= enter_year) #%>%
  #select(-years_after_onet55)

sts_vet <- as.matrix(sts_vet)
sts_vet <- as.data.frame(sts_vet)
  
sts_vet <- seqformat(sts_vet, from = "SPELL", to = "STS",
            id = "id",  begin = "start_year", end = "end_year", 
            status = "onet_job_zone", process = FALSE)

sts_vet <- seqdef(sts_vet, left="DEL", gaps="Civilian unemployment", right="DEL", cpal = c("#CBBEB5",  uva_color_palette[1], uva_color_palette[2],uva_color_palette[4],  uva_color_palette[6], uva_color_palette[9], "#CBBEB5"))

names(sts_vet) <- paste0(1:ncol(sts_vet))

assign(name, sts_vet, envir = .GlobalEnv)
}

#career_sequence_exit_year("sts_non_1964", 1987, 1964)

statd <- seqstatd(sts_non_1988)
states <- statd$Frequencies
states <- as.data.frame(states)
names(states) <- paste0(1:31)
states <- states %>% rownames_to_column(var = "state")
states <- melt(states, id = "state")

v_states <- statd$ValidStates
v_states <- as.data.frame(v_states)
v_states <- v_states %>% rownames_to_column(var = "variable")

meant <- seqmeant(sts_non_1988)
meant <- as.data.frame(meant)
meant <- meant %>% rownames_to_column(var = "state")

# Plot state frequency over time ---------------

library(scales)
ggplot() + 
  geom_area(data = states, aes(x = as.numeric(variable), y = value, fill = state, position = "fill")) +
  geom_line(data = v_states, aes(x = as.numeric(variable), y = v_states/nrow(sts_non_1988)),
            size = 2) +
  scale_y_continuous(name = paste0("Share of careers in a job state, (n = ", nrow(sts_non_1988), ")"),
                     sec.axis = sec_axis(trans = ~.*(nrow(sts_non_1964)), name = "Number of careers in year of career")) +
  xlim(0,31) +
  scale_fill_manual(values = c("#CBBEB5",  uva_color_palette[1], uva_color_palette[2], uva_color_palette[4],  uva_color_palette[6], uva_color_palette[9], "#CBBEB5"), labels = c("Missing zone", "Zone 1", "Zone 2", "Zone 3", "Zone 4", "Zone 5", "Civilian unemployed")) +
  xlab("Year of career") +
  labs(title = "Veteran career paths after military exit",
       fill = "Job state") +
  theme_classic() +
  theme(legend.position = "bottom")

# Plot mean time in each state --------------

ggplot(data = meant, aes(x = state, y = Mean, fill = state)) +
  geom_col() +
  scale_fill_manual(values = c("#CBBEB5",  uva_color_palette[1], uva_color_palette[2], uva_color_palette[4],  uva_color_palette[6], uva_color_palette[9], "#CBBEB5", "#CBBEB5"), labels = c("Missing zone", "Zone 1", "Zone 2", "Zone 3", "Zone 4", "Zone 5", "Civilian unemployed")) +
  scale_x_discrete(name = "Job state", labels = c("Missing zone", "Zone 1", "Zone 2", "Zone 3", "Zone 4", "Zone 5", "Civilian unemployed")) +
  labs(title = "Mean time spent in each state for veteran's post-military career",
       fill = "Job state",
       caption = "Data: BGT Resumes") +
  ylab("Mean time (years)") +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 40, hjust = 1))

library(WeightedCluster)

#agg <- wcAggregateCases(sts_non_1964)
#print(agg)
unique <- sts_non_1964[agg_1964$aggIndex, ]
#seq <- seqdef(unique, left="DEL", gaps="DEL", right="DEL", weights = agg$aggWeights)
diss <- seqdist(unique, method = "OM", indel = "auto", sm = cost_matrix, with.missing = TRUE, norm = TRUE)

#seqstatl(seq)

#non.seq.OM <- seqdist(sts_non_1964, method = "OM", indel = "auto", sm = cost_matrix, with.missing = TRUE, norm = TRUE)
clusterpam <- wcKMedoids(diss, k = 7)
clusterpam7 <- clusterpam$clustering

clusterpam7_fac <- factor(clusterpam7[agg$disaggIndex], labels = paste("Type", 1:7))

#uniqueCluster7 <- clusterpam[clustering]
sts_non_1964$cluster7 <- clusterpam[agg$disaggIndex]

#longitudinal plot
seqdplot(sts_non_1964, group = clusterpam7_fac)
#another cluster plot
seqmtplot(sts_non_1964, group = clusterpam7_fac)
#index plot
seqIplot(sts_non_1964, group = clusterpam7_fac)
```
```{r eval = FALSE}
## Computing transition rates
#counts
transition_matrix <- seqtrate(sts_non_1964, weighted=FALSE, count=TRUE)
round(transition_matrix,2)
#standardize counts

#make the diagnal 0
diag(transition_matrix) = 0
round(transition_matrix,2)

#calculate rowsum
rowsum <- apply(transition_matrix, 1, sum)

#calculate percent 
for(i in 1:nrow(transition_matrix)) {
  for (j in 1:ncol(transition_matrix)){
    transition_matrix[i,j] = transition_matrix[i,j]/rowsum[i]
  }
}

#code NAs as 0
for(j in 1:ncol(transition_matrix)){
      transition_matrix[7,j] = transition_matrix[7,j] = 0
}
transition_matrix <- round(transition_matrix,2)


#convert transition matrix to cost matrix
cost_matrix <- 2-transition_matrix
round(cost_matrix,2)

#make the diagnal 0
diag(cost_matrix) = 0
cost_matrix <- round(cost_matrix,2)
```

