---
title: "Using CIP/Major to Impute Minedu (2019)"
author: "Sarah McDonald"
date: "7/13/2020"
output: md_document
always_allow_html: true
---

```{r,echo = F, message = F, warning=F}
library(dplyr)
conn <- RPostgreSQL::dbConnect(drv = RPostgreSQL::PostgreSQL(),
                               dbname = "sdad",
                               host = "postgis1",
                               port = 5432,
                               user = Sys.getenv(x = "DB_USR"),
                               password = Sys.getenv(x = "DB_PWD"))



bgt <- RPostgreSQL::dbGetQuery(
  conn = conn, 
  statement = "SELECT DISTINCT A.id, A.soc, A.minedu, B.onet, 
  CASE WHEN C.cip IS NULL THEN FALSE ELSE TRUE END AS hascip,
  CASE WHEN D.stdmajor IS NULL THEN FALSE ELSE TRUE END AS hasstdmajor
      
  FROM bgt_job.jolts_comparison_2019 A
      LEFT JOIN bgt_job.main B ON A.id = B.id 
      LEFT JOIN bgt_job.cip C ON A.id = C.id 
      LEFT JOIN bgt_job.maj D ON A.id = D.id 
  ")


```
Share of Job Ads with missing MINEDU: 
```{r, echo =  F}
minedu_na <- bgt %>% filter(is.na(minedu) == TRUE)

print(nrow(minedu_na)/nrow(bgt))
```

Share of Job Ads with missing MINEDU, where STDMAJOR OR CIP is present (I think these variables are one-to-one).
```{r, echo =  F}
print(sum(minedu_na$hasstdmajor)/nrow(minedu_na))
print(sum(minedu_na$hascip)/nrow(minedu_na))
```

Having a CIP or STDMAJOR might not necessarily indicate that at least a bachelor degree is required.
Share of BGT job ads with less than a bachelors degree (minedu < 16), that have a CIP code:

```{r,echo = F}
nrow(bgt %>% filter(minedu < 16 & hascip== TRUE))/nrow(bgt %>% filter(minedu <16))
```




A few surprising Onet Titles of Job Ads with missing MINEDU (having a bachelor degree may be implied by job title, but not explicit in job posting):
```{r, echo = F}
titlesOfNonMatches <- RPostgreSQL::dbGetQuery(
  conn = conn, 
  statement = "SELECT DISTINCT B.onetname
  FROM bgt_job.jolts_comparison_2019 A
  LEFT JOIN bgt_job.main B ON A.id = B.id
  WHERE A.minedu IS NULL
  ")
titlesOfNonMatches[6,]
titlesOfNonMatches[431,]
titlesOfNonMatches[301,]
titlesOfNonMatches[338,]
titlesOfNonMatches[91,]
```

