---
title: "three-measure-comparison"
author: "Sarah McDonald"
date: "7/14/2020"
#output: html_document
output: md_document
always_allow_html: true
  
---

```{r, warning=F,message = F, echo = F}
library(readxl)
library(dplyr)
conn <- RPostgreSQL::dbConnect(drv = RPostgreSQL::PostgreSQL(),
                               dbname = "sdad",
                               host = "postgis1",
                               port = 5432,
                               user = Sys.getenv(x = "DB_USR"),
                               password = Sys.getenv(x = "DB_PWD"))
```

Gather BGT data and ONET/SOC/OCC/NCSES crosswalk:
```{r}
bgt <- RPostgreSQL::dbGetQuery(
  conn = conn, 
  statement = paste("SELECT A.minedu, B.onet
  FROM bgt_job.jolts_comparison_2010 A
  JOIN bgt_job.main B
  ON A.id = B.id
  WHERE A.state IN" , paste("(", paste(shQuote(c(state.name, "District of Columbia"), type="sh"), collapse=", "), ")", sep = ""), sep = ""))

xwalk <- read_xlsx("/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ONETSOC_SOC_OCC_NCSES_JUL2020.xlsx", na ="NA")
```

Create an STW column based off of the PER.HS column. If PER.HS is greater than or equal to 50, categorize the ONET code as STW. 
```{r}
xwalk$PER.HS.STW <- ifelse(xwalk$PER.HS >= 50 & is.na(xwalk$PER.HS) == FALSE, 1, 
                           ifelse(xwalk$PER.HS < 50 & is.na(xwalk$PER.HS) == FALSE, 0, NA))
```

Create a table with the following:

- ONET Code

- Mean BGT minedu

- Median BGT minedu

- Min BGT minedu

- Max BGT minedu

- Number of BGT Observations

- Number of Missing BGT Observations

- Proportion of Missing BGT Observations to Total BGT Observations

- SDAD STW 

- NCSES STW

- Job Zone

- ONET Survey Percent of Respondents with No More than a High School Degree

- STW Based on ONET Survey

```{r, message = F, warning= F}
onet_distribution <- bgt %>%
  group_by(onet) %>%
  summarize(mean = round(mean(minedu, na.rm = T), 3), median = median(minedu, na.rm = T), min = ifelse(is.nan(mean) == TRUE, NA, min(minedu, na.rm = T)), 
            max = ifelse(is.nan(mean) == TRUE, NA, max(minedu, na.rm = T)), n_obs = n(), sum_nas = sum(is.na(minedu) ==TRUE), prop_nas = round(sum_nas/n_obs, 3)) %>%
  merge(xwalk[,c("SDADSTW", "NCSESSTW", "JobZone", "PER.HS", "ONETSOC", "PER.HS.STW")], by.x = "onet", by.y = "ONETSOC", all.x = T)

head(onet_distribution)
```

How many BGT ONETs are not in the crosswalk?
```{r}
list_onets_not_in_xwalk <- unique(onet_distribution$onet)[!(unique(onet_distribution$onet) %in% xwalk$ONETSOC)]
length(list_onets_not_in_xwalk)
```

