---
title: "Model Evaluation"
output: html_document
---

```{r}
load(here::here("data", "working", "model_objects", "basic_model_bayes_no_interact.RData"))
load(here::here("data", "working", "model_objects", "basic_model_bayes_yes_interact.RData"))
load(here::here("data", "working", "model_objects", "basic_model_freq_no_interact.RData"))
load(here::here("data", "working", "model_objects", "basic_model_freq_yes_interact.RData"))
load(here::here("data", "working", "model_objects", "neighbor_model_bayes_no_interact.RData"))
load(here::here("data", "working", "model_objects", "neighbor_model_bayes_yes_interact.RData"))

library(rstanarm)
library(bayesplot)
library(dplyr)
library(ggplot2)
library(sf)
library(broom)
library(purrr)
library(glue)
library(stringr)
library(viridis)

prepared_data <- readr::read_csv(here::here("data", "final", "response_time_model_data_prepared.csv"))
prepared_data_sp <- sf::st_read(here::here("data", "final", "response_time_model_data_prepared_sp.geojson"))


freq_models <- list("basic_model_freq_no_interact" = basic_model_freq_no_interact,
                    "basic_model_freq_yes_interact" = basic_model_freq_yes_interact)

bayes_models <- stanreg_list("basic_model_bayes_no_interact" = basic_model_bayes_no_interact,
                             "basic_model_bayes_yes_interact" = basic_model_bayes_yes_interact,
                             "neighbor_model_bayes_no_interact" = neighbor_model_bayes_no_interact,
                             "neighbor_model_bayes_yes_interact" = neighbor_model_bayes_yes_interact)
```


## Examine Coefficients

Setup everything
```{r}

freq_model_coefs <- map(freq_models, tidy,
                        conf.int = TRUE,
                        conf.level = 0.95,
                        exponentiate = FALSE)

bayes_model_coefs <- map(bayes_models, ~tidy(.x$stanfit,
                                             estimate.method = "median",
                                             conf.int = TRUE,
                                             conf.level = 0.95)) %>%
  map(~filter(.x, !(term %in% c("sigma", "mean_PPD", "log-posterior"))))


freq_model_coefs_trans <- freq_model_coefs %>%
  map(~mutate(.x, across(c(estimate,
                           conf.low,
                           conf.high),
                         list(scale_factor = ~exp(.x * (.x != .x[1])), # scales to exponential, expression insures intercept is right
                              time_to_incident = ~exp(.x * (.x != .x[1]) + .x[1]))))) # scales to exponential, expression insures intercept is right

bayes_model_coefs_trans <- bayes_model_coefs %>%
  map(~mutate(.x, across(c(estimate,
                           conf.low,
                           conf.high),
                         list(scale_factor = ~exp(.x * (.x != .x[1])),
                              time_to_incident = ~exp(.x * (.x != .x[1]) + .x[1])))))


plot_scale_factor <- function(data) {
  ggplot(data, aes(y = term)) +
    geom_point(aes(x = estimate_scale_factor)) +
    geom_errorbar(aes(xmin = conf.low_scale_factor, xmax = conf.high_scale_factor)) +
    theme_minimal() +
    labs(x = NULL, y = glue("Scale Factor from {round(data$estimate_time_to_incident[1],
                                                      digits = 1)} Minutes")) +
    coord_cartesian(xlim = c(0.25, 1.75)) +
    geom_vline(xintercept = 1, alpha = 0.5)
}

plot_time_to_incident <- function(data) {
  ggplot(data) +
    geom_point(aes(x = term, y = estimate_time_to_incident)) +
    geom_errorbar(aes(x = term, ymin = conf.low_time_to_incident, ymax = conf.high_time_to_incident)) +
    theme_minimal() +
    coord_flip() +
    labs(x = NULL, y = glue("Travel Time in Minutes"))
}

```

Actually make the plots:
```{r}
map(bayes_model_coefs_trans,
    plot_scale_factor)
```

```{r}
map(bayes_model_coefs_trans,
    plot_time_to_incident)

```

Looks like the after covid interaction term doesn't do a lot in the non neighborhood models. In the neighborhood models however, it looks like times consistently increase. 

Now for frequentist models:

```{r}
map(freq_model_coefs_trans,
    plot_scale_factor)
```

```{r}
map(freq_model_coefs_trans,
    plot_time_to_incident)
```

Basically identical to the bayesian models

### After Covid Interaction Term

```{r}

cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")



bayes_model_coefs_trans[[2]] %>%
  mutate(after_covid = str_detect(term, "(after_covidTRUE.*)")) %>%
  mutate(term = str_replace(term, "after_covidTRUE:", "")) %>%
  ggplot(aes(y = term, color = after_covid)) +
  geom_point(aes(x = estimate_scale_factor)) +
  geom_errorbar(aes(xmin = conf.low_scale_factor, xmax = conf.high_scale_factor)) +
  theme_minimal() +
  scale_color_manual(values = cbbPalette) +
  coord_cartesian(xlim = c(0.5, 1.5)) +
  geom_vline(xintercept = 1, alpha = 0.5) +
  labs(y = NULL, x = glue("Scale Factor from {round(bayes_model_coefs_trans[[2]]$estimate_time_to_incident[1],
                                                      digits = 1)} Minutes"),
       color = "During Covid")

bayes_model_coefs_trans[[4]] %>%
  mutate(after_covid = str_detect(term, "(after_covidTRUE.*)")) %>%
  mutate(term = str_replace(term, "after_covidTRUE:", "")) %>%
  ggplot(aes(y = term, color = after_covid)) +
  geom_point(aes(x = estimate_scale_factor)) +
  geom_errorbar(aes(xmin = conf.low_scale_factor, xmax = conf.high_scale_factor)) +
  theme_minimal() +
  scale_color_manual(values = cbbPalette) +
  coord_cartesian(xlim = c(0.5, 1.5)) +
  geom_vline(xintercept = 1, alpha = 0.5) +
  labs(y = NULL, x = glue("Scale Factor from {round(bayes_model_coefs_trans[[4]]$estimate_time_to_incident[1],
                                                      digits = 1)} Minutes"),
       color = "During Covid")

```


### Neighborhood Coefficients

```{r}
neighborhood_coefs <- map(bayes_model_coefs_trans, ~.x %>% 
    filter(str_detect(term, r"(b\[)")) %>% 
    mutate(NAME = str_extract(term, r"((?<=NAME:).*(?=\]))")) %>% 
    mutate(NAME = str_replace_all(NAME, "_", " ")))[3:4]

neighborhood_coefs_sp <- map(neighborhood_coefs, ~left_join(prepared_data_sp %>% 
  group_by(NAME) %>% 
  slice_head(1) %>% 
  ungroup(), .x, by = "NAME"))

neighborhood_coefs_sp$neighbor_model_bayes_no_interact %>% 
  ggplot() +
  geom_sf(aes(fill = estimate_scale_factor))
```



## Technical Examination

I will only be examining bayesian fits because they are almost identical to the frequentist fits. 

First we'll look at the loo statistics

```{r}
(loo_list <- map(bayes_models, loo))
loo_compare(loo_list)
```

Looks like the neighborhood level models do a lot! better, and they all have reasonable specifications

```{r}
map(bayes_models, pp_check)
```

The capture the majority of model variation, the remaining portion is likely due to it not being truly normal. Maybe a proper spatial mixed effects model could solve that?

Lets now check the residuals on a map:

```{r}
bayes_residuals <- map(bayes_models, residuals)

augmented_data <- prepared_data_sp %>% 
  mutate(resid_basic_no = bayes_residuals$basic_model_bayes_no_interact,
         resid_basic_yes = bayes_residuals$basic_model_bayes_yes_interact,
         resid_neighbor_no = bayes_residuals$neighbor_model_bayes_no_interact,
         resid_neighbor_yes = bayes_residuals$neighbor_model_bayes_yes_interact)
```

Basic Linear Model with No Interaction Term

```{r}
augmented_data %>% 
  ggplot() +
  stat_summary_hex(aes(x = scene_gps_longitude, 
                       y = scene_gps_latitude,
                       z = resid_basic_no)) +
  geom_sf(fill = NA,
          color = "#444444",
          alpha = 0.3,
          size = 0.1) +
  scale_fill_viridis() +
  theme_minimal() +
  labs(x = "Longitude", 
       y = "Latitude",
       fill = "Mean Residuals",
       title = "Basic Linear Model, No Interactions") 
```


Basic Linear Model with Interaction Term

```{r}
augmented_data %>% 
  ggplot() +
  stat_summary_hex(aes(x = scene_gps_longitude, 
                       y = scene_gps_latitude,
                       z = resid_basic_yes)) +
  geom_sf(fill = NA,
          color = "#444444",
          alpha = 0.3,
          size = 0.1) +
  scale_fill_viridis() +
  theme_minimal() +
  labs(x = "Longitude", 
       y = "Latitude",
       fill = "Mean Residuals",
       title = "Basic Linear Model, Interactions") 
```

Neighborhood Model with No Interaction Term

```{r}
augmented_data %>% 
  ggplot() +
  stat_summary_hex(aes(x = scene_gps_longitude, 
                       y = scene_gps_latitude,
                       z = resid_neighbor_no)) +
  geom_sf(fill = NA,
          color = "#444444",
          alpha = 0.3,
          size = 0.1) +
  scale_fill_viridis() +
  theme_minimal() +
  labs(x = "Longitude", 
       y = "Latitude",
       fill = "Mean Residuals",
       title = "Neighborhood Model, No Interactions") 
```

Neighborhood Model with Interaction Term

```{r}
augmented_data %>% 
  ggplot() +
  stat_summary_hex(aes(x = scene_gps_longitude, 
                       y = scene_gps_latitude,
                       z = resid_neighbor_yes)) +
  geom_sf(fill = NA,
          color = "#444444",
          alpha = 0.3,
          size = 0.1) +
  scale_fill_viridis() +
  theme_minimal() +
  labs(x = "Longitude", 
       y = "Latitude",
       fill = "Mean Residuals",
       title = "Neighborhood Model, Interactions") 
```

