---
title: "ONET Matching Attempt"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# load packages 
for (pkg in c("tidyverse", "igraph", "visNetwork", "data.table", "R.utils", "RPostgreSQL", "cowplot", "maditr", "stringr", "stringi", "mosaic", "dplyr", "ggplot2", "lubridate", "readxl")) {
  library(pkg, character.only = TRUE)
}
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
get_db_conn <-
  function(db_name = "sdad",
           db_host = "postgis1",
           db_port = "5432",
           db_user = Sys.getenv("db_usr"),
           db_pass = Sys.getenv("db_pwd")) {
    RPostgreSQL::dbConnect(
      drv = RPostgreSQL::PostgreSQL(),
      dbname = db_name,
      host = db_host,
      port = db_port,
      user = db_user,
      password = db_pass
    )
  }

con <- get_db_conn()

#query onet data
onet <- DBI::dbGetQuery(con, "SELECT * FROM onet.job_zones")

# query job data
job <- DBI::dbGetQuery(con, "SELECT *
FROM bgt_res.job A
JOIN bgt_res.id_msa_dc B
ON A.id = B.id"
)

DBI::dbDisconnect(con)
```

```{r}
#check that id columns are identical and drop one
ncol(job)
identical(job[[1]],job[[10]])
job <- job[-10]
```


There are `r length(unique(job$onet))` unique ONET codes in the job table and `r length(unique(onet$onetsoc_code))` unique ONET codes in the ONET table. 

```{r}
inner_onet <- merge(job, onet, by.x = "onet", by.y = "onetsoc_code")
```

An inner join (below) drops the number of observations in the table from `r nrow(job)` to `r nrow(inner_onet)`.

Looking at the rows that did not join successfully with an anti-join:
```{r}
anti_onet <- anti_join(job, onet, by = c("onet" = "onetsoc_code"))
```


```{r}
unique_anti <- unique(anti_onet$onet)
```
There are `r nrow(unique_anti)` unique ONET codes that did not join.

The following are the codes that most commonly did not join:
```{r}
grouped_anti_onet <- anti_onet %>%
  group_by(onet) %>%
  summarize(n=n()) %>%
  arrange(desc(n))
grouped_anti_onet
```

Using the ONET codes found online (which are slightly different from the codes in the pgAdmin database), we can attempt to further investigate the ONET codes that didn't join.
```{r}
onet_web <- read_excel('~/Occupation_Data.xlsx')
colnames(onet_web)
```

```{r}
onet_web <- rename(onet_web, c("soc_code" = "O*NET-SOC Code"))
```

```{r}
colnames(onet_web)
```

```{r}
onet_web_soc <- onet_web$soc_code
```

```{r}
grouped_anti_merged <- merge(grouped_anti_onet, onet_web, by.x = 'onet', by.y = 'soc_code')
```

Below are the most common values that *did* exist online, but did *not* merge with the job table. There are still codes that were in the pgAdmin ONET data that are not in the online data.
```{r}
grouped_anti_merged %>%   
  group_by(onet) %>%
  arrange(desc(n))
```

