---
title: "Health_Science_HT"
author: "Haleigh Tomlin"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(network)
library(igraph)
```


colorBlindness: https://cran.r-project.org/web/packages/colorBlindness/vignettes/colorBlindness.htm


```{r}
hs_data <- read_xlsx("Health_Science.xlsx")
```

```{r}
table(hs_data$...6)
```

^^ I put these SOCs in emily's code.
```{r}
hs_stem_creds <- read.csv("hs_stem_credentials.csv")
```

```{r}
#hs_stem_creds <- hs_stem_creds %>% select(SOC, Cert = Certification.Name)
```

joining back to the HS data.

```{r}
hs_data
```

```{r}
hs_data <- hs_data %>% mutate(ONETSOC = substr(Code, start = 1, stop = 7))
hs_data <- hs_data %>% mutate(STEM = `...6`) 
hs_data <- hs_data %>% select(ONETSOC, Code, STEM, Occupation)
```

```{r}
hs_stem_creds <- hs_stem_creds %>% select(Certification_Name = Certification.Name, SOC) %>% mutate(ONETSOC = substr(SOC, start = 1, stop = 7))
```

```{r}
hs_stem_creds
```
```{r}
hs <- hs_data %>% left_join(hs_stem_creds, by = c("Code" = "SOC", "ONETSOC"))
```


```{r}
hs2
```

```{r}
stw_certs <- d %>% select(ONETSOC = SOC, Cert = Certification.Name)
```

```{r}
hs2 <- hs %>% left_join(stw_certs, by = "ONETSOC")
```

```{r}
hs2$Certification_Name <- as.character(hs2$Certification_Name)
hs2$Certification <- ifelse(is.na(hs2$Certification_Name), hs2$Cert, hs2$Certification_Name)

```

```{r}
hs2 <- hs2 %>% mutate(Occupation = gsub(" ", "\n", Occupation))
```


```{r}
hs2 <- hs2 %>% select(ONETSOC, Certification, Occupation, STEM)
```

```{r}
hs_attributes1 <- hs2 %>% select(name = ONETSOC, Occupation, STEM)
hs_attributes2 <- hs2 %>% select(name = Certification, Occupation, STEM)

hs_attributes1 <- hs_attributes1[!duplicated(hs_attributes1$name),]
hs_attributes2 <- hs_attributes2[!duplicated(hs_attributes2$name),]

hs_attributes <- rbind(hs_attributes1, hs_attributes2)
```


```{r}
hs2 <- na.omit(hs2)
hs2 <- hs2[!duplicated(hs2),]
hsci <- graph.data.frame(hs2)


nodelist <- data.frame(name = igraph::V(hsci)$name)

nodelist <- nodelist %>% left_join(hs_attributes, by = "name")

health_sci <- graph.data.frame(hs2, directed = F, vertices = nodelist)
V(health_sci)$type <- bipartite.mapping(health_sci)$type
```


ifelse(V(health_sci)$type, "circle", "square")

```{r, fig.height = 20, fig.width = 20}
plot(health_sci,
     vertex.shape = ifelse(V(health_sci)$type, "circle", "square"),
     vertex.color = ifelse(V(health_sci)$type == F, ifelse(V(health_sci)$STEM == "STW", cbPalette[5], cbPalette[1]), cbPalette[4]),
     vertex.size = ifelse(V(health_sci)$type, 2, 10),
     vertex.label = ifelse(V(health_sci)$type == F, V(health_sci)$Occupation, NA),
     vertex.label.color = "black",
     vertex.label.cex = 0.5,
     edge.width = 1.5,
     layout = layout_nicely, xaxs = i, yaxs = i)
legend("topright", legend = c("STW: yellow squares", "STEM: gray squares", "Certifications: green circles"), bty = "n", title = "Legend", xjust = 0, cex = 1.5)
```

```{r}
get_network_stats(health_sci)
```


```{r}
hs[hs$Occupation == "Nursing Assistants",]
```



Health Sciences graph instead of nursing

For Health sciences, cybersecurity, manu we will have stem/stw


```{r}
all_df <- read_xlsx("All_Career_Clusters.xlsx")
```

```{r}
all_df
```

Vicki: "Occupation clusters include occupations from different major occupation groups."

So, I should join this to the bigger DF and then I can make a drop down?

```{r}
unique(all_df$`Code`[all_df$`Career Cluster` == "Manufacturing"])
```


```{r}
table(d$SOC %in% all_df$Code)
table(d$SOCNAME %in% all_df$Occupation)
```

```{r}
test <- substr(all_df$Code, start = 1, stop = 7)
```

```{r}
table(d$SOC %in% test)
```

