---
title: "HT_network_analysis"
author: "Haleigh Tomlin"
date: "7/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(igraph)
library(network)
library(readxl)
library(tidyr)
library(tidyverse)
library(ggplot2)
```


```{r}
#Color Blind Palettes

#color blind palette with grey

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

#color blind palette with black

cbPalette.blk<-c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

#display the palette

wheel <- function(col, radius = 1, ...)

  pie(rep(1, length(col)), col = col, radius = radius, ...)

wheel(cbPalette)
wheel(cbPalette.blk)
```

```{r, fig.height = 10}
#making a legend with Vicki's code
#install.packages("wesanderson")

library(wesanderson)
png('test.png',width=1000,height=800,units="px",bg = "transparent")

Zissou1<-c("#3B9AB2", "#78B7C5", "#EBCC2A", "#E1AF00", "#F21A00")

WA.pal<-wes_palette("Zissou1", 5, type = c("discrete"))

legend.network<- function(col, radius = 1, ...)

{

pie(rep(1, length(col)), col=col, radius=radius,

labels=c("Decrease\n[-35.7% to -20.0%]", "Decrease\n[-19.9% to -0.2%]", "Increase\n[0.2% to 19.9%]",

"Increase\n[20.0% to 30.5%]", "Increase\n[50.5% to 60.7%]"),

cex=1.2, init.angle=18,

main="Bureau of Labor Statistics: 2019-2029 Employment Projection Categories")

}

legend.network(WA.pal)
plot.background = element_rect(fill = "transparent", color = NA)
```

```{r}
legend.network(WA.pal)
```

```{r}
WA.pal
```


#function for printing network stats

```{r}
get_network_stats <- function(network.obj) {
  net_size <- vcount(network.obj)
  net_dens <- edge_density(network.obj)
  net_comp <- count_components(network.obj)
  net_diam <- diameter(network.obj)
  net_cluster <- transitivity(network.obj)
  return(paste("Size: ", net_size, "Density: ", net_dens, "Number of Components: ", net_comp, "Diameter: ", net_diam, "Clustering Coefficient: ", net_cluster))
}
```


## Making the overall network data and its attributes

```{r}
cert <- read_xlsx("~/dspg21stw/src/ONET_SOC_credentials.xlsx")
stw_change_pct <- read_xlsx("~/dspg21stw/src/STW_2021_Change.xlsx")
```

```{r}
cert <- cert[-1]

stw_change_pct <- stw_change_pct %>% select(SOC, GROWTH, PER_CHG, SOCNAME = NAME)

cert <- cert %>% select(SOC, Certification.Name, Certifying.Organization, Type)

d <- cert %>% left_join(stw_change_pct, by = "SOC")
```



```{r}
d <- d %>% select(SOCNAME, Certification.Name, SOC, Certifying.Organization, Type, GROWTH, PER_CHG)
```


```{r}
d <- d %>% mutate(SOC_type = substr(SOC, start = 1, stop = 2))
```


```{r}
d <- d %>% mutate(soc_type_name = case_when(SOC_type == "11" ~ "Management Occupations",
                                            SOC_type == "13" ~ "Business and Financial Operations Occupations",
                                            SOC_type == "15" ~ "Computer and Mathematical Occupations",
                                            SOC_type == "17" ~ "Architecture and Engineering Occupation",
                                            SOC_type == "19" ~ "Life, Physical, and Social Science Occupations",
                                            SOC_type == "27" ~ "Art, Design, Entertainment, Sports, And Media Occupations",
                                            SOC_type == "29" ~ "Healthcare Practitioners and Technical Occupation",
                                            SOC_type == "35" ~ "Protective Service Occupations",
                                            SOC_type == "43" ~ "Office and Administrative Support Occupations",
                                            SOC_type == "45" ~ "Farming, Fishing, and Forestry Occupations",
                                            SOC_type == "47" ~ "Construction and Extraction Occupations",
                                            SOC_type == "49" ~ "Installation, Maintenance, and Repair Occupations",
                                            SOC_type == "51" ~ "Production Occupations",
                                            SOC_type == "53" ~ "Transportation and Material Moving Occupations"))
```


```{r}
d <- d %>% mutate(soc_type_name = gsub(" ", "\n", soc_type_name),
                  SOCNAME = gsub(" ", "\n", SOCNAME))
```


^^ use this as the original df.  An attributes matrix now:


labels=c("Decrease\n[-35.7% to -20.0%]", "Decrease\n[-19.9% to -0.2%]", "Increase\n[0.2% to 19.9%]",

"Increase\n[20.0% to 30.5%]", "Increase\n[50.5% to 60.7%]")

```{r}
d <- d %>% mutate(PER_CHG_CAT = case_when(PER_CHG <= -20.0 ~ "highly decreasing",
                                          PER_CHG > -20.0 & PER_CHG <= -0.2 ~ "decreasing",
                                          PER_CHG > -0.2 & PER_CHG <= 19.9 ~ "increasing",
                                          PER_CHG > 19.9 & PER_CHG <= 30.5 ~ "highly increasing",
                                          PER_CHG > 30.5 ~ "very highly increasing"))

```

```{r}
d$PER_CHG_CAT <- as.factor(d$PER_CHG_CAT)
levels(d$PER_CHG_CAT) <- c("highly decreasing", "decreasing", "increasing", "highly increasing", "very highly increasing")
```

```{r}
d_attributes1 <- d %>% select(name = SOCNAME, SOC, SOC_type, Certifying.Organization, SOC_type, soc_type_name, GROWTH, PER_CHG_CAT)
d_attributes2 <- d %>% select(name = Certification.Name, SOC, SOC_type, Certifying.Organization, SOC_type, soc_type_name, GROWTH, PER_CHG_CAT)

d_attributes1 <- d_attributes1[!duplicated(d_attributes1$name),]
d_attributes2 <- d_attributes2[!duplicated(d_attributes2$name),]

d_attributes <- rbind(d_attributes1, d_attributes2)

d <- d[!duplicated(d),]
```


```{r}
ggplot(d, aes(PER_CHG_CAT)) + geom_bar(fill = WA.pal[5])
```


```{r}
range(as.numeric(d$PER_CHG), na.rm = T)
```


```{r}
stw_network <- graph.data.frame(d, directed = F, vertices = d_attributes)
V(stw_network)$type <- bipartite.mapping(stw_network)$type
stw_network.pr <- bipartite.projection(stw_network)
```

## Cybersecurity Only

```{r}
cyber <- read_xlsx("Cyber.xlsx")
```

```{r}
cybermat_attributes1 <- cyber %>% select(name = ONETSOCTITLE, STW, Level, InDemand)
cybermat_attributes2 <- cyber %>% select(name = Certification, STW, Level, InDemand)
cybermat_attributes1 <- cybermat_attributes1[!duplicated(cybermat_attributes1$name),]
cybermat_attributes2 <- cybermat_attributes2[!duplicated(cybermat_attributes2$name),]
cybermat_attributes <- rbind(cybermat_attributes1, cybermat_attributes2)
cybermat_attributes <- cybermat_attributes[!duplicated(cybermat_attributes$name),]

cybermat_attributes <- cybermat_attributes %>% mutate(socname = case_when(name == "Forensic Science Technicians" ~ "Forensic\nScience\nTechnicians",
                                                                          name == "Network and Computer Systems Administrators" ~ "Network\nand Computer\nSystems\nAdministrators",
                                                                          name == "Information Security Analysts" ~ "Information\nSecurity\nAnalysts",
                                                                          name == "Computer Systems Analysts" ~ "Computer\nSystems\nAnalysts",
                                                                          name == "Software Developer" ~ "Software\nDeveloper",
                                                                          name == "Software Quality Assurance Analysts and Testers"~ "Software\nQuality\nAssurance\nAnalysts\nand Testers",
                                                                          name == "Computer Programmers" ~ "Computer\nProgrammers",
                                                                          name == "Information Technology Project Managers" ~ "Information\nTechnology\nProject\nManagers",
                                                                          name == "Computer and Information Systems Managers" ~ "Computer\nand\nInformation\nSystems\nManagers",
                                                                          name == "Computer User Support Specialists" ~ "Computer\nUser\nSupport\nSpecialists",
                                                                          name == "Computer Network Support Specialists" ~ "Computer\nNetwork\nSupport\nSpecialists",
                                                                          name == "Electrical and Electronic Engineering Technologists and Technicians" ~ "Electronic\nEngineering\nTechnicians",
                                                                          name == "Information Security Engineers" ~ "Information\nSecurity\nEngineers",
                                                                          name == "Intelligence Analyst" ~ "Intelligence\nAnalyst",
                                                                          name == "Detectives and Criminal Investigators" ~ "Detectives\nand Criminal\nInvestigators",
                                                                          name == "Police Identification and Records Officers" ~ "Police\nIdentification\nand Records\nOfficers"))

cyber <- cyber[!duplicated(cyber[1:2]),]

cyber <- cyber %>% select(ONETSOCTITLE, Certification, STW, Origin, Certifier, Level, InDemand)
```





## Health Sciences


#creating a graph of only the nursing credentials

```{r nurse graph, fig.height = 20, fig.width = 20}
plot(nurse_net,
     vertex.shape = ifelse(V(nurse_net)$type, "circle", "square"),
     vertex.color = ifelse(V(nurse_net)$type, cbPalette[4], cbPalette[5]),
     vertex.size = ifelse(V(nurse_net)$type, 5, 15),
     vertex.label = ifelse(V(nurse_net)$type == F, V(nurse_net)$socname, V(nurse_net)$name),
     vertex.label.color = "black",
     vertex.label.dist = ifelse(V(nurse_net)$type, -.75, 0),
     vertex.label.cex = ifelse(V(nurse_net)$type, .75, 1),
     edge.color = cbPalette.blk[1],
     edge.width = 1.5)
title(main = "Nursing Credential and Occupation Network", sub = "data from O*Net OnLine")
legend("bottomleft", legend = c("Credential", "Occupation"), col = c(cbPalette[4], cbPalette[5]), pch = c(19,15), pt.cex = 1.5, bty = "n", title = "Node Type", xjust = 0)
```


## Repeat the above method to subset to any type of occupation using SOC_type or some other detail



# Entire Network

```{r}
d
```




```{r, fig.height = 20, fig.width = 20}
growth_color <- as.factor(E(stw_network)$GROWTH)
plot(stw_network,
     vertex.label = ifelse(V(stw_network)$type == F, V(g_test)$name, NA),
     vertex.label.color = "black",
     vertex.label.cex = 0.5,
     vertex.shape = ifelse(V(stw_network)$type, "circle", "square"),
     vertex.color = ifelse(V(stw_network)$type, cbPalette[8], cbPalette[5]),
     vertex.size = ifelse(V(stw_network)$type, .6, 2),
     edge.width = 2,
     edge.color = ifelse(E(stw_network)$GROWTH == "NEG", cbPalette[7], cbPalette[4]),
     layout = layout_with_kk)
title(main = "STW Credential and Occupation Network", sub = "data from O*Net OnLine")
legend("bottomleft", legend = c("Credential", "Occupation"), col = c(cbPalette[8], cbPalette[5]), pch = c(19,15), pt.cex = 1.5, bty = "n", title = "Node Type", xjust = 0)
legend("bottomleft", legend = c("Positive Projected Growth", "Negative Projected Growth"), col = c(cbPalette[4], cbPalette[7]), pch = c(4,4), pt.cex = 1.5, bty = "n", title = "Edge Type", inset = c(0, .05), xjust = 0)
```

So, we can see that the ones that have a positive projected growth are more connected to one another in the large component, and the ones that are less connected tend to be farther out.

consider how we can make this one interactive^^

```{r}
vcount(stw_network)
```

```{r}
ecount(stw_network)
```


```{r}
head(d)
```

```{r}
E(stw_network)$PER_CHG <- as.factor(E(stw_network)$PER_CHG)
```


```{r}
E(stw_network)$color <- colour_values(E(stw_network)$PER_CHG, palette = "rdylgn")
```


```{r, fig.height = 20, fig.width = 20}

plot(stw_network,
     vertex.label = NA,
     vertex.label.color = "black",
     vertex.label.cex = 0.5,
     vertex.shape = ifelse(V(stw_network)$type, "circle", "square"),
     vertex.color = ifelse(V(stw_network)$type, cbPalette[8], cbPalette[5]),
     vertex.size = ifelse(V(stw_network)$type, .75, 2),
     edge.width = 2,
     edge.color = E(stw_network)$color,
     layout = layout_with_graphopt)
title(main = "STW Credential and Occupation Network", sub = "data from O*Net OnLine")
legend("bottomleft", legend = c("Credential", "Occupation"), col = c(cbPalette[8], cbPalette[5]), pch = c(19,15), pt.cex = 1.5, bty = "n", title = "Node Type", xjust = 0)
legend("bottomleft", legend = c("Positive Projected Growth", "Negative Projected Growth"), col = c(cbPalette[4], cbPalette[7]), pch = c(4,4), pt.cex = 1.5, bty = "n", title = "Edge Type", inset = c(0, .05), xjust = 0)
```

```{r}

unique(E(stw_network)$PER_CHG_CAT)
```

```{r}
E(stw_network)$color <- ifelse(E(stw_network)$PER_CHG_CAT == "highly decreasing", WA.pal[1], ifelse(E(stw_network)$PER_CHG_CAT == "decreasing", WA.pal[2], ifelse(E(stw_network)$PER_CHG_CAT == "increasing", WA.pal[3], WA.pal[4])))
```



```{r, fig.height = 20, fig.width = 20}
#E(stw_network)$color <- colour_values(E(stw_network)$PER_CHG_CAT, palette = WA.pal)
plot(stw_network,
     vertex.label = NA,
     vertex.label.color = "black",
     vertex.label.cex = 0.5,
     vertex.shape = ifelse(V(stw_network)$type, "circle", "square"),
     vertex.color = ifelse(V(stw_network)$type, cbPalette[8], cbPalette[5]),
     vertex.size = ifelse(V(stw_network)$type, 1, 4),
     edge.width = 2,
     edge.color = E(stw_network)$color,
     layout = layout_with_gem, xaxs = i, yaxs = i)
legend("topright", legend = c("STW Occupations: yellow squares", "Certifications: pink circles"), bty = "n", title = "Legend", xjust = 0, cex = 1.5)
```

```{r}
get_network_stats(stw_network)
```


## subsetting to each broad occupation group

#Already did it for Healthcare Practitioners and Technical Occupations, Art, Design, Entertainment, Sports, and Media Occupations, Architechture and Engineering Occupations

```{r}
unique(d$soc_type_name)
```


```{r}
length(unique(d_temp$SOCNAME)) + length(unique(d_temp$Certification.Name))
length(unique(d_attr_temp$name))
```

```{r}
d_temp$Certification.Name[which(! d_temp$Certification.Name %in% d_attr_temp$name)]
```

```{r}
d_temp[d_temp$Certification.Name ==  "Commercial Electrical Inspector",]
d_attributes
```



ideas: make each of these independently, so that if I pick a color for each subgroup, we could maintain those colors in the bigger graph, or draw lines around them too I suppose.
if I make them independently, it will only be a matter of naming them or lumping them together.  Then we could just direct the shiny app to be to the code for that one.

#Architecture and Engineering

```{r}
d_arch<- d[d$soc_type_name == "Architecture\nand\nEngineering\nOccupation",]
#d_arch <- na.omit(d_repair)
```


```{r}
d_arch <- d_arch %>% select(-Type)
d_arch <- na.omit(d_arch)
d_arch <- d_arch[!duplicated(d_arch),]
arch <- graph.data.frame(d_arch)


nodelist <- data.frame(name = igraph::V(arch)$name)

nodelist <- nodelist %>% left_join(d_attributes, by = "name")

arch_net <- graph.data.frame(d_arch, directed = F, vertices = nodelist)
V(arch_net)$type <- bipartite.mapping(arch_net)$type
```


```{r}
nodelist
d_arch
```

```{r}
d_arch <- d_arch %>% select(-Type)

```

```{r}
d_arch[d_arch$SOCNAME == "Surveying\n&\nMapping\nTechnicians",]
table(duplicated(d_arch))
```


```{r, fig.height = 25, fig.width= 25}
E(arch_net)$color <- colour_values(E(arch_net)$PER_CHG, palette = "WA.pal")
plot(arch_net,
     vertex.size = ifelse(V(arch_net)$type, 2, 9),
     vertex.shape = ifelse(V(arch_net)$type, "circle", "square"),
     vertex.color = ifelse(V(arch_net)$type, cbPalette[6], cbPalette[5]),
     #vertex.color = cbPalette[V(manu_net)$SOC_type],
     vertex.label = ifelse(V(arch_net)$type == F, V(arch_net)$name, NA),
     vertex.label.color = "black",
     vertex.label.cex = 0.5,
     edge.width = 2,
     edge.color = E(arch_net)$color,
     layout = layout_with_kk)
title("Architecture and Engineering Occupation Network", sub = "data from O*Net OnLine")
legend("bottomleft", legend = c("Credential", "Occupation"), col = c(cbPalette[6], cbPalette[5]), pch = c(19,15), pt.cex = 1.5, bty = "n", title = "Node Type", xjust = 0)
```



#Construction and Extraction Occupations

```{r}
d_const <- d[d$soc_type_name == "Construction\nand\nExtraction\nOccupations",]
d_const <- d_const %>% select(-Type)
d_const <- d_const[!duplicated(d_const),]

d_const <- na.omit(d_const)

```

```{r}
d_const[d_const$SOCNAME == "Construction\nand\nBuilding\nInspectors",]
```


```{r}
const <- graph.data.frame(d_const)


nodelist <- data.frame(name = igraph::V(const)$name)

nodelist <- nodelist %>% left_join(d_attributes, by = "name")

const_net <- graph.data.frame(d_const, directed = F, vertices = nodelist)
V(const_net)$type <- bipartite.mapping(const_net)$type
```


```{r, fig.height = 20, fig.width= 20}
E(const_net)$color <- colour_values(E(const_net)$PER_CHG, palette = "viridis")
plot(const_net,
     vertex.size = ifelse(V(const_net)$type, 2, 10),
     vertex.shape = ifelse(V(const_net)$type, "circle", "square"),
     vertex.color = ifelse(V(const_net)$type, cbPalette[6], cbPalette[5]),
     #vertex.color = cbPalette[V(manu_net)$SOC_type],
     vertex.label = ifelse(V(const_net)$type == F, V(const_net)$name, NA),
     vertex.label.color = "black",
     vertex.label.cex = 0.5,
     edge.width = 2,
     edge.color = E(const_net)$color,
     layout = layout_nicely)
title("Construction and Extraction Occupation Network", sub = "data from O*Net OnLine")
legend("bottomleft", legend = c("Credential", "Occupation"), col = c(cbPalette[6], cbPalette[5]), pch = c(19,15), pt.cex = 1.5, bty = "n", title = "Node Type", xjust = 0)
```

```{r}
get_network_stats(const)
```


#Installation, Maintenance, and Repair Occupations

```{r}
d_repair<- d[d$soc_type_name == "Installation,\nMaintenance,\nand\nRepair\nOccupations",]
d_repair <- na.omit(d_repair)
```


```{r}
rep <- graph.data.frame(d_repair)

nodelist <- data.frame(name = igraph::V(rep)$name)

nodelist <- nodelist %>% left_join(d_attributes, by = "name")

rep_net <- graph.data.frame(d_repair, directed = F, vertices = nodelist)
V(rep_net)$type <- bipartite.mapping(rep_net)$type
```

```{r, fig.height = 20, fig.width=20}
E(rep_net)$color <- colour_values(E(rep_net)$PER_CHG, palette = "viridis")
plot(rep_net,
     vertex.size = ifelse(V(rep_net)$type, 2, 13),
     vertex.shape = ifelse(V(rep_net)$type, "circle", "square"),
     vertex.color = ifelse(V(rep_net)$type, cbPalette[6], cbPalette[5]),
     #vertex.color = cbPalette[V(manu_net)$SOC_type],
     vertex.label = ifelse(V(rep_net)$type == F, V(rep_net)$name, NA),
     vertex.label.color = "black",
     vertex.label.cex = 0.6,
     edge.width = 2,
     #edge.color = E(prod)$color,
     layout = layout_nicely)
title("Installation, Maintenance, and Repair Occupation Network", sub = "data from O*Net OnLine")
legend("bottomleft", legend = c("Credential", "Occupation"), col = c(cbPalette[6], cbPalette[5]), pch = c(19,15), pt.cex = 1.5, bty = "n", title = "Node Type", xjust = 0)
```

```{r}
get_network_stats(rep)
```


```{r}
d_repair
```


# "Production\nOccupations"

```{r}
d_prod <- d[d$soc_type_name == "Production\nOccupations",]
d_prod <- na.omit(d_prod)
d_prod_attr <- d_attributes[d_attributes$soc_type_name == "Production\nOccupations",]
```

```{r}
prod <- graph.data.frame(d_prod)
#prod <- na.omit(prod)
#d_prod_attr <- na.omit(d_prod_attr)
```

```{r}
nodelist <- data.frame(name = igraph::V(prod)$name)

nodelist <- nodelist %>% left_join(d_attributes, by = "name")

```

```{r}
prod <- graph.data.frame(d_prod, directed = F, vertices = nodelist)
V(prod)$type <- bipartite.mapping(prod)$type
```


```{r, fig.height = 20, fig.width=20}
E(prod)$color <- colour_values(E(prod)$PER_CHG, palette = "viridis")
plot(prod,
     vertex.size = ifelse(V(prod)$type, 2, 15),
     vertex.shape = ifelse(V(prod)$type, "circle", "square"),
     vertex.color = ifelse(V(prod)$type, cbPalette[6], cbPalette[5]),
     #vertex.color = cbPalette[V(manu_net)$SOC_type],
     vertex.label = ifelse(V(prod)$type == F, V(prod)$name, NA),
     vertex.label.color = "black",
     vertex.label.cex = 0.6,
     edge.width = 2,
     edge.color = E(prod)$color,
     layout = layout_nicely)
title("Production Occupation Network", sub = "data from O*Net OnLine")
legend("bottomleft", legend = c("Credential", "Occupation"), col = c(cbPalette[6], cbPalette[5]), pch = c(19,15), pt.cex = 1.5, bty = "n", title = "Node Type", xjust = 0)
```

```{r}
get_network_stats(prod)
```


#"Transportation\nand\nMaterial\nMoving\nOccupations"

```{r}
d_transport <- d[d$soc_type_name == "Transportation\nand\nMaterial\nMoving\nOccupations",]
d_transport <- na.omit(d_transport)
```

```{r}
trans <- graph.data.frame(d_transport)

nodelist <- data.frame(name = igraph::V(trans)$name)

nodelist <- nodelist %>% left_join(d_attributes, by = "name")

trans_net <- graph.data.frame(d_transport, directed = F, vertices = nodelist)
V(trans_net)$type <- bipartite.mapping(trans_net)$type
```

```{r, fig.height = 20, fig.width=20}
E(trans_net)$color <- colour_values(E(trans_net)$PER_CHG, palette = "viridis")
plot(trans_net,
     vertex.size = ifelse(V(trans_net)$type, 2, 10),
     vertex.shape = ifelse(V(trans_net)$type, "circle", "square"),
     vertex.color = ifelse(V(trans_net)$type, cbPalette[6], cbPalette[5]),
     #vertex.color = cbPalette[V(manu_net)$SOC_type],
     vertex.label = ifelse(V(trans_net)$type == F, V(trans_net)$name, NA),
     vertex.label.color = "black",
     vertex.label.cex = 0.6,
     edge.width = 2,
     edge.color = E(prod)$color,
     layout = layout_nicely)
title("Transportation and Material Moving Occupation Network", sub = "data from O*Net OnLine")
legend("bottomleft", legend = c("Credential", "Occupation"), col = c(cbPalette[6], cbPalette[5]), pch = c(19,15), pt.cex = 1.5, bty = "n", title = "Node Type", xjust = 0)
```

```{r}
get_network_stats(trans_net)
```




