---
title: "HT_network_analysis"
author: "Haleigh Tomlin"
date: "7/7/2021"
output: html_document
---

# Exploratory Network Analysis of Credentials and SOC codes


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# install.packages("statnet")
# install.packages("igraph")
# install.packages("tidyverse")
# install.packages("devtools")
# install.packages("intergraph")
```

```{r}
library(statnet)
#library(tidyverse)
library(devtools)
library(readxl)
library(intergraph)
```


```{r}
cert <- read_xlsx("ONET_SOC_credentials.xlsx")
```

```{r}
head(cert)
```


We are interested in the usefullness of certifications.  How portable are they?  If you get a certification, how many different jobs can you do?  What are they?

To answer this, we will have "certification" as our nodes, and edges will be jobs.  This way, we can assess the centrality of a certification to jobs, and how connected certifications are within job areas.  Then, we can create a list of certifications that are only useful in unique occupations.

Making an edge-list:

```{r}
netmat <- cbind(cert$Certification.Name, cert$SOC)
```

```{r}
head(netmat)
```

```{r}
soc_type <- substr(netmat[,2], start = 1, stop = 2)
```

```{r}
unique(soc_type)
```

```{r}
netmat2 <- cbind(netmat, soc_type)
```

```{r}
head(netmat2)
```



```{r}
n_orig <- network(netmat, matrix.type = "edgelist", directed = F)
```


```{r}
detach(package:statnet)
library(igraph)

```

```{r}
net.df <- data.frame(netmat)

affiliation <- graph.data.frame(net.df, directed = FALSE)

V(affiliation)$type <- V(affiliation)$name %in% net.df[,1]
```



```{r}
op <- par(mar = rep(0,4))
shapes <- c("circle", "square")
colors <- c("blue", "red")

plot(affiliation, vertex.color = colors[V(affiliation)$type+1],
     vertex.shape = shapes[V(affiliation)$type+1],
     vertex.size = 5, vertex.label = NA, layout = layout_with_kk)#, vertex.label.degree = -pi/20,
     #vertex.label.dist = 1.2, vertex.label.cex = 0.9)
par(op)
```

```{r}
affiliation.pr <- bipartite.projection(affiliation)
```

```{r}
affiliation.pr
```

```{r}
class(affiliation)
```

```{r}
plt.x <- c(rep(2,6), rep(4,4))
plt.y <- c(7:2,6:3)
lay <- as.matrix(cbind(plt.x, plt.y))
```

```{r}
summary(affiliation)
```

```{r, fig.height = 10}
shapes <- c("circle", "square")
colors <- c("blue", "red")

plot(affiliation, vertex.color = colors[V(affiliation)$type + 1],
     vertex.shape = shapes[V(affiliation)$type + 1],
     vertex.size = 8, vertex.label = NA, layout = layout_with_kk)
```


```{r}
list.edge.attributes(affiliation.pr)
```


```{r}
#proj 1 gives the edges
affiliation.pr$proj1
```

```{r}
#proj 2 gives the nodes
affiliation.pr$proj2
```



```{r}
graph.density(affiliation.pr$proj1)
```

```{r}
graph.density(affiliation.pr$proj2)
```

```{r}
net.soc <- affiliation.pr$proj1
net.cred <- affiliation.pr$proj2
```

```{r}
soc_adjacency <- get.adjacency(net.soc, sparse = FALSE, attr = "weight")
```

```{r}
cred_adjacency <- get.adjacency(net.cred, sparse = FALSE, attr = "weight")
```

```{r, fig.height = 10}
shapes <- c("circle", "square")
colors <- c("blue", "red")

op <- par(mfrow = c(1,2))
plot(net.cred, vertex.color = "blue",
     vertex.shape = "circle", main = "Credentials",
     vertex.size = 3, vertex.label = NA)

plot(net.soc, vertex.color = "red",
     vertex.shape = "square", main = "SOC codes",
     vertex.size = 7, vertex.label = NA)
par(op)
```



```{r}
detach(package:igraph)
library(statnet)
net <- network(cred_adjacency, matrix.type = "adjacency", directed = F)
```

```{r}
summary(net, print.adj = F)
```

```{r}
gplot(net, usearrows = F, displaylabels = F)
```



```{r}
summary(net, print.adj = F)
```

Not a super dense network but there are certainly clusters.

```{r}
gden(net)
```


```{r}
components(net)
```

That doesnt make sense, there should be less components than there are nodes

```{r}
lgc <- component.dist(net)
gd <- geodist(lgc)
max(gd$gdist)
```
 Having some trouble getting the # of components, and the diameter of the largest component.
 
```{r}
component.dist(net)
```

Clustering coefficient:

```{r}
gtrans(net, mode = "graph")
```

```{r, fig.height = 10}
op <- par(mar = rep(1,4))
gplot(net, displaylabels = F,
      usearrows = F,
      main = "Network of ONET Certifications",
      vertex.col = "lightblue", vertex.cex = 1.3,
      label.cex = 0.4, label.pos = 5)

par(op)
```

```{r}
net_igraph <- asIgraph(net)
```

Next I will make nodes bigger if they have higher centrality

```{r}
deg <- degree(net_igraph)
cls <- closeness(net_igraph)
bet <- betweenness(net_igraph)
```

```{r}
rescale <- function(nchar, low, high) {
  min_d <- min(nchar)
  max_d <- max(nchar)
  rscl <- ((high - low)* (nchar - min_d)) / (max_d - min_d) + low
  rscl
}
```


```{r, fig.height = 10, dev = 'png'}
op <- par(mar = rep(1,4))
gplot(net, displaylabels = F,
      usearrows = F,
      main = "Network of ONET Certifications",
      vertex.col = "lightblue", vertex.cex = rescale(deg,1,6))

par(op)
```




trying to make a table of the Centrality of each credential (modeled off page 99 in Luke networks book. https://link.springer.com/book/10.1007/978-3-319-23883-8)


#detach(package:statnet)
library(igraph)

df.prom <- data.frame(
  degree = degree(net_igraph),
  closeness = closeness(net_igraph),
  betweenness = betweenness(net_igraph)
)

#row.names(df.prom) <- net_igraph %v% "vertex.names"

df.promsort <- df.prom[order(-df.prom$degree),]
cd <- centralization(net_igraph, degree)
cc <- centralization(net_igraph, closeness)
cb <- centralization(net_igraph, betweenness)

df.promsort <- rbind(df.promsort, c(cd,cc,cb))

row.names(df.promsort)[18] <- "\\emph{Centralization}"


```{r}
#detach(package:igraph)
library(statnet)
```

```{r}
list.network.attributes(net)
```

```{r}
list.vertex.attributes(net)
```



