---
title: "CoreLogic Fairfax Profile"
output: html_document
---

```{r libraries, include=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE, warning = FALSE)
library(dplyr)
library(ggplot2)
library(purrr)
library(sf)  
# install.packages('venn')
library(venn)
# install.packages('ggthemes')
library(ggthemes)
# install.packages('DataExplorer')
library(DataExplorer)
library(maditr)
```

```{r functions}
# functions to calculate completeness (from Devika)
check_calc <- function(vec) {
  blanks <- 0L
  true_na <- 0L
  written_na <- 0L
  len <- length(x = vec)
  for (elem in vec) {
    if (is.na(x = elem)) {
      true_na <- true_na + 1L
    } else if (elem == "na") {
      written_na <- written_na + 1L
    } else if (elem == "") {
      blanks <- blanks + 1L
    }
  }
  percent_complete <- round(((len - (blanks + true_na + written_na)) / len) * 100, 2)
  unique_values <- length(unique(vec))
  tibble(blanks = blanks,
         true_na = true_na,
         written_na = written_na,
         percent_complete = percent_complete,
         unique_values = unique_values)
}
check_complete <- function(df) {
  z <- deparse(substitute(df))
  map_df(.x = df, .f = check_calc) %>%
    mutate(column = colnames(df)) %>%
    mutate(set = print(z))  %>%
    select(set, column, blanks, true_na, written_na, percent_complete, unique_values)
}
```

```{r}
get_db_conn <-
  function(db_name = "sdad",
           db_host = "postgis1",
           db_port = "5432",
           db_user = Sys.getenv("db_usr"), # requires you to setup environmental vars (above)
           db_pass = Sys.getenv("db_pwd")) {
    RPostgreSQL::dbConnect(
      drv = RPostgreSQL::PostgreSQL(),
      dbname = db_name,
      host = db_host,
      port = db_port,
      user = db_user,
      password = db_pass
    )
  }

con <- get_db_conn()

cols_usda_curr_tax <- "fips_code,unformatted_apn,apn_sequence_nbr,original_apn,census_tract,lot_number,township,municipality_code,situs_house_number_prefix,situs_house_number,situs_house_number_2, situs_house_number_suffix,situs_direction,situs_street_name,situs_mode,situs_quadrant,situs_unit_number,situs_city,situs_state,situs_zip_code,situs_carrier_code, property_centroid_longitude,property_centroid_latitude,block_level_latitude,block_level_longitude,owner_corporate_indicator,owner1_last_name,owner1_first_name_mi,owner2_last_name,owner2_first_name_mi, mail_house_number_prefix,mail_house_number,mail_house_number_2,mail_house_number_suffix,mail_direction,mail_street_name,mail_mode,mail_quadrant,mail_unit_number,mail_city,mail_state,mail_zip_code,mail_carrier_code,mailing_opt_out_code,total_value_calculated,land_value_calculated,improvement_value_calculated,total_value_calculated_ind,land_value_calculated_ind,improvement_value_calculated_ind,assd_total_value,assd_land_value,assd_improvement_value,mkt_total_value,mkt_land_value,mkt_improvement_value,appr_total_value,appr_land_value,appr_improvement_value,tax_amount,tax_year,assessed_year,tax_code_area,sale_date,sale_price,sale_code,rior_sale_date,prior_sale_price,prior_sale_code,acres,land_square_footage,universal_building_square_feet,building_square_feet_ind,building_square_feet,living_square_feet,ground_floor_square_feet,gross_square_feet,adjusted_gross_square_feet,year_built,effective_year_built,bedrooms,total_rooms,total_baths_calculated,total_baths,full_baths,half_baths,1qtr_baths,3qtr_baths FROM corelogic_usda.corelogic_usda_current_tax_2020_06_27"

#cols_sdad_tax_hist <- "fips_code, apn__parcel_number_unformatted_, apn_sequence_number, original_apn, census_tract, legal_lot_number, township, municipality_code, property_indicator_code, assessed_total_value, market_total_value, tax_amount, tax_year, assessed_year, acres, land_square_footage, building_square_feet, living_square_feet, year_built, effective_year_built, bedrooms, total_baths, full_baths, half_baths, parcel_level_latitude, parcel_level_longitude, block_level_latitude, block_level_longitude, situs_house_number, situs_street_name, situs_mode, situs_city, situs_state, situs_zip_code FROM corelogic_sdad.tax_hist_2_51"

state_fips <-
  c(
  "02",
  "06",
  "08",
  "13",
  "16",
  "19",
  "26",
  "27",
  "36",
  "37",
  "39",
  "40",
  "41",
  "48",
  "51",
  "53"
  )

for (sfp in state_fips[1:3]) {
  state_county_fips_sql <-
  paste0(
  "select distinct fips_code from corelogic_usda.corelogic_usda_current_tax_2020_06_27 where fips_code like '",
  sfp,
  "%'"
  )
  state_county_fips <-
  DBI::dbGetQuery(con, statement = state_county_fips_sql)
  
  #fips <- "51059"
  for (f in state_county_fips$fips_code) {
  sql <- paste0("SELECT ",
  cols_usda_curr_tax,
  " WHERE fips_code='", f, "'")
  
  assign(paste0("_", f), DBI::dbGetQuery(con, statement = sql))
  }
}

#ffx <- DBI::dbGetQuery(con, statement = sql)
DBI::dbDisconnect(con)
```



```{r}
ffx_subset <- select(ffx, 1:22)
```

```{r}
#check completeness
lfo <- list()
for (f in ls(pattern = "^_")) {
  assign(paste0(f, "_"), check_complete(get(f)))
  t <- get(paste0(f, "_"))
  t$st_cnty_fips <- substr(f, 2, 6)
  lfo <- c(lfo, list(t))
}

lfo_dt <- data.table::rbindlist(lfo)

#check_complete(`_51001`)
```

```{r}
dt <- lfo_dt[,.(st_cnty_fips, column, blanks, true_na, written_na, percent_complete, unique_values)]
con <- get_db_conn()
DBI::dbWriteTable(con, name = c("corelogic_usda", "corelogic_usda_current_tax_2020_06_27_counts"), value = dt, row.names = F, append = T)
DBI::dbDisconnect(con)
```


```{r}
introduce(ffx_subset)
```

```{r}
plot_missing(ffx_subset)
```



```{r}
plot_bar(ffx_subset)
```


```{r}
ffx_sub_numeric <- mutate_all(ffx_subset, function(x) as.numeric(as.character(x)))
sapply(ffx_sub_numeric, class)
```

```{r}
ffx_sub_numeric %>%
  ggplot(aes(year_built)) +
  geom_histogram()
```

**completeness of variables**

```{r}
ffx_1 <- ffx[1:10]

ffx_1 <- dt_mutate(ffx_1, assessed_total_value = as.numeric(assessed_total_value))
```


```{r}
#completeness of variables
check_complete(ffx_1) %>% knitr::kable()
```

```{r}
plot_intro(ffx_1)
```

```{r}
plot_missing(ffx_1)
```

**Distribution of property_indicators/ property types**
Most properties are single family residences (10) with the next most popular property type being condominiums (11) and vacant (80). There are no miscellaneous properties (00).

```{r}
prop_indicator_dist_plot <- ggplot(ffx_1, aes(x=property_indicator_code, fill=property_indicator_code)) +
  geom_bar(position="stack") 
prop_indicator_dist_plot
```

**Distribution of assd_total_value for the different property types**
Single family homes have the greatest range in assessed total value, followed by condos and vacant properties (among the three most common property types).
```{r}
dist <- (ffx_1) %>%
  filter(!is.na(assessed_total_value)) %>% 
  filter(property_indicator_code == c("10", "11", "80")) %>%
  group_by(property_indicator_code) %>%
  ggplot(aes(x= assessed_total_value, y=property_indicator_code)) + geom_point() 

dist 
```

Average assessed total value for the three most common property types.
```{r}
dist <- (ffx_1) %>%
  filter(!is.na(assessed_total_value)) %>% 
  filter(property_indicator_code == c("10", "11", "80")) %>%
  group_by(property_indicator_code) %>%
  summarise(mean = mean(assessed_total_value)) %>%
  ggplot(aes(x= mean, y=property_indicator_code)) + geom_point() 

dist 
```

Assd and Mkt values are the same throughout
```{r include= FALSE}
# ffx_1 %>%
#   filter(!is.na(assessed_total_value))%>%
#   filter(!is.na(market_total_value)) %>%
#   mutate(diff = assessed_total_value - market_total_value) 
# ffx_1 <- ffx_1[!grepl("0", ffx_1$diff)]
```


