---
title: "Full Fairfax Profiling"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE, warning = FALSE)
library(dplyr)
library(ggplot2)
library(purrr)
library(sf)  
# install.packages('venn')
library(venn)
# install.packages('ggthemes')
library(ggthemes)
# install.packages('DataExplorer')
library(DataExplorer)
```


## First 6 Variables

```{r functions}
# functions for completeness from Devika
check_calc <- function(vec) {
  blanks <- 0L
  true_na <- 0L
  written_na <- 0L
  len <- length(x = vec)
  for (elem in vec) {
    if (is.na(x = elem)) {
      true_na <- true_na + 1L
    } else if (elem == "na") {
      written_na <- written_na + 1L
    } else if (elem == "") {
      blanks <- blanks + 1L
    }
  }
  percent_complete <- (len - (blanks + true_na + written_na)) / len
  unique_values <- length(unique(vec))
  tibble(blanks = blanks,
         true_na = true_na,
         written_na = written_na,
         percent_complete = percent_complete,
         unique_values = unique_values)
}
check_complete <- function(df) {
  z <- deparse(substitute(df))
  map_df(.x = df, .f = check_calc) %>%
    mutate(column = colnames(df)) %>%
    mutate(set = print(z))  %>%
    select(set, column, blanks, true_na, written_na, percent_complete, unique_values)
}
```

```{r}
#Load data 
library(readr)
fairfax_housing_2018_geo <- read_csv("~/git/dspg20broadbandERS/data/fairfax-data/fairfax_housing_2018_geo.csv")

first_six <- fairfax_housing_2018_geo %>%
            select(PARCEL_ID, NUM_UNITS, HOUSING_TYPE, YEAR_BUILT, LONGITUDE, LATITUDE)
```


```{r eval = F}
# To find the no. of NAs in the variable 
sum(is.na(fairfax_housing_2018_geo$PARCEL_ID))
sum(is.na(fairfax_housing_2018_geo$NUM_UNITS))
sum(is.na(fairfax_housing_2018_geo$HOUSING_TYPE))
sum(is.na(fairfax_housing_2018_geo$YEAR_BUILT))
# 453 NAs in the YEAR_BUILT variable 
sum(is.na(fairfax_housing_2018_geo$LONGITUDE))
sum(is.na(fairfax_housing_2018_geo$LATITUDE))

```


Table for completeness of Variables

```{r}
#completeness of variables
check_complete(first_six) %>% knitr::kable()
```



General exploration of the first 6 variables

```{r}
#General exploration of the variables
plot_intro(first_six)

#distribution of types of housing units
housing_dist_plot <- ggplot(first_six, aes(x=HOUSING_TYPE)) +
  geom_bar(fill="light blue", position="stack") +
                   geom_text(stat = "count", aes(y = ..count.., label = ..count..), position = "identity", vjust=-0.3 ) +
  labs(x = "Hosuing Type", y = "Count")+
  theme(
    axis.title.x = element_text(size = 10, face = "bold"),
    axis.title.y = element_text(size = 10, face = "bold.italic")
  )
housing_dist_plot

# DataExplorer method to plot Housing Type
plot_bar(first_six$HOUSING_TYPE) 

#There are only 8 mobile houses in the entire dataset. 

################################################################################

#distribution of the year houses were built in
plot_histogram(first_six$YEAR_BUILT)

################################################################################

#number of units distribution by Year built
#Most houses were built in the mid to late 1900s. 
num_units <- first_six[, c("NUM_UNITS", "YEAR_BUILT")]
plot_scatterplot(num_units, by="NUM_UNITS")

# 26 houses were built before the year 1800
house_1700s <- first_six %>%
                filter(YEAR_BUILT < "1800")
house_1700s
count(house_1700s)

#Only 1 house was built in most years
house <- house_1700s[, c("NUM_UNITS", "YEAR_BUILT")]
plot_scatterplot(house, by="NUM_UNITS")


```



Distribution of Housing Type Variable

```{r}
#distribution of types of housing units
housing_dist_plot <- ggplot(first_six, aes(x=HOUSING_TYPE, fill=HOUSING_TYPE)) +
  geom_bar( position="stack") +
                   geom_text(stat = "count", aes(y = ..count.., label = ..count..), position = "identity", vjust=-0.3 ) +
  labs(x = "Hosuing Type", y = "Count")+
  theme(
    axis.title.x = element_text(size = 0, face = "bold"),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
    axis.title.y = element_text(size = 10, face = "bold.italic")
  )
housing_dist_plot
```


Distribution of the year houses were built in

```{r} 
#distribution of the year houses were built in
plot_histogram(first_six$YEAR_BUILT)
```


Distribution of the number of units

```{r}
#number of units distribution
#returns strange results
plot_histogram(first_six$NUM_UNITS)

```


Number of units built per year

```{r}
#scatterplot form of the same
num_units <- first_six[, c("NUM_UNITS", "YEAR_BUILT")]
plot_scatterplot(num_units, by="NUM_UNITS") 
```

## Second 6 Variables

Profiling fairfax data, specifically the variables WATER, SEWER, GAS, VALUE_LAND, VALUE_BUILDING, & VALUE_TOTAL.


```{r}
check_complete(select(fairfax_housing_2018_geo, 7:12)) %>% knitr::kable()
```

Trying out DataExplorer package.

```{r}
introduce(fairfax_housing_2018_geo)
plot_missing(fairfax_housing_2018_geo)
```

Plotting frequencies for categorical variables WATER, SEWER, and GAS.

```{r}
f <- ggplot(data = fairfax_housing_2018_geo)
f + geom_bar(mapping = aes(x = WATER, fill = WATER))
f + geom_bar(mapping = aes(x = SEWER, fill = SEWER))
f + geom_bar(mapping = aes(x = GAS, fill = GAS))
```

Looking at value distributions. Lowest and highest 1% are removed.

```{r}
value_total <- na.omit(fairfax_housing_2018_geo[, "VALUE_TOTAL"])
total <- sort(data.matrix(value_total))
lower <- length(total) %/% 100
upper <- length(total) - lower
total_mid <- total[lower:upper]
hist(total_mid)
```

```{r}
value_building <- na.omit(fairfax_housing_2018_geo[, "VALUE_BUILDING"])
building <- sort(data.matrix(value_building))
lower <- length(building) %/% 100
upper <- length(building) - lower
building_mid <- building[lower:upper]
hist(building_mid)
```

```{r}
value_land <- na.omit(fairfax_housing_2018_geo[, "VALUE_LAND"])
land <- sort(data.matrix(value_land))
lower <- length(land) %/% 100
upper <- length(land) - lower
land_mid <- land[lower:upper]
hist(land_mid, xlim = c(29000, 743000))
```

Quick check to see if VALUE_LAND and VALUE_BUILDING sum to VALUE_TOTAL.

```{r}
sum(land) + sum(building)
sum(total)
```


## Final 6 Variables 

```{r}
# select six variables in question
d_six <- select(fairfax_housing_2018_geo, 13:18)
```

We use a custom function to check for completion of the six variables in question. The results are below:

```{r}
# check completeness
check_complete(d_six) %>% knitr::kable()
```

We confirm these results for the six variables in question using the DataExplorer package:

```{r}
introduce(d_six)
```

We plot the missing values of the six variables in question to see proportions:

```{r}
plot_missing(d_six)
```


```{r}
d_six <- update_columns(d_six, "GEOID", as.factor)
```

We plot the continuous variables:

```{r}
plot_histogram(d_six)
```

To read the continuous variables better we plot them individually:

```{r}
ggplot(data = d_six, aes(x=BEDROOMS))+
  geom_histogram()
```


```{r}
ggplot(data = d_six, aes(x=BATHROOMS))+
  geom_histogram()
```

```{r}
ggplot(data = d_six, aes(x=LIVING_AREA))+
  geom_histogram()
```

We plot the categorical variables:

```{r}
plot_bar(d_six)
```

