---
title: "ERS Property Indicator Exploration"
output: html_document
---

```{r echo = FALSE}
library(tidyverse)
library(readr)
library(scales)
library(readxl)
```

```{r include = FALSE}
ffx_original <- read.csv('~/git/dspg20broadbandERS/data/fairfax-data/fairfax_housing_2018_geo.csv')
nk <- readxl::read_excel('NK ASSESSMENT REPORT 2020.xlsx')
```

```{r include = FALSE}
get_db_conn <-
  function(db_name = "sdad",
           db_host = "postgis1",
           db_port = "5432",
           db_user = Sys.getenv("db_usr"), # requires you to setup environmental vars (above)
           db_pass = Sys.getenv("db_pwd")) {
    RPostgreSQL::dbConnect(
      drv = RPostgreSQL::PostgreSQL(),
      dbname = db_name,
      host = db_host,
      port = db_port,
      user = db_user,
      password = db_pass
    )
  }

con <- get_db_conn()

ffx_allyears <- DBI::dbGetQuery(con, statement = paste(
  "SELECT fips_code, apn__parcel_number_unformatted_, apn_sequence_number, original_apn, census_tract, legal_lot_number, township, municipality_code,
  property_indicator_code, assessed_total_value, market_total_value, tax_amount, tax_year, assessed_year, 
  acres, land_square_footage, building_square_feet, living_square_feet, year_built, effective_year_built, 
  bedrooms, total_baths, full_baths, half_baths, parcel_level_latitude, parcel_level_longitude, block_level_latitude, block_level_longitude,
  situs_house_number, situs_street_name, situs_mode, situs_city, situs_state, situs_zip_code",
  "FROM corelogic_sdad.tax_hist_2_51",
  "WHERE fips_code='51059'"))

nk_allyears <- DBI::dbGetQuery(con, statement = paste(
  "SELECT fips_code, apn__parcel_number_unformatted_, apn_sequence_number, original_apn, census_tract, legal_lot_number, township, municipality_code,
  property_indicator_code, assessed_total_value, market_total_value, tax_amount, tax_year, assessed_year, 
  acres, land_square_footage, building_square_feet, living_square_feet, year_built, effective_year_built, 
  bedrooms, total_baths, full_baths, half_baths, parcel_level_latitude, parcel_level_longitude, block_level_latitude, block_level_longitude,
  situs_house_number, situs_street_name, situs_mode, situs_city, situs_state, situs_zip_code",
  "FROM corelogic_sdad.tax_hist_2_51",
  "WHERE fips_code='51127'"))

DBI::dbDisconnect(con)

#check if the assessed_year and the tax_year columns are identical
identical(ffx_allyears[[13]],ffx_allyears[[14]])

#hooray! they are (**IN THIS CASE FOR FAIRFAX** - need to check if applying other counties)

colnames(ffx_allyears)
sapply(ffx_allyears, class)

#filter by 2018
ffx_2018 <- ffx_allyears[ffx_allyears$tax_year == '2018',]
nk_2018 <- nk_allyears[nk_allyears$tax_year == '2018',]
```

# Fairfax County 
## Categories Provided by Fairfax County
```{r echo = FALSE}
ffx_original$HOUSING_TYPE <- as.factor(ffx_original$HOUSING_TYPE)
knitr::kable(summary(ffx_original$HOUSING_TYPE))
```

## CoreLogic Property Indicator Codes Included in Fairfax
```{r}
ffx_2018$property_indicator_code <- as.factor(ffx_2018$property_indicator_code)
knitr::kable(summary(ffx_2018$property_indicator_code))
```

The codes that were excluded from the graphs presented were 00 (Miscellaneous), 20 (Commercial), 23 (Hotel, Motel), 25 (Retail), 26 (Service), 27 (Office Building), 28 (Warehouse), 29 (Financial Institution), 30 (Hospital), 31 (Parking), 32 (Amusement-Recreation), 50 (Industrial), 53) (Transport), 54 (Utilities), 70 (Agricultural), 80 (Exempt), and 90 (Vacant).

Those included were 10 (Single Family Dwelling), 11 (Condo), 21 (Duplex), 22 (Apartment), 24 (Commercial Condo). These are referred to as "residential" rows.

```{r echo=FALSE}
ffx_residential <- filter(ffx_2018, property_indicator_code == 10 | property_indicator_code == 11 | property_indicator_code == 21 | property_indicator_code == 22 | property_indicator_code == 24)
```

## Analysis of residential vs. non-residential (Fairfax County, both datasets)
There are `r sum(is.na(ffx_2018$property_indicator_code))` NA values for the property indicator code in this dataset. The variables that were not included in the plot (non-residential) and were not NA values account for `r nrow(ffx_2018) - nrow(ffx_residential) - sum(is.na(ffx_2018$property_indicator_code))` rows in the CoreLogic data. There are `r nrow(ffx_residential)` residential rows in the CoreLogic data (were included in the plot) and `r nrow(ffx_2018)` total rows in the CoreLogic data. This contrasts with `r nrow(ffx_original)` rows in the data we have from Fairfax County.

# New Kent County

Note that New Kent does *not* provide its own property indicator variable. The following is all analysis of New Kent CoreLogic.

```{r}
nk_2018$property_indicator_code <- as.factor(nk_2018$property_indicator_code)
knitr::kable(summary(nk_2018$property_indicator_code))
```

Like Fairfax, the codes that were excluded from the graphs presented were 20 (Commercial), 23 (Hotel, Motel), 25 (Retail), 26 (Service), 27 (Office Building), 28 (Warehouse), 29 (Financial Institution), 30 (Hospital), 31 (Parking), 32 (Amusement-Recreation), 50 (Industrial), 53) (Transport), 54 (Utilities), 70 (Agricultural), 80 (Exempt), and 90 (Vacant).

Those included were 10 (Single Family Dwelling), 11 (Condo), 21 (Duplex), 22 (Apartment), 24 (Commercial Condo). These are referred to as "residential" rows.

```{r echo = FALSE}
nk_residential <- filter(nk_2018, property_indicator_code == 10 | property_indicator_code == 11 | property_indicator_code == 21 | property_indicator_code == 22 | property_indicator_code == 24)
```


## Analysis of residential vs. non-residential (New Kent County, both datasets)
There are `r sum(is.na(nk_2018$property_indicator_code))` NA values for the CoreLogic property indicator code in this dataset. The variables that were not included in the plot (non-residential) and were not NA values account for `r nrow(nk_2018) - nrow(nk_residential) - sum(is.na(nk_2018$property_indicator_code))` rows in the CoreLogic data. There are `r nrow(nk_residential)` residential rows in the CoreLogic data (were included in the plot) and `r nrow(nk_2018)` total rows in the CoreLogic data. This compares to `r nrow(nk)` rows in the original New Kent data, though again we do not know the property types.


