---
title: "Examining NAs"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

```{r, warning = FALSE}
library(naniar)
library(ggplot2)
library(visdat)
library(dplyr)
ems_full = read.csv(here::here("data", "working", "ems_clean_data.csv"))
```

## NA Counts

```{r}
miss_var_summary(ems_full)
```

### Core Administrative Variables

Get a list of variables that could reasonably exist on every call:

```{r}
# non patient vars
ems_full %>% 
  select(incident_complaint_reported_by_dispatch, 
         response_incident_number,
         response_ems_unit_call_sign,
         incident_date,
         scene_gps_latitude,
         scene_gps_longitude,
         agency_name) %>% 
  gg_miss_var()
```

#### NA Co-ocurrence

```{r}
# non patient vars
ems_full %>% 
  select(incident_complaint_reported_by_dispatch,
         response_incident_number,
         response_ems_unit_call_sign,
         incident_date,
         scene_gps_latitude,
         scene_gps_longitude,
         agency_name) %>%  
  gg_miss_upset(nsets = 7)
```

Show for just albemarle and charlottesville respectively


### Variables Relevant to Analysis

Get a list of variables that would be useful to have for our analysis:

```{r}
# patient vars
ems_full %>% 
  select(patient_race_list, 
         patient_age,
         patient_age_range_in_years,
         patient_gender,
         situation_provider_primary_impression_code_and_description,
         total_unit_response_time,
         scene_gps_latitude,
         scene_gps_longitude,
         incident_psap_call_date_time,
         patient_initial_pulse_oximetry,
         patient_last_pulse_oximetry) %>% 
  gg_miss_var()
```

#### NA Co-ocurrence

```{r}
# non patient vars
ems_full %>% 
  select(patient_race_list, 
         patient_age,
         patient_age_range_in_years,
         patient_gender) %>% 
  # rename(primary_impression = situation_provider_primary_impression_code_and_description) %>%
  gg_miss_upset()
```

If any demographic variable is missing, highly likely all the others are as well. 

### Time Variables
Seems that when total unit time is missing so is arrival time to scene.
```{r}
ems_full %>% 
  select(incident_date,
         incident_psap_call_date_time,
         incident_dispatch_notified_to_unit_arrived_on_scene_in_minutes,
         total_unit_response_time) %>% 
  gg_miss_upset(nsets = 4)

```

```{r}
ems_full %>% 
  select(situation_provider_primary_impression_code_and_description,
        situation_primary_complaint_statement_list,
        situation_chief_complaint_anatomic_location,
        situation_provider_secondary_impression_description_and_code_list,
        situation_complaint_duration,
        situation_complaint_duration_time_units)%>%
  gg_miss_upset(nsets = 7)

```



```{r}
ems_full %>% 
  select(patient_suspected_influenza_type_illness,
         patient_last_pulse_oximetry,
         patient_initial_pulse_oximetry,
         patient_initial_body_temperature_in_fahrenheit,
         patient_last_body_temperature_in_fahrenheit,
         situation_primary_complaint_statement_list) %>%  
  gg_miss_upset(nsets = 6)
```

We see that across races, the density of na's by variable is realitively stable across however by age, there are some varaibles that are fuller. 
```{r}
ems_full$patient_race_list = gsub("(.*?)\\|.*", "\\1", ems_full$patient_race_list)

ems_full %>%
  replace_with_na(replace = list(patient_race_list = c("Not Recorded", "Not Applicable"))) %>% 
  gg_miss_fct(fct = patient_race_list)

ems_full %>% 
  gg_miss_fct(fct = patient_age_range_in_years)

```

## NA Distribution

### Over Index

Do the number of NAs vary with location in the table?

```{r, fig.height= 8}
# note these are printed out of order
map(0:5, function(i) {
  ems_full %>% 
    select(incident_complaint_reported_by_dispatch,
           response_incident_number,
           response_ems_unit_call_sign,
           incident_date,
           scene_gps_latitude,
           scene_gps_longitude,
           agency_name) %>% 
    slice((1 + 20000*i):min((20000*(i + 1)), nrow(ems_full))) %>% 
    vis_miss() +
    labs(title = paste((i*20000 + 1), min((20000*(i + 1)), nrow(ems_full)), sep = ":"))
})
```

Looks like albemarle has a lot more NAs, possibly due to the shifted values on some rows. 

Now lets look at number of NAs per row over index. 

```{r}
ems_full %>% 
  mutate(total_na = rowSums(is.na(ems_full))) %>%
  ggplot() + 
  geom_point(aes(x = 1:nrow(ems_full), y = total_na), alpha = 0.3) + 
  labs(y = "Number of NA's", x = "index") +
  ylim(0, 90) +
  geom_hline(yintercept = 50, color = "green") + # min number of variables to contain all admin data
  theme_minimal()

# As density

ems_full %>% 
  mutate(total_na = rowSums(is.na(ems_full))) %>%
  ggplot() + 
  geom_hex(aes(x = 1:nrow(ems_full), y = total_na), alpha = 1) + 
  labs(y = "Number of NA's", x = "index") +
  ylim(0, 90) +
  geom_hline(yintercept = 50, color = "green") +
  scale_fill_continuous(high = "#132B43", low = "#56B1F7")+
  theme_minimal()
```


```{r}

samp_dat = ems_full %>%
  sample_frac(size=.1)%>%
  vis_dat()
samp_dat

samp_missing = ems_full %>%
  sample_frac(size=.2)%>%
  vis_miss()
samp_missing

ems_full$incident_date = as.POSIXct(ems_full$incident_date)

nas_by_case = ems_full %>% arrange(desc(incident_date)) %>%
  gg_miss_case()
nas_by_case

### part below is not working as desired
ems_full%>%
gg_miss_span(patient_last_pulse_oximetry, span_every = 3000, facet = patient_race_list)

covid_vars = c("patient_suspected_influenza_type_illness",
              "patient_initial_body_temperature_in_fahrenheit",
              "patient_last_body_temperature_in_fahrenheit",
              "situation_primary_complaint_statement_list",
              "patient_last_pulse_oximetry",
              "patient_initial_pulse_oximetry",
              "patient_initial_body_temperature_in_fahrenheit",
              "patient_last_body_temperature_in_fahrenheit")

for (i in covid_vars){
  plot = 
  gg_miss_span(ems_full, as.name(i), span_every = 3000, facet = patient_race_list)+labs(title = as.name(covid_vars[i]))
  print(plot)
}

n=5
ems_full%>%
  gg_miss_span(as.name(covid_vars[n]), span_every = 3000, facet = patient_race_list) + labs(title = as.name(covid_vars[n]))

```
