---
title: "Tournament Theory"
description: "Applying tournament theory to post-military veteran career sequences"
output: html_document
---


#### **Tournament Theory**

Rosenbaum's (1979) tournament model (described in more detail in the Methodology section) argues that career mobility can be represented as rounds of a tournament, where each round has winners and losers and the results of each round affect subsequent career opportunities. We use 10-year career sequences of veterans beginning after their last military job to test hypotheses about career mobility.

We are particularly interested in two potential tournament "results" for veterans: early promotion (promotion within two years of leaving the military) and transitional unemployment (unemployment occurring directly after leaving the military). If these two results can be shown to be associated with positive or negative career outcomes, this could provide evidence in favor of certain types of career development interventions for veterans. As in the rest of our analysis, we use O\*NET job zones (described further in the Data section) to represent different career levels; an increase in job zone is considered a promotion, and a decrease is considered a demotion. We also include "left" unemployment in our sequences to represent transitional unemployment after leaving the military.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE
)

for (pkg in c("tidyverse",  "data.table", "R.utils", "maditr", "stringr", "stringi", "dplyr", "ggplot2", "lubridate", "gt", "DataExplorer", "TraMineR", "TraMineRextras", "cluster", "MASS", "kableExtra")) {
  library(pkg, character.only = TRUE)
}

uva_color_palette <- 
c("#232D4B", #space cadet
  "#2C4F6B", #indigo dye
  "#0E879C", #blue munsell
  "#60999A", #cadet blue
  "#D1E0BF", #tea green
  "#D9E12B", #pear
  "#E6CE3A", #citrine
  "#E6A01D", #marigold
  "#E57200" #princeton orange
)

```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=FALSE}
source("~/git/dspg20career/src/burningglass/clean_post_military_job.R", echo = T, prompt.echo = "", spaced = F)

# bg_vet_job <-read_csv("~/git/dspg20career/data/04_bg_vet_job.csv")
bg_vet_job <-read_csv("../../tmpdata/04_bg_vet_job.csv")
bg_vet_job_seq <- clean_post_military(bg_vet_job)

bg_vet_job_seq <- as.matrix(bg_vet_job_seq)
bg_vet_job_seq <- as.data.frame(bg_vet_job_seq)

sts_vet <- seqformat(bg_vet_job_seq, from = "SPELL", to = "STS",
                     id = "id",  begin = "start_year", end = "end_year", 
                     status = "onet_job_zone", process = FALSE)



# Here we are renaming columns to be in format "yn" (year in the job market)
names(sts_vet) <- paste0("y", 1:ncol(sts_vet))

vet.seq <- seqdef(sts_vet, left="NA", gaps="NA", right="DEL")

# limit to first 5 years
vet.seq.5 <- vet.seq[,1:5]
vet.seq.5 <- as.data.frame(vet.seq.5)
names <- colnames(vet.seq.5)

vet.seq.5[, names] <- sapply(vet.seq.5[, names],function(x) str_replace_all(x,fixed("%"), "NA"))

vet.seq.5.numeric <- vet.seq.5

vet.seq.5.numeric[,names] = data.frame(sapply(vet.seq.5.numeric[,names], as.character))
vet.seq.5.numeric[,names] = data.frame(sapply(vet.seq.5.numeric[,names], as.numeric))


# limit to first 10 years
vet.seq.10 <- vet.seq[,1:10]
vet.seq.10 <- as.data.frame(vet.seq.10)
names <- colnames(vet.seq.10)

vet.seq.10[, names] <- sapply(vet.seq.10[, names],function(x) str_replace_all(x,fixed("%"), "NA"))

vet.seq.10.numeric <- vet.seq.10
vet.seq.10.numeric[,names] = data.frame(sapply(vet.seq.10.numeric[,names], as.character))
vet.seq.10.numeric[,names] = data.frame(sapply(vet.seq.10.numeric[,names], as.numeric))

# limit to first 20 years
vet.seq.20 <- vet.seq[,1:20]
vet.seq.20 <- as.data.frame(vet.seq.20)

vet.seq.20[, names] <- sapply(vet.seq.20[, names],function(x) str_replace_all(x,fixed("%"), "NA"))


vet.seq.20.numeric <- vet.seq.20
names <- colnames(vet.seq.20.numeric)
vet.seq.20.numeric[,names] = data.frame(sapply(vet.seq.20.numeric[,names], as.character))
vet.seq.20.numeric[,names] = data.frame(sapply(vet.seq.20.numeric[,names], as.numeric))
```


```{r include = FALSE}
vet.seq.10 <- vet.seq.10.numeric %>%
  mutate(promoted_first_period = ifelse(((y3 > y2) & !is.na(y3)),
                                        'Promoted First Period', 'Not Promoted First Period'),
        promoted_second_period = ifelse(((y5 > y4) & !is.na(y5)),
                                        'Promoted Second Period', 'Not Promoted Second Period'),
        promoted_third_period = ifelse(((y7 > y5) & !is.na(y7)),
                                       'Promoted Third Period', 'Not Promoted Third Period'),
        promoted_fourth_period = ifelse(((y9 > y5) & !is.na(y9)),
                                         'Promoted Fourth Period', 'Not Promoted Fourth Period'),
        promoted_fifth_period = ifelse(((y10 > y9) & !is.na(y10)),
                                       'Promoted Fifth Period', 'Not Promoted Fifth Period'),
        unemployed_first_period = ifelse((is.na(y1) & is.na(y2)), 
                                         'Unemployed First Period', 'Employed First Period'),
        employed_later_period = ifelse((!is.na(y3) & !is.na(y4) & !is.na(y5) & !is.na(y6) & !is.na(y7) & !is.na(y8)
                                          & !is.na(y9) & !is.na(y10)), 'Emp. Full Y3-Y10 Period','Unemp. Sometime Y3-Y10'),
        unemployed_later_period = ifelse((is.na(y3) | is.na(y4) | is.na(y5) | is.na(y6) | is.na(y7) | is.na(y8)
                                          | is.na(y9) | is.na(y10)), 'Emp. Full Y3-Y10','Unemp. Sometime Y3-Y10'),
        finish_zone_1_2_unemp = ifelse((y10 == 1 | y10 == 2 | is.na(y10)),
                                   'Finish Unemp. or Zone 1, 2', 'Finish Zone 3, 4, or 5'),
        finish_zone_4_5 = ifelse((y10 == 4 | y10 == 5), 'Finish Zone 4 or 5', 'Finish Unemp. or Zone 1, 2, 3')
        
)
```




#### **Hypothesis: Promotion in the first period is associated with promotion in the second period**

To test our first hypothesis, we group our sequences by individuals promoted in the first period (promoted after two years) and individuals promoted in the second period (promoted after four years). If this association exists, it would be evidence that promotion in early "tournament rounds" improves the chances of promotion later.

```{r echo = FALSE}
# promoted first period, promoted second period
tblA = table(vet.seq.10$promoted_first_period, vet.seq.10$promoted_second_period)
tblA %>%
  knitr::kable(caption = "Two Way Frequency Table") %>%
  kable_styling("striped", full_width = FALSE)
```



```{r echo = FALSE}
chisq.test(tblA)
```

The results of a chi-squared test (above) do not show a strong statistical assocation between promotion in the first period and promotion in the second period. This may be because our population (the DC MSA veterans) is already highly educated, and so a veteran is more likely to start in a higher job zone. If an individual starts in job zone four or five, there is not much upward mobility possible.



#### **Hypothesis: Promotion in the first period is associated with employment in a later period**

To test our second hypothesis, we group sequences by veterans who were promoted in the first employment period (promoted after two years), and individuals who were employed for the full remainder of the period (from year 3 to year 10, represented in the table below as Y3-Y10). If this association exists, it is evidence that promotion in early rounds of the tournament results in a positive career outcome (staying employed for the duration of the 10-year career).

```{r echo = FALSE}
# promoted first period, employed full later period
tblB = table(vet.seq.10$promoted_first_period, vet.seq.10$employed_later_period)
tblB %>%
  knitr::kable(caption = "Two Way Frequency Table") %>%
  kable_styling("striped", full_width = FALSE)
```



```{r echo = FALSE}
chisq.test(tblB)
```

The results of the chi-squared test (above) show strong statistical evidence (p < 0.01) that early promotion is associated with employment for the remaining duration of the sequence.



#### **Hypothesis: Transitional unemployment is associated with unemployment later in the sequence**

To test our third hypothesis, we group sequences by veterans who were unemployed for two years following their last military job (the first two years of the sequence - transitional unemployment) and veterans who were unemployed sometime in the remainder of the sequence (year 3 to year 10, represented in the table as Y3-Y10). If this association exists, it is evidence that early unemployment (a "loss" in a tournament round) can cause "losses" in later rounds of the tournament.

```{r echo = FALSE}
# unemployed first period, unemployed later period
tblC = table(vet.seq.10$unemployed_first_period, vet.seq.10$unemployed_later_period)
tblC %>%
  knitr::kable(caption = "Two Way Frequency Table") %>%
  kable_styling("striped", full_width = FALSE)
```



```{r echo = FALSE}
chisq.test(tblC)
```

The results of the chi-squared test (above) show a strong statistical association (p < 0.01) between transitional unemployment and unemployment in the later years of a career sequence; in fact, this is the strongest association we present.


#### **Hypothesis: Transitional unemployment is associated with finishing unemployed or in a lower job zone (job zone 1 or 2)**

To test our third hypothesis, we group sequences by veterans who were unemployed for two years following their last military job (the first two years of the sequence - transitional unemployment) and veterans who finished the ten-year period in either job zone one or two (or were unemployed). If this association exists, it is evidence that early unemployment lowers the career "ceiling" (highest possible career) of a given veteran.

```{r echo = FALSE}
# unemployed first period, zone 1 or 2 or unemployed at end of 10 years
tblD = table(vet.seq.10$unemployed_first_period, vet.seq.10$finish_zone_1_2_unemp)
tblD %>%
  knitr::kable(caption = "Two Way Frequency Table") %>%
  kable_styling("striped", full_width = FALSE)
```



```{r echo = FALSE}
chisq.test(tblD)
```

The results of the chi-squared test (above) show a strong statistical association (p < 0.01) between transitional unemployment and finishing the sequence either unemployed or in a lower job zone.



#### **Hypothesis: Promotion in the first period is associated with holding a job in the highest job zones at the end of a career.**

To test our final hypothesis, we group veterans by those who were promoted in the first period (promoted after two years) and those who held a job in job zone four or five at the end of the 10-year period. If this association exists, it could be evidence that an early promotion results in a higher career "ceiling"; that is, higher potential outcomes for "winners" of the first round of the career tournament.

```{r echo = FALSE}
# promoted first period, zone 4 or 5 at end of 10 years
tblE = table(vet.seq.10$promoted_first_period, vet.seq.10$finish_zone_4_5)
tblE %>%
  knitr::kable(caption = "Two Way Frequency Table") %>%
  kable_styling("striped", full_width = FALSE)
```

```{r echo = FALSE}
chisq.test(tblE)
```

The results of the chi-squared test (above) show a weak statistical association (p < 0.1, but not < 0.05) between early promotion and a high final job zone. This may again be because our sample of veterans in the DC MSA tends to start out highly trained and educated, so there is not much additional mobility possible once a promotion occurs.

#### **Policy Implications**

In order to best improve veteran career mobility and outcomes, the first two years after the military exit are crucial. While early promotion does not have as strong of an association with positive outcomes, early unemployment has very strong associations with negative outcomes, such as additional unemployment or a lower career ceiling. Policy efforts that target "new" veterans should consider the strong effects of transitional unemployment on later career options and mitigate this unemployment where possible.