---
title: "Fairfax profiling"
author: "Morgan Klutzke"
date: "6/18/2020"
output: html_document
---

Profiling fairfax data, specifically the variables WATER, SEWER, GAS, VALUE_LAND, VALUE_BUILDING, & VALUE_TOTAL.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE, message=FALSE, echo=FALSE}
library(DataExplorer)
library(ggplot2)
library(scales)
library(dplyr)
library(purrr)
```

```{r import data}
library(readr)
fairfax_housing_2018_geo <- read_csv("~/git/dspg20broadbandERS/data/fairfax-data/fairfax_housing_2018_geo.csv")
```

Checking for completeness, using functions from Devika.

```{r functions}
check_calc <- function(vec) {
  blanks <- 0L
  true_na <- 0L
  written_na <- 0L
  len <- length(x = vec)
  for (elem in vec) {
    if (is.na(x = elem)) {
      true_na <- true_na + 1L
    } else if (elem == "na") {
      written_na <- written_na + 1L
    } else if (elem == "") {
      blanks <- blanks + 1L
    }
  }
  percent_complete <- (len - (blanks + true_na + written_na)) / len
  unique_values <- length(unique(vec))
  tibble(blanks = blanks,
         true_na = true_na,
         written_na = written_na,
         percent_complete = percent_complete,
         unique_values = unique_values)
}
check_complete <- function(df) {
  z <- deparse(substitute(df))
  map_df(.x = df, .f = check_calc) %>%
    mutate(column = colnames(df)) %>%
    mutate(set = print(z))  %>%
    select(set, column, blanks, true_na, written_na, percent_complete, unique_values)
}
```

```{r}
check_complete(select(fairfax_housing_2018_geo, 7:12)) %>% knitr::kable()
```

Trying out DataExplorer package.

```{r}
introduce(fairfax_housing_2018_geo)
plot_missing(fairfax_housing_2018_geo)
```

Plotting frequencies for categorical variables WATER, SEWER, and GAS.

```{r}
f <- ggplot(data = fairfax_housing_2018_geo)
f + geom_bar(mapping = aes(x = WATER, fill = WATER))
f + geom_bar(mapping = aes(x = SEWER, fill = SEWER))
f + geom_bar(mapping = aes(x = GAS, fill = GAS))
```

Looking at value distributions. Lowest and highest 1% are removed.

```{r}
value_total <- na.omit(fairfax_housing_2018_geo[, "VALUE_TOTAL"])
total <- sort(data.matrix(value_total))
lower <- length(total) %/% 100
upper <- length(total) - lower
total_mid <- total[lower:upper]
hist(total_mid)
```

```{r}
value_building <- na.omit(fairfax_housing_2018_geo[, "VALUE_BUILDING"])
building <- sort(data.matrix(value_building))
lower <- length(building) %/% 100
upper <- length(building) - lower
building_mid <- building[lower:upper]
hist(building_mid)
```

```{r}
value_land <- na.omit(fairfax_housing_2018_geo[, "VALUE_LAND"])
land <- sort(data.matrix(value_land))
lower <- length(land) %/% 100
upper <- length(land) - lower
land_mid <- land[lower:upper]
hist(land_mid, xlim = c(29000, 743000))
```

Quick check to see if VALUE_LAND and VALUE_BUILDING sum to VALUE_TOTAL.

```{r}
sum(land) + sum(building)
sum(total)
```

