---
title: "Partial Fairfax Profiling"
output: html_document
---

This is a partial profiling of the 2018 Fairfax housing data, including variables for:

- Bedrooms

- Bathrooms

- Living area

- GeoID

- District

- High School

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE, warning = FALSE)
library(dplyr)
library(ggplot2)
library(purrr)
library(sf)  
# install.packages('venn')
library(venn)
# install.packages('ggthemes')
library(ggthemes)
# install.packages('DataExplorer')
library(DataExplorer)
```



```{r}
# read in full fairfax housing data
d_full <- read.csv('~/git/dspg20broadbandERS/data/fairfax-data/fairfax_housing_2018_geo.csv')
```

```{r}
# select six variables in question
d_six <- select(d_full, 13:18)
```


```{r functions}
# functions to calculate completeness (from Devika)
check_calc <- function(vec) {
  blanks <- 0L
  true_na <- 0L
  written_na <- 0L
  len <- length(x = vec)
  for (elem in vec) {
    if (is.na(x = elem)) {
      true_na <- true_na + 1L
    } else if (elem == "na") {
      written_na <- written_na + 1L
    } else if (elem == "") {
      blanks <- blanks + 1L
    }
  }
  percent_complete <- (len - (blanks + true_na + written_na)) / len
  unique_values <- length(unique(vec))
  tibble(blanks = blanks,
         true_na = true_na,
         written_na = written_na,
         percent_complete = percent_complete,
         unique_values = unique_values)
}
check_complete <- function(df) {
  z <- deparse(substitute(df))
  map_df(.x = df, .f = check_calc) %>%
    mutate(column = colnames(df)) %>%
    mutate(set = print(z))  %>%
    select(set, column, blanks, true_na, written_na, percent_complete, unique_values)
}
```


We use a custom function to check for completion of the six variables in question. The results are below:

```{r}
# check completeness
check_complete(d_six) %>% knitr::kable()
```


We confirm these results for the full data:

```{r}
#check completeness of full data (they match)
check_complete(d_full) %>% knitr::kable()
```


We confirm these results for the six variables in question using the DataExplorer package:

```{r}
introduce(d_six)
```

We plot the missing values of the six variables in question to see proportions:

```{r}
plot_missing(d_six)
```


```{r}
d_six <- update_columns(d_six, "GEOID", as.factor)
```

We plot the continuous variables:

```{r}
plot_histogram(d_six)
```

We plot the categorical variables:

```{r}
plot_bar(d_six)
```

