---
title: "New_Kent_Profile"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

CoreLogic variables not included here- APN, census tract, fips code. Instead there is a GPIN, map, tax map, subdivision and zip code. I also assumed consideration is similar to assd value.
```{r}
library(readxl)
library(tidygeocoder)
library(dplyr)
library(stringr)
library(dplyr)
library(ggplot2)
library(purrr)
library(sf)  
library(venn)
library(ggthemes)
library(DataExplorer)
library(maditr)
library(gridExtra)

new_kent <- read_excel("~/Data.xlsx")

nk_sub <- new_kent %>%
  select(Lot, City, GPIN, Unit, Subdivision, Consideration)
```

```{r functions, include=FALSE}

check_calc <- function(vec) {
  blanks <- 0L
  true_na <- 0L
  written_na <- 0L
  len <- length(x = vec)
  for (elem in vec) {
    if (is.na(x = elem)) {
      true_na <- true_na + 1L
    } else if (elem == "na") {
      written_na <- written_na + 1L
    } else if (elem == "") {
      blanks <- blanks + 1L
    }
  }
  percent_complete <- (len - (blanks + true_na + written_na)) / len
  unique_values <- length(unique(vec))
  tibble(blanks = blanks,
         true_na = true_na,
         written_na = written_na,
         percent_complete = percent_complete,
         unique_values = unique_values)
}
check_complete <- function(df) {
  z <- deparse(substitute(df))
  map_df(.x = df, .f = check_calc) %>%
    mutate(column = colnames(df)) %>%
    mutate(set = print(z))  %>%
    select(set, column, blanks, true_na, written_na, percent_complete, unique_values)
}
```


**completeness of variables**
1) Unit is only 22% complete, consideration (which is similar to value) is 86% complete 
2) remaining variable are between 99% and 100% complete. Lot is 100% complete.

```{r}
#completeness of variables
check_complete(nk_sub) %>% knitr::kable()
```

Basic data intro
```{r}
plot_intro(nk_sub)
```

```{r}
lot <- nk_sub[, c("Consideration", "Subdivision")]
plot_scatterplot(lot, by="Consideration") 
```

consideration historgram
```{r}
grid.arrange(ggplot(nk_sub, aes(x = `Consideration`)) + geom_histogram() + scale_x_log10())
```

unit distribution
```{r}
grid.arrange(ggplot(nk_sub, aes(x = `Unit`)) + geom_bar())
```


