---
title: "CoreLogic ffx"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE, message=FALSE, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE, warning = FALSE)
# install.packages("dplyr")
library(dplyr)
library(ggplot2)
library(purrr)
library(sf)  
library(venn)
library(ggthemes)
library(DataExplorer)
library(maditr)
```

```{r include=FALSE}
# read in the CoreLogic data
```

```{r include=FALSE}
ffx_1 <- ffx[1:8]

ffx_1 <- dt_mutate(ffx_1, assd_total_value = as.numeric(assd_total_value))

```


```{r functions, include=FALSE}

check_calc <- function(vec) {
  blanks <- 0L
  true_na <- 0L
  written_na <- 0L
  len <- length(x = vec)
  for (elem in vec) {
    if (is.na(x = elem)) {
      true_na <- true_na + 1L
    } else if (elem == "na") {
      written_na <- written_na + 1L
    } else if (elem == "") {
      blanks <- blanks + 1L
    }
  }
  percent_complete <- (len - (blanks + true_na + written_na)) / len
  unique_values <- length(unique(vec))
  tibble(blanks = blanks,
         true_na = true_na,
         written_na = written_na,
         percent_complete = percent_complete,
         unique_values = unique_values)
}
check_complete <- function(df) {
  z <- deparse(substitute(df))
  map_df(.x = df, .f = check_calc) %>%
    mutate(column = colnames(df)) %>%
    mutate(set = print(z))  %>%
    select(set, column, blanks, true_na, written_na, percent_complete, unique_values)
}
```


**completeness of variables**
1) fips_code, unformatted_apn, formatted_apn are complete.
2) census_tract, lot_number, property_indicator, ass_total_value are between 95.5% and 99.3% complete.
3) township is fully incomplete; 0% complete.

```{r}
#completeness of variables
check_complete(ffx_1) %>% knitr::kable()
```

```{r}
plot_intro(ffx_1)
```

```{r}
plot_missing(ffx_1)
```

**Distribution of property_indicators/ property types**
Most properties are single family residences (10) with the next most popular property type being condominiums (11) and vacant (80). There are no miscellaneous properties (00).

```{r}
prop_indicator_dist_plot <- ggplot(ffx_1, aes(x=property_indicator, fill=property_indicator)) +
  geom_bar(position="stack") 
prop_indicator_dist_plot
```

**Distribution of assd_total_value for the different property types**
Single family homes have the greatest range in assessed total value, followed by condos and vacant properties (among the three most common property types).
```{r}
dist <- (ffx_1) %>%
  filter(!is.na(assd_total_value)) %>% 
  filter(property_indicator == c("10", "11", "80")) %>%
  group_by(property_indicator) %>%
  ggplot(aes(x= assd_total_value, y=property_indicator)) + geom_point() 

dist 
```

Average assessed total value for the three most common property types.
```{r}
dist <- (ffx_1) %>%
  filter(!is.na(assd_total_value)) %>% 
  filter(property_indicator == c("10", "11", "80")) %>%
  group_by(property_indicator) %>%
  summarise(mean = mean(assd_total_value)) %>%
  ggplot(aes(x= mean, y=property_indicator)) + geom_point() 

dist 
```

Assd and Mkt values are the same throughout
```{r include= FALSE}
ffx_1 %>%
  filter(!is.na(assd_total_value))%>%
  filter(!is.na(mkt_total_value)) %>%
  mutate(diff = assd_total_value - mkt_total_value) 
ffx_1 <- ffx_1[!grepl("0", ffx_1$diff)]
```


