---
title: "edu_knowledge_rothwell"
author: "Sarah McDonald"
date: "7/20/2020"
#output: html_document
output: md_document
---

```{r, include =F}
library(readxl)
library(dplyr)
library(data.table)
```



The construction of the crosswalk follows Rothwell’s (2015)  definition of the STW.  
Data sources used to construct the crosswalk are:
•	2010 O*NET-SOC occupations to 2010 SOC occupations :   https://www.onetcenter.org/taxonomy/2010/soc/2010_to_SOC_Crosswalk.xls?fmt=xls
•	O\*NET 15.1 Database : Content Model Knowledge data associated with each O*NET-SOC occupation: https://www.onetcenter.org/dl_files/db_15_1.zip
•	O\*NET 15.1 Database: Education percent frequency data associated with each O*NET-SOC occupation: https://www.onetcenter.org/dl_files/db_15_1.zip
•	2010 Census Occupation Codes with Crosswalk : https://www2.census.gov/programs-surveys/demo/guidance/industry-occupation/2010-occ-codes-with-crosswalk-from-2002-2011.xls


Following Rothwell’s paper, the following steps were taken to determine which occupation codes to designate as skilled technical workforce (STW). His analysis, as outlined in the paper, utilizes the O*NET Knowledge survey, which asks workers to rate the level of knowledge needed to perform their job across 33 distinct knowledge domains on a scale from 1¬ to 7. The STW crosswalk was constructed by completing the following steps:
1.	The O*NET Version 15.1 (released February 2011) Knowledge and Education survey data were downloaded. 



Knowledge and Education/Training/Experience Tables Download
```{r, warning=F, message=F}
#download.file(url = "https://www.onetcenter.org/dl_files/db_15_1.zip", destfile = "/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_15_1.zip")

#system(command = "unzip -c /sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_15_1.zip 'db_15_1/Knowledge.txt' > /sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/Knowledge.txt" )

#system(command = "unzip -c /sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_15_1.zip 'db_15_1/Education, Training, and Experience.txt' > /sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/Education.txt" )

```

OCC and SOC Crosswalk Download
```{r}
#download.file(url = "https://www2.census.gov/programs-surveys/demo/guidance/industry-occupation/2010-occ-codes-with-crosswalk-from-2002-2011.xls",
 #             destfile = "/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/2010-occ-codes-with-crosswalk-from-2002-2011.xls")

#download.file(url = "https://www.onetcenter.org/taxonomy/2010/soc/2010_to_SOC_Crosswalk.xls?fmt=xls", 
      #        destfile = "/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/2010_to_SOC_Crosswalk.xls")
```


```{r}
knowledge <- fread("/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/Knowledge.txt")
xwalk_soc <- read_xls("/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/2010_to_SOC_Crosswalk.xls", skip =3)

knowledge <- knowledge[knowledge$`Scale ID` == "LV", ]


knowledge <- merge(knowledge[, c("O*NET-SOC Code", "Element Name", "Data Value")], xwalk_soc[, c("O*NET-SOC 2010 Code", "2010 SOC Code")],
      by.x = "O*NET-SOC Code", by.y = "O*NET-SOC 2010 Code", all =  T)

```

Potentially STW competency categories are listed in the variable _element\_stw_.

The mean knowledge score is computed for each SOC and competency category (Step 2, Step 4). O\*NET Categories with at least a 4.5 and competency categories defined in _element\_stw_ will be further considered for STW.

```{r}

element_stw <- c(
"Biology",                    "Building and Construction",
"Chemistry",                  "Computers and Electronics", 
 "Design",                     "Economics and Accounting",  
"Engineering and Technology", "Food Production",           
 "Mathematics",                "Mechanical",                
"Medicine and Dentistry",     "Physics",                   
"Production and Processing",  "Telecommunications")

knowledge_soc <- knowledge %>% 
  group_by(`2010 SOC Code`, `Element Name`) %>% # group by SOC, competency category
  summarize(`Mean Data Value` = mean(`Data Value`)) %>% #take mean for each SOC, competency category grouping
  mutate(`Category STW Indicator` = ifelse(`Element Name` %in% element_stw & `Mean Data Value`>= 4.5, 1, 
                                           ifelse(is.na(`Element Name`)==T|is.na(`Mean Data Value`)==TRUE, NA, 0))) #%>% # if STW competency category and mean value is greater than 4.5, that specific SOC, competency category combination is considered STW for knowledge

knowledge_soc <- knowledge_soc %>%
  select(!`Element Name` & !`Mean Data Value`) %>%
  group_by(`2010 SOC Code`) %>%
  summarize(knowledge_STW = ifelse(length(`2010 SOC Code`)==1 & is.na(`Category STW Indicator`)==T, NA, 
                                   ifelse(length(`2010 SOC Code`)>1 & sum(`Category STW Indicator`, na.rm = T) > 0,1, 0))) %>% summarize(knowledge_STW = unique(knowledge_STW))
# to pass the knowledge STW definition, only one of the SOC, copetency category combinations must have a 1

head(knowledge_soc,10)
```

We then download the Education data and calculate the percentage of respondents without a bachelor's degree (Step 3). We then take the mean percentage of respondents with less than a bachelor’s degree for each SOC code (Step 4).
```{r}
education <- fread("/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/Education.txt")
education <- education[education$`Element Name` == "Required Level of Education", c("O*NET-SOC Code", "Category", "Data Value")]

education <- merge(education, xwalk_soc[, c("O*NET-SOC 2010 Code", "2010 SOC Code")],
      by.x = "O*NET-SOC Code", by.y = "O*NET-SOC 2010 Code", all = T)

education <- education  %>%
  mutate(hasbach = ifelse(Category <= 5, "nobach", ifelse(Category > 5, "bach", NA))) %>% # if category is greater than 5, the percentage cooresponds to bachelor's degree or higher.
  group_by(`O*NET-SOC Code`,  `2010 SOC Code`, hasbach) %>%
  summarize(`Percent No Bachelors` = sum(`Data Value`)) %>%
  filter(hasbach == "nobach"|is.na(hasbach)==TRUE) %>%
  select(!hasbach)
  
# calculate the percentage of respondents for each O*NET without a bachelor's degree

education_soc <- education %>%
  group_by(`2010 SOC Code`) %>%
  summarize(`Mean Percent No Bachelors` = mean(`Percent No Bachelors`, na.rm = T), 
            education_STW = ifelse(`Mean Percent No Bachelors`> 50, 1, ifelse(is.na(`Mean Percent No Bachelors`)==T, NA, 0)))%>%
  select(!`Mean Percent No Bachelors`)# take the mean percentage of respondents with less than a bachelor’s degree for each SOC code

head(education_soc, 10)

```

Finally, we merge our knowledge and education findings together. If there is a "1" in both the knowledge and education tables, then a SOC code is categorized as STW. 
```{r}
rothwell_soc <- merge(education_soc, knowledge_soc, by = "2010 SOC Code", all =  T)

rothwell_soc$rothwell_STW <- ifelse(rothwell_soc$knowledge_STW == 1 & rothwell_soc$education_STW ==1, 1, ifelse(is.na(rothwell_soc$knowledge_STW) == T & is.na(rothwell_soc$education_STW) ==T, NA, 0))

#We fix the case where registered nurses fail to meet the knowledge and education criteria (Step 5).
rothwell_soc[rothwell_soc$`2010 SOC Code` == "29-1141", "rothwell_STW"] <- 1

head(rothwell_soc, 10)
```

Merge with OCC
```{r}
rothwell <- merge(xwalk_soc, rothwell_soc, by = "2010 SOC Code")
occ <- read_xls("/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/2010-occ-codes-with-crosswalk-from-2002-2011.xls",
                skip = 4)
occ <- occ[,-1]

occ <- occ %>% filter(grepl("^\\d{4}$", x = `2010 Census Code`))

# some of the SOC codes in the OCC data table have "Xs" where multiple SOC codes is relevant to one OCC codes.
# ex) if 13-1111 and 13-1112 corresponded to occ code, 3783, the SOC codes would show as 13-111X

occ_XX <- occ %>% filter(grepl(pattern = "^\\d{2}\\-\\d{2}XX$", x = `2010 SOC Code`)) %>%
  mutate(`2010 SOC Code X` = substr(`2010 SOC Code`, start = 1, stop = 5))

occ_X <- occ %>% filter(grepl(pattern = "^\\d{2}\\-\\d{3}X$", x = `2010 SOC Code`)) %>%
  mutate(`2010 SOC Code X` = substr(`2010 SOC Code`, start = 1, stop = 6))

occ_not_X <- occ %>% filter(grepl(pattern = "^\\d{2}\\-\\d{4}$", x = `2010 SOC Code`)) 

rothwell <- rothwell %>%
  merge(occ, by= "2010 SOC Code", all.x = T)

  
for(j in occ_X$`2010 SOC Code X`){ 
    
  for(i in rothwell[is.na(rothwell$`2010 Census Code`) == T, "2010 SOC Code"]){

    if( i %like% j){
      
      rothwell[rothwell$`2010 SOC Code` == i, "2010 Census Code"] <- occ_X[occ_X$`2010 SOC Code X` == j, "2010 Census Code"]
      rothwell[rothwell$`2010 SOC Code` == i, "Occupation 2010 Description"] <- occ_X[occ_X$`2010 SOC Code X` == j, "Occupation 2010 Description"]

      }
  
    }


  }

for(j in occ_XX$`2010 SOC Code X`){ 
    
  for(i in rothwell[is.na(rothwell$`2010 Census Code`) == T, "2010 SOC Code"]){

    if( i %like% j){
      
      rothwell[rothwell$`2010 SOC Code` == i, "2010 Census Code"] <- occ_XX[occ_XX$`2010 SOC Code X` == j, "2010 Census Code"]
      rothwell[rothwell$`2010 SOC Code` == i, "Occupation 2010 Description"] <- occ_XX[occ_XX$`2010 SOC Code X` == j, "Occupation 2010 Description"]

      }
  
    }


  }

```

```{r}
#write.csv(rothwell, "/sfs/qumulo/qhome/sm9dv/dspg20STW/src/edu_knowledge_rothwell/rothwell.csv", row.names = F)
```
