---
title: "Add 2018 Data to Crosswalk"
author: "Sarah McDonald"
date: "10/30/2020"
output: html_document
---

First, we load the existing crosswalk, called _rothwell_, and the 2010 to 2018 crosswalk, called _xwalk_.
```{r, message = F, warning = F}
library(readxl)
library(dplyr)
library(data.table)
```

```{r, warning=F, message=F, eval = F}

# Knowledge and Ed tabel Download 25.1
download.file(url = "https://www.onetcenter.org/dl_files/database/db_25_1_text.zip", 
              destfile = "/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_25_1_text.zip")

system(command = "unzip -c /sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_25_1_text.zip 'db_25_1_text/Knowledge.txt' > /sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_25_1_knowledge.txt")

system(command = "unzip -c /sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_25_1_text.zip 'db_25_1_text/Education, Training, and Experience.txt' > /sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_25_1_education.txt")

# Knowledge and Ed table Download 25.0
download.file(url = "https://www.onetcenter.org/dl_files/database/db_25_0_text.zip", 
              destfile = "/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_25_0_text.zip")

system(command = "unzip -c /sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_25_0_text.zip 'db_25_0_text/Knowledge.txt' > /sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_25_0_knowledge.txt")

system(command = "unzip -c /sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_25_0_text.zip 'db_25_0_text/Education, Training, and Experience.txt' > /sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_25_0_education.txt")

# Knowledge and Ed table Download 15.1
download.file(url = "https://www.onetcenter.org/dl_files/db_15_1.zip", 
              destfile = "/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_15_1.zip")

system(command = "unzip -c /sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_15_1.zip 'db_15_1/Knowledge.txt' > /sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_15_1_knowledge.txt")

system(command = "unzip -c /sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_15_1.zip 'db_15_1/Education, Training, and Experience.txt' > /sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/db_15_1_education.txt")
  
  
# 2010 SOC to OCC Code Download
download.file(url = "https://www2.census.gov/programs-surveys/demo/guidance/industry-occupation/2010-occ-codes-with-crosswalk-from-2002-2011.xls",
             destfile = "/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/2010-occ-codes-with-crosswalk-from-2002-2011.xls")

# 2010 O*NET-SOC to 2010 SOC
download.file(url = "https://www.onetcenter.org/taxonomy/2010/soc/2010_to_SOC_Crosswalk.xls?fmt=xls", 
              destfile = "/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/2010_to_SOC_Crosswalk.xls")

# 2010 O*NET SOC to 2018 SOC
download.file(url = "https://www.onetcenter.org/taxonomy/2010/soc2018/2010_to_2018_SOC_Crosswalk.xls?fmt=xls",
              destfile = "/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/2010_to_2018_SOC_Crosswalk.xls")

# 2010 O*NET SOC to 2019 O*NET SOC
download.file(url = "https://www.onetcenter.org/taxonomy/2019/walk/2010_to_2019_Crosswalk.csv?fmt=csv",
              destfile = "/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/2010_to_2019_Crosswalk.csv")

# 2018 SOC to 2019 O*NET SOC
download.file(url = "https://www.onetcenter.org/taxonomy/2019/soc/2019_to_SOC_Crosswalk.csv?fmt=csv", 
              destfile = "/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/2019_to_SOC_Crosswalk.csv")
```

The 2018 Census STEM, STEM-Related and Non-STEM related Code List provides its designations with highlight colors, so we download the file in excel, and make two columns to designate STEM and STEM-Related Census codes. We upload the edited spreadsheet.



Read in tables needed

```{r}
# Knowledge
knowledge_25_1 <- fread("~/dspg20STW/data/ncses_stw/original/onet/db_25_1_knowledge.txt")
knowledge_25_0 <- fread("~/dspg20STW/data/ncses_stw/original/onet/db_25_0_knowledge.txt")
knowledge_15_1 <- fread("~/dspg20STW/data/ncses_stw/original/onet/db_15_1_knowledge.txt")

# Education
education_25_1 <- fread("~/dspg20STW/data/ncses_stw/original/onet/db_25_1_education.txt")
education_25_0 <- fread("~/dspg20STW/data/ncses_stw/original/onet/db_25_0_education.txt")
education_15_1 <- fread("~/dspg20STW/data/ncses_stw/original/onet/db_15_1_education.txt")

# O*NET-SOC 2010 to SOC 2010
onet_2010_to_soc_2010 <- read_xls("~/dspg20STW/data/ncses_stw/original/onet/2010_to_SOC_Crosswalk.xls", skip =3)

# 2010 OCC to 2010 SOC
occ_2010_to_soc_2010 <- read_xls("/sfs/qumulo/qhome/sm9dv/dspg20STW/data/ncses_stw/original/onet/2010-occ-codes-with-crosswalk-from-2002-2011.xls",
                 range = "2010OccCodeList!B5:D638",
                 .name_repair = ~ vctrs::vec_as_names(..., quiet = TRUE)) %>%
  filter(grepl("^\\d{4}$", x = `2010 Census Code`)) %>%
  rename("2010 Census Title" = "Occupation 2010 Description")

# 2018  OCC to 2018 SOC
occ_2018_to_soc_2018 <- read_xlsx("~/dspg20STW/data/ncses_stw/2018-occupation-code-list-and-crosswalk.xlsx", 
                                  range= "2018 Census Occ Code List!B5:D667",
                    .name_repair = ~ vctrs::vec_as_names(..., quiet = TRUE)) %>% 
  filter(grepl("^\\d{4}$", x = `2018 Census Code`))

# O*NET-SOC 2010 to SOC 2018 crosswalk
onet_2010_to_soc_2018 <- read_xls("~/dspg20STW/data/ncses_stw/original/onet/2010_to_2018_SOC_Crosswalk.xls", skip = 2)

# STEM Category 2018

stem_2018 <- read_xlsx("~/dspg20STW/data/ncses_stw/original/onet/edited-2018-census-stem-related-and-non-stem-occupation-code-list.xlsx", skip = 7, .name_repair = ~ vctrs::vec_as_names(..., quiet = TRUE)) %>%
  select(2:6) %>% 
  filter(grepl("^\\d{4}$", x = `2018 Census Code`)) %>% 
  mutate(`2018 STEM Category` = ifelse(`2018 STEM` == 1, "STEM", 
                                       ifelse(`2018 STEM-related` ==1, "STEM-related", "Non-STEM"))) %>%
  select(-`2018 STEM`, -`2018 STEM-related`)

# O*NET-SOC 2010 to O*NET-SOC 2019
onet_2010_onet_2019 <- read.csv("~/dspg20STW/data/ncses_stw/original/onet/2010_to_2019_Crosswalk.csv")

# SOC 2018 to O*NET-SOC 2019
soc_2018_onet_2019 <- read.csv("~/dspg20STW/data/ncses_stw/original/onet/2019_to_SOC_Crosswalk.csv")

soc_2010_soc_2018 <-read_xlsx("~/dspg20STW/data/ncses_stw/original/onet/soc_2010_to_2018_crosswalk.xlsx", skip = 8)


#rothwell <- merge(onet_2010_to_soc_2010, onet_2010_onet_2019, by.x = "O*NET-SOC 2010 Code", by.y = "O.NET.SOC.2010.Code", all = T )

#rothwell <- merge(rothwell, soc_2018_onet_2019[, c("O.NET.SOC.2019.Code", "X2018.SOC.Code", "X2018.SOC.Title")], by = "O.NET.SOC.2019.Code", all = T)

#colnames(rothwell)[colnames(rothwell) == "X2018.SOC.Code"] <- "2018 SOC Code"
#colnames(rothwell)[colnames(rothwell) == "X2018.SOC.Title"] <- "2018 SOC Title"

rothwell <- merge(onet_2010_to_soc_2010, onet_2010_to_soc_2018[, c("O*NET-SOC 2010 Code", "2018 SOC Code", "2018 SOC Title")], by = "O*NET-SOC 2010 Code", all = T)

rothwell <- merge(rothwell, onet_2010_onet_2019[, c("O.NET.SOC.2019.Code", "O.NET.SOC.2010.Code", "O.NET.SOC.2019.Title")],by.x = "O*NET-SOC 2010 Code", by.y = "O.NET.SOC.2010.Code", all = T)

rothwell <- rothwell %>% select(1,3,5, 7)

rothwell2 <- merge(onet_2010_onet_2019, soc_2018_onet_2019, by = "O.NET.SOC.2019.Code", all =T)
rothwell2<-rothwell2 %>% select(1,2,6)

# i think this is the best way
rothwell2 <- merge(rothwell2, onet_2010_to_soc_2010[, c("O*NET-SOC 2010 Code","2010 SOC Code")],by.x = "O.NET.SOC.2010.Code", by.y = "O*NET-SOC 2010 Code", all =  T)

rothwell2 <- rothwell2 %>% select(1,4,3,2)

```


Classify by 2010 O*NET

Merge with 2010 Census Occupation Table

The 2010 Census Occupation to 2010 SOC does not perfectly merge with the rothwell data frame. This is because the Census Occupation crosswalk includes broad group SOC codes (ie xx-xxx0), minor group SOC codes (xx-x000), and grouped SOC codes, where groups are represented with "Xs" (ie xx-xxxX, xx-xxXX, xx-xXXX).

We subset these groups and merge in the following order, matching to the most specific SOC code matches to least specific:

1. xx-xxxx (perfect matches)

2. xx-xxx0 (broad groups)

3. xx-xxxX

4. xx-xxXX

5. xx-x000 (minor groups)


```{r}
xx <- occ_2010_to_soc_2010 %>% filter(grepl(pattern = "^\\d{2}\\-\\d{2}XX$", x = `2010 SOC Code`)) %>%
  mutate(`2010 SOC Code XX` = substr(`2010 SOC Code`, start = 1, stop = 5))

x <- occ_2010_to_soc_2010 %>% filter(grepl(pattern = "^\\d{2}\\-\\d{3}X$", x = `2010 SOC Code`)) %>%
  mutate(`2010 SOC Code X` = substr(`2010 SOC Code`, start = 1, stop = 6))

# this just checks that we account for all cases
nrow(occ_2010_to_soc_2010) - nrow(x) - nrow(xx) - length(occ_2010_to_soc_2010$`2010 SOC Code`[!(occ_2010_to_soc_2010$`2010 SOC Code` %in% union(x$`2010 SOC Code`, xx$`2010 SOC Code`))])

broad <- occ_2010_to_soc_2010 %>% 
  filter(grepl(pattern = "^\\d{2}-\\d{2}[1-9]0$", x = `2010 SOC Code`) ) %>%
  mutate(`broad` = substr(`2010 SOC Code`, start = 1, stop = 6)) 

minor <- occ_2010_to_soc_2010 %>% 
  filter(grepl(pattern = "^\\d{2}-\\d{1}0{3}$", x = `2010 SOC Code`) ) %>%
  mutate(`minor` = substr(`2010 SOC Code`, start = 1, stop = 4)) 

# make sure that broad and x do not overwrite each other
intersect(broad$broad, x$`2010 SOC Code X`)

```

This is the function we use to fill in the table:

```{r}
fill_table <- function(data, column, year){
  
    for(j in data[[column]]){ 
        
      for(i in rothwell[is.na(rothwell[[paste(year, "Census Code", sep = " ")]]) == T & is.na(rothwell[[paste(year, "SOC Code", sep = " ")]]) == F, paste(year, "SOC Code",sep = " ")]){
    
        if(i %like% j){
          
          rothwell[rothwell[[paste(year, "SOC Code", sep = " ")]] == i & is.na(rothwell[[paste(year, "SOC Code", sep = " ")]]) == F, paste(year, "Census Code", sep = " ")] <- 
            data[data[[column]] == j, paste(year, "Census Code",sep = " ")]
          
          rothwell[rothwell[[paste(year, "SOC Code", sep = " ")]] == i & is.na(rothwell[[paste(year, "SOC Code", sep = " ")]]) == F, paste(year, "Census Title", sep = " ")] <- 
            data[data[[column]] == j, paste(year, "Census Title",sep = " ")]
    
        }
    
      }
  
    }
  
  assign("rothwell", rothwell, envir = .GlobalEnv)
  
}
```


```{r}
rothwell <- merge(rothwell, occ_2010_to_soc_2010, by = "2010 SOC Code", all.x = T)

fill_table(data = x, column = "2010 SOC Code X", year = 2010)
fill_table(data = broad, column = "broad", year = 2010)
fill_table(data = xx, column = "2010 SOC Code XX", year = 2010)
fill_table(data = minor, column = "minor", year = 2010)
```

Now, merge with 2018 SOC and OCC codes


some of the SOC codes in the OCC data table have "Xs" where multiple SOC codes is relevant to one OCC codes.
ex) if 13-1111 and 13-1112 corresponded to occ code, 3783, the SOC codes would show as 13-111X

```{r}
xxx <- occ_2018_to_soc_2018 %>% filter(grepl(pattern = "^\\d{2}\\-\\d{1}XXX$", x = `2018 SOC Code`)) %>%
  mutate(`2018 SOC Code XXX` = substr(`2018 SOC Code`, start = 1, stop = 4))

xx <- occ_2018_to_soc_2018 %>% filter(grepl(pattern = "^\\d{2}\\-\\d{2}XX$", x = `2018 SOC Code`)) %>%
  mutate(`2018 SOC Code XX` = substr(`2018 SOC Code`, start = 1, stop = 5))

x <- occ_2018_to_soc_2018 %>% filter(grepl(pattern = "^\\d{2}\\-\\d{3}X$", x = `2018 SOC Code`)) %>%
  mutate(`2018 SOC Code X` = substr(`2018 SOC Code`, start = 1, stop = 6))

occ_2018_to_soc_2018_not_X <- occ_2018_to_soc_2018 %>% filter(grepl(pattern = "^\\d{2}\\-\\d{4}$", x = `2018 SOC Code`)) 


# this just checks that we account for all cases
nrow(occ_2018_to_soc_2018) - nrow(x) - nrow(xx) - nrow(xxx) - length(occ_2018_to_soc_2018$`2018 SOC Code`[!(occ_2018_to_soc_2018$`2018 SOC Code` %in% union(x$`2018 SOC Code`, union(xx$`2018 SOC Code`, xxx$`2018 SOC Code`)))])


broad <- occ_2018_to_soc_2018 %>% 
  filter(grepl(pattern = "^\\d{2}-\\d{2}[1-9]0$", x = `2018 SOC Code`) ) %>%
  mutate(`broad` = substr(`2018 SOC Code`, start = 1, stop = 6)) 

minor <- occ_2018_to_soc_2018 %>% 
  filter(grepl(pattern = "^\\d{2}-\\d{1}0{3}$", x = `2018 SOC Code`) ) %>%
  mutate(`minor` = substr(`2018 SOC Code`, start = 1, stop = 4)) 

# check that nothing gets overwritten
intersect(broad$broad, x$`2018 SOC Code X`)
intersect(minor$minor, xxx$`2018 SOC Code XXX`)
```


```{r}
rothwell <- merge(rothwell, occ_2018_to_soc_2018, by = "2018 SOC Code", all.x = T)

fill_table(data = x, column = "2018 SOC Code X", year = 2018)
fill_table(data = broad, column = "broad", year = 2018)
fill_table(data = xx, column = "2018 SOC Code XX", year = 2018)
fill_table(data = minor, column = "minor", year = 2018)
fill_table(data = xxx, column = "2018 SOC Code XXX", year = 2018)
```

Merge Census STEM data

```{r}
rothwell <- merge(rothwell, stem_2018[, c("2018 SOC Code", "2018 Census Code", "2018 STEM Category")], by = c("2018 SOC Code", "2018 Census Code"), all.x = T)
```


```{r}
xxx <- stem_2018 %>% filter(grepl(pattern = "^\\d{2}\\-\\d{1}XXX$", x = `2018 SOC Code`)) %>%
  mutate(`2018 SOC Code XXX` = substr(`2018 SOC Code`, start = 1, stop = 4))

xx <- stem_2018 %>% filter(grepl(pattern = "^\\d{2}\\-\\d{2}XX$", x = `2018 SOC Code`)) %>%
  mutate(`2018 SOC Code XX` = substr(`2018 SOC Code`, start = 1, stop = 5))

x <- stem_2018 %>% filter(grepl(pattern = "^\\d{2}\\-\\d{3}X$", x = `2018 SOC Code`)) %>%
  mutate(`2018 SOC Code X` = substr(`2018 SOC Code`, start = 1, stop = 6))

# this just checks that we account for all cases
nrow(stem_2018) - nrow(x) - nrow(xx)- nrow(xxx) - length(stem_2018$`2018 SOC Code`[!(stem_2018$`2018 SOC Code` %in% union(x$`2018 SOC Code`, union(xx$`2018 SOC Code`, xxx$`2018 SOC Code`)))])

broad <- stem_2018 %>% 
  filter(grepl(pattern = "^\\d{2}-\\d{2}[1-9]0$", x = `2018 SOC Code`) ) %>%
  mutate(`broad` = substr(`2018 SOC Code`, start = 1, stop = 6)) 

minor <- stem_2018 %>% 
  filter(grepl(pattern = "^\\d{2}-\\d{1}0{3}$", x = `2018 SOC Code`) ) %>%
  mutate(`minor` = substr(`2018 SOC Code`, start = 1, stop = 4)) 

# make sure that broad and x do not overwrite each other
intersect(broad$broad, x$`2018 SOC Code X`)
intersect(minor$minor, xxx$`2018 SOC Code XXX`)
```

This is the function we use to fill in the table:

```{r}
fill_stem_table <- function(data, column, year){
  
    for(j in data[[column]]){ 
        
      for(i in rothwell[is.na(rothwell[[paste(year, "STEM Category", sep = " ")]]) == T & is.na(rothwell[[paste(year, "SOC Code", sep = " ")]]) == F, paste(year, "SOC Code",sep = " ")]){
    
        if(i %like% j){
          
          rothwell[rothwell[[paste(year, "SOC Code", sep = " ")]] == i & is.na(rothwell[[paste(year, "SOC Code", sep = " ")]]) == F, paste(year, "STEM Category", sep = " ")] <- 
            data[data[[column]] == j, paste(year, "STEM Category",sep = " ")]
 
    
        }
    
      }
  
    }
  
  assign("rothwell", rothwell, envir = .GlobalEnv)
  
}
```


```{r}
fill_stem_table(data = x, column = "2018 SOC Code X", year = 2018)
fill_stem_table(data = broad, column = "broad", year = 2018)
fill_stem_table(data = xx, column = "2018 SOC Code XX", year = 2018)
fill_stem_table(data = minor, column = "minor", year = 2018)
fill_stem_table(data = xxx, column = "2018 SOC Code XXX", year = 2018)
```


### ROTHWELL STW
Knowledge Criteria

Following Rothwell’s paper, the following steps were taken to determine which occupation codes to designate as skilled technical workforce (STW). His analysis, as outlined in the paper, utilizes the O*NET Knowledge survey, which asks workers to rate the level of knowledge needed to perform their job across 33 distinct knowledge domains on a scale from 1 to 7.

O\*NET Database version 25.0 is the last O\*NET-SOC 2010 based version. See [https://www.onetcenter.org/taxonomy.html](https://www.onetcenter.org/taxonomy.html)

```{r}
knowledge_25_1 <- knowledge_25_1[knowledge_25_1$`Scale ID` == "LV", ]

knowledge_25_0 <- knowledge_25_0[knowledge_25_0$`Scale ID` == "LV", ]
knowledge_15_1 <- knowledge_15_1[knowledge_15_1$`Scale ID` == "LV", ]
```

```{r}
knowledge_25_1 <- merge(knowledge_25_1[, c("O*NET-SOC Code", "Element Name", "Data Value")], soc_2018_onet_2019[, c("O.NET.SOC.2019.Code", "X2018.SOC.Code")],
      by.x = "O*NET-SOC Code", by.y = "O.NET.SOC.2019.Code", all =  T)

knowledge_25_0 <- merge(knowledge_25_0[, c("O*NET-SOC Code", "Element Name", "Data Value")], onet_2010_to_soc_2010[, c("O*NET-SOC 2010 Code", "2010 SOC Code")],
      by.x = "O*NET-SOC Code", by.y = "O*NET-SOC 2010 Code", all =  T)

knowledge_15_1 <- merge(knowledge_15_1[, c("O*NET-SOC Code", "Element Name", "Data Value")], onet_2010_to_soc_2010[, c("O*NET-SOC 2010 Code", "2010 SOC Code")],
      by.x = "O*NET-SOC Code", by.y = "O*NET-SOC 2010 Code", all =  T)
```
Potentially STW competency categories are listed in the variable _element\_stw_. The mean knowledge score is computed for each SOC and competency category (Step 2, Step 4). O\*NET Categories with at least a 4.5 and competency categories defined in _element\_stw_ will be further considered for STW.

```{r}
element_stw <- c(
"Biology",                    "Building and Construction",
"Chemistry",                  "Computers and Electronics", 
 "Design",                     "Economics and Accounting",  
"Engineering and Technology", "Food Production",           
 "Mathematics",                "Mechanical",                
"Medicine and Dentistry",     "Physics",                   
"Production and Processing",  "Telecommunications")
```


```{r}
knowledge_25_1_soc <- knowledge_25_1 %>% 
  group_by(`X2018.SOC.Code`, `Element Name`) %>% # group by SOC, competency category
  summarize(`Mean Data Value` = mean(`Data Value`)) %>% #take mean for each SOC, competency category grouping
  mutate(`Category STW Indicator` = ifelse(`Element Name` %in% element_stw & `Mean Data Value`>= 4.5, 1, 
                                           ifelse(is.na(`Element Name`)==T|is.na(`Mean Data Value`)==TRUE, NA, 0))) #%>% # if STW competency category and mean value is greater than 4.5, that specific SOC, competency category combination is considered STW for knowledge

knowledge_25_0_soc <- knowledge_25_0 %>% 
  group_by(`2010 SOC Code`, `Element Name`) %>% # group by SOC, competency category
  summarize(`Mean Data Value` = mean(`Data Value`)) %>% #take mean for each SOC, competency category grouping
  mutate(`Category STW Indicator` = ifelse(`Element Name` %in% element_stw & `Mean Data Value`>= 4.5, 1, 
                                           ifelse(is.na(`Element Name`)==T|is.na(`Mean Data Value`)==TRUE, NA, 0))) #%>% # if STW competency category and mean value is greater than 4.5, that specific SOC, competency category combination is considered STW for knowledge

knowledge_15_1_soc <- knowledge_15_1 %>% 
  group_by(`2010 SOC Code`, `Element Name`) %>% # group by SOC, competency category
  summarize(`Mean Data Value` = mean(`Data Value`)) %>% #take mean for each SOC, competency category grouping
  mutate(`Category STW Indicator` = ifelse(`Element Name` %in% element_stw & `Mean Data Value`>= 4.5, 1, 
                                           ifelse(is.na(`Element Name`)==T|is.na(`Mean Data Value`)==TRUE, NA, 0))) #%>% # if STW competency category and mean value is greater than 4.5, that specific SOC, competency category combination is considered STW for knowledge


knowledge_25_1_soc <- knowledge_25_1_soc %>%
  select(!`Element Name` & !`Mean Data Value`) %>%
  group_by(`X2018.SOC.Code`) %>%
  summarize(knowledge_stw_25_1 = ifelse(length(`X2018.SOC.Code`)==1 & is.na(`Category STW Indicator`)==T, NA, 
                                   ifelse(length(`X2018.SOC.Code`)>1 & sum(`Category STW Indicator`, na.rm = T) > 0,1, 0))) %>% 
  summarize(knowledge_stw_25_1 = unique(knowledge_stw_25_1))

knowledge_25_0_soc <- knowledge_25_0_soc %>%
  select(!`Element Name` & !`Mean Data Value`) %>%
  group_by(`2010 SOC Code`) %>%
  summarize(knowledge_stw_25_0 = ifelse(length(`2010 SOC Code`)==1 & is.na(`Category STW Indicator`)==T, NA, 
                                   ifelse(length(`2010 SOC Code`)>1 & sum(`Category STW Indicator`, na.rm = T) > 0,1, 0))) %>% 
  summarize(knowledge_stw_25_0 = unique(knowledge_stw_25_0))
# to pass the knowledge STW definition, only one of the SOC, copetency category combinations must have a 1

knowledge_15_1_soc <- knowledge_15_1_soc %>%
  select(!`Element Name` & !`Mean Data Value`) %>%
  group_by(`2010 SOC Code`) %>%
  summarize(knowledge_stw_15_1 = ifelse(length(`2010 SOC Code`)==1 & is.na(`Category STW Indicator`)==T, NA, 
                                   ifelse(length(`2010 SOC Code`)>1 & sum(`Category STW Indicator`, na.rm = T) > 0,1, 0))) %>% 
  summarize(knowledge_stw_15_1 = unique(knowledge_stw_15_1))
head(knowledge_25_1_soc,10)
head(knowledge_25_0_soc,10)
head(knowledge_15_1_soc,10)
```


We then download the Education data and calculate the percentage of respondents without a bachelor's degree (Step 3). We then take the mean percentage of respondents with less than a bachelor’s degree for each SOC code (Step 4).

```{r}
education_25_1 <- fread("~/dspg20STW/data/ncses_stw/original/onet/db_25_1_education.txt")
education_25_0 <- fread("~/dspg20STW/data/ncses_stw/original/onet/db_25_0_education.txt")
education_15_1 <- fread("~/dspg20STW/data/ncses_stw/original/onet/db_15_1_education.txt")

education_25_1$Category <- as.numeric(education_25_1$Category)
education_25_0$Category <- as.numeric(education_25_0$Category)
education_15_1$Category <- as.numeric(education_15_1$Category)

education_25_1 <- education_25_1[education_25_1$`Element Name` == "Required Level of Education", c("O*NET-SOC Code", "Category", "Data Value")]
education_25_0 <- education_25_0[education_25_0$`Element Name` == "Required Level of Education", c("O*NET-SOC Code", "Category", "Data Value")]
education_15_1 <- education_15_1[education_15_1$`Element Name` == "Required Level of Education", c("O*NET-SOC Code", "Category", "Data Value")]



education_25_1 <- merge(education_25_1, soc_2018_onet_2019[, c("O.NET.SOC.2019.Code", "X2018.SOC.Code")],
      by.x = "O*NET-SOC Code", by.y = "O.NET.SOC.2019.Code", all =  T)
education_25_0 <- merge(education_25_0, onet_2010_to_soc_2010[, c("O*NET-SOC 2010 Code", "2010 SOC Code")],
      by.x = "O*NET-SOC Code", by.y = "O*NET-SOC 2010 Code", all = T)
education_15_1 <- merge(education_15_1, onet_2010_to_soc_2010[, c("O*NET-SOC 2010 Code", "2010 SOC Code")],
      by.x = "O*NET-SOC Code", by.y = "O*NET-SOC 2010 Code", all = T)

education_25_1 <- education_25_1  %>%
  mutate(hasbach = ifelse(Category <= 5, "nobach", ifelse(Category > 5, "bach", NA))) %>%
  select(!Category) %>%
  group_by(`O*NET-SOC Code`, `X2018.SOC.Code`, hasbach) %>%
  summarize(per= sum(`Data Value`)) %>%
  filter(hasbach == "nobach"|is.na(hasbach) == T) %>%
  rename("per_no_bach" = "per")

education_25_0 <-education_25_0  %>%
  mutate(hasbach = ifelse(Category <= 5, "nobach", ifelse(Category > 5, "bach", NA))) %>%
  select(!Category) %>%
  group_by(`O*NET-SOC Code`, `2010 SOC Code`, hasbach) %>%
  summarize(per= sum(`Data Value`)) %>%
  filter(hasbach == "nobach"|is.na(hasbach) == T) %>%
  rename("per_no_bach" = "per")

education_15_1 <-education_15_1  %>%
  mutate(hasbach = ifelse(Category <= 5, "nobach", ifelse(Category > 5, "bach", NA))) %>%
  select(!Category) %>%
  group_by(`O*NET-SOC Code`, `2010 SOC Code`, hasbach) %>%
  summarize(per= sum(`Data Value`)) %>%
  filter(hasbach == "nobach"|is.na(hasbach) == T) %>%
  rename("per_no_bach" = "per")
```

```{r}
# calculate the percentage of respondents for each O*NET without a bachelor's degree

education_25_1_soc <- education_25_1 %>%
  group_by(`X2018.SOC.Code`) %>%
  summarize(mean_per_no_bach = mean(per_no_bach, na.rm = T), 
            education_stw_25_1 = ifelse(mean_per_no_bach> 50, 1, ifelse(is.na(mean_per_no_bach)==T, NA, 0)))%>%
  select(!mean_per_no_bach)# take the mean percentage of respondents with less than a bachelor’s degree for each SOC code

education_25_0_soc <- education_25_0 %>%
  group_by(`2010 SOC Code`) %>%
  summarize(mean_per_no_bach = mean(per_no_bach, na.rm = T), 
            education_stw_25_0 = ifelse(mean_per_no_bach> 50, 1, ifelse(is.na(mean_per_no_bach)==T, NA, 0)))%>%
  select(!mean_per_no_bach)# take the mean percentage of respondents with less than a bachelor’s degree for each SOC code

education_15_1_soc <- education_15_1 %>%
  group_by(`2010 SOC Code`) %>%
  summarize(mean_per_no_bach = mean(per_no_bach, na.rm = T), 
            education_stw_15_1 = ifelse(mean_per_no_bach> 50, 1, ifelse(is.na(mean_per_no_bach)==T, NA, 0)))%>%
  select(!mean_per_no_bach)# take the mean percentage of respondents with less than a bachelor’s degree for each SOC code

head(education_25_1_soc, 10)
head(education_25_0_soc, 10)
head(education_15_1_soc, 10)
```


Finally, we merge our knowledge and education findings together. If there is a "1" in both the knowledge and education tables, then a SOC code is categorized as STW. 
```{r}
rothwell_soc_25_1 <- merge(education_25_1_soc, knowledge_25_1_soc,  by = "X2018.SOC.Code", all =  T)
rothwell_soc_25_0 <- merge(education_25_0_soc, knowledge_25_0_soc,  by = "2010 SOC Code", all =  T)
rothwell_soc_15_1 <- merge(education_15_1_soc, knowledge_15_1_soc,  by = "2010 SOC Code", all =  T)


rothwell_soc_25_1$rothwell_stw_25_1 <- 
           ifelse(rothwell_soc_25_1$knowledge_stw_25_1 == 1 & rothwell_soc_25_1$education_stw_25_1 ==1, 1, 
                  ifelse(is.na(rothwell_soc_25_1$knowledge_stw_25_1) == T |is.na(rothwell_soc_25_1$education_stw_25_1) ==T, NA, 0))

rothwell_soc_25_0$rothwell_stw_25_0 <- 
           ifelse(rothwell_soc_25_0$knowledge_stw_25_0 == 1 & rothwell_soc_25_0$education_stw_25_0 ==1, 1, 
                  ifelse(is.na(rothwell_soc_25_0$knowledge_stw_25_0) == T |is.na(rothwell_soc_25_0$education_stw_25_0) ==T, NA, 0))

rothwell_soc_15_1$rothwell_stw_15_1 <- 
           ifelse(rothwell_soc_15_1$knowledge_stw_15_1 == 1 & rothwell_soc_15_1$education_stw_15_1 ==1, 1, 
                  ifelse(is.na(rothwell_soc_15_1$knowledge_stw_15_1) == T |is.na(rothwell_soc_15_1$education_stw_15_1) ==T, NA, 0))

#We fix the case where registered nurses fail to meet the knowledge and education criteria (Step 5).
rothwell_soc_25_1[rothwell_soc_25_1$X2018.SOC.Code == "29-1141", "rothwell_stw_25_1"] <- 1
rothwell_soc_25_0[rothwell_soc_25_0$`2010 SOC Code` == "29-1141", "rothwell_stw_25_0"] <- 1
rothwell_soc_15_1[rothwell_soc_15_1$`2010 SOC Code` == "29-1141", "rothwell_stw_15_1"] <- 1
```

Merge Rothwell STW Definition with Table

```{r}
rothwell <- merge(rothwell, rothwell_soc_15_1, by = "2010 SOC Code", all = T)
rothwell <- merge(rothwell, rothwell_soc_25_0, by = "2010 SOC Code", all = T)
rothwell <- merge(rothwell, rothwell_soc_25_1, by.x = "2018 SOC Code", by.y = "X2018.SOC.Code", all = T)
```

```{r}
rothwell <- rothwell %>% 
  mutate(rothwell_stw_25_1 = 
           ifelse(rothwell_stw_25_1 == 1, "STW", 
                  ifelse(rothwell_stw_25_1 == 0, "Non-STW", NA)),
         education_stw_25_1 = 
           ifelse(education_stw_25_1 == 1, "Yes",
                  ifelse(education_stw_25_1 ==0, "No", NA)),
         knowledge_stw_25_1 = 
           ifelse(knowledge_stw_25_1 == 1, "Yes", 
                  ifelse(knowledge_stw_25_1 == 0, "No", NA)),
         rothwell_stw_25_0 = 
           ifelse(rothwell_stw_25_0 == 1, "STW", 
                  ifelse(rothwell_stw_25_0 == 0, "Non-STW", NA)),
         education_stw_25_0 = 
           ifelse(education_stw_25_0 == 1, "Yes",
                  ifelse(education_stw_25_0 ==0, "No", NA)),
         knowledge_stw_25_0 = 
           ifelse(knowledge_stw_25_0 == 1, "Yes", 
                  ifelse(knowledge_stw_25_0 == 0, "No", NA)),
         rothwell_stw_15_1 = 
           ifelse(rothwell_stw_15_1 == 1, "STW", 
                  ifelse(rothwell_stw_15_1 == 0, "Non-STW", NA)),
         education_stw_15_1 = 
           ifelse(education_stw_15_1 == 1, "Yes",
                  ifelse(education_stw_15_1 ==0, "No", NA)),
         knowledge_stw_15_1 = 
           ifelse(knowledge_stw_15_1 == 1, "Yes", 
                  ifelse(knowledge_stw_15_1 == 0, "No", NA)))
```


```{r}
rothwell <- rothwell %>%
  rename(
    "soc_code_2010" = "2010 SOC Code", 
    "soc_code_2018" = "2018 SOC Code", 
         "census_code_2018" = "2018 Census Code", 
         "onet_code_2010" = "O*NET-SOC 2010 Code", 
         "onet_title_2010" = "O*NET-SOC 2010 Title", 
         "soc_title_2010" = "2010 SOC Title",
         "census_title_2010" = "2010 Census Title",
         "census_code_2010" = "2010 Census Code", 
         "soc_title_2018" = "2018 SOC Title", 
         "census_title_2018" = "2018 Census Title",
         "onet_code_2019" = "O.NET.SOC.2019.Code",
         "onet_title_2019" = "O.NET.SOC.2019.Title",
         "stem_category_2018" = "2018 STEM Category") %>% 
  select(rothwell_stw_25_1, education_stw_25_1, knowledge_stw_25_1, rothwell_stw_25_0, education_stw_25_0, knowledge_stw_25_0, rothwell_stw_15_1, education_stw_15_1, knowledge_stw_15_1,
         stem_category_2018, soc_code_2018, soc_title_2018, onet_code_2019, onet_title_2019, census_code_2018, census_title_2018,
         soc_code_2010, soc_title_2010, onet_code_2010, onet_title_2010, census_code_2010, census_title_2010)
```


```{r, eval = F}
write.csv(rothwell, "~/dspg20STW/src/edu_knowledge_rothwell/rothwell_revised_test.csv", row.names = F)
```



