---
title: "Full New Kent Profiling"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(readxl)
library(DataExplorer)
library(ggplot2)
library(dplyr)
library(purrr)
library(data.table)
library(maditr)
library(gridExtra)
library(Hmisc)
```

```{r functions}
check_calc <- function(vec) {
  blanks <- 0L
  true_na <- 0L
  written_na <- 0L
  len <- length(x = vec)
  for (elem in vec) {
    if (is.na(x = elem)) {
      true_na <- true_na + 1L
    } else if (elem == "na") {
      written_na <- written_na + 1L
    } else if (elem == "") {
      blanks <- blanks + 1L
    }
  }
  percent_complete <- (len - (blanks + true_na + written_na)) / len
  unique_values <- length(unique(vec))
  tibble(blanks = blanks,
         true_na = true_na,
         written_na = written_na,
         percent_complete = percent_complete,
         unique_values = unique_values)
}
check_complete <- function(df) {
  z <- deparse(substitute(df))
  map_df(.x = df, .f = check_calc) %>%
    mutate(column = colnames(df)) %>%
    mutate(set = print(z))  %>%
    select(set, column, blanks, true_na, written_na, percent_complete, unique_values)
}
```

```{r}
new_kent <- read_excel('~/git/dspg20broadbandERS/new-kent-data/NK ASSESSMENT REPORT 2020.xlsx')
```

Looking at: Building Area (Effective), Building Area (Gross), Building Area (Living), Land Area (Acres)

```{r}
check_complete(select(new_kent,53:56))
introduce(select(new_kent, 53:56))
plot_missing(select(new_kent, 53:56))
```
```{r}
options(scipen = 10)
grid.arrange(ggplot(new_kent, aes(x = `Land Area (Acres)`)) + geom_histogram() + scale_x_log10(), 
             ggplot(new_kent, aes(x = `Building Area (Effective)`)) + geom_histogram() + scale_x_log10(), 
             ggplot(new_kent, aes(x = `Building Area (Gross)`)) + geom_histogram() + scale_x_log10(), 
             ggplot(new_kent, aes(x = `Building Area (Living)`)) + geom_histogram() + scale_x_log10(),
             ncol = 2)
```

Looking at: Land, Improvement, Total

```{r}
check_complete(select(new_kent, 19:21))
introduce(select(new_kent, 19:21))
plot_missing(select(new_kent, 19:21))
```
```{r}
grid.arrange(ggplot(new_kent, aes(x = Land)) + geom_histogram() + scale_x_log10(), 
             ggplot(new_kent, aes(x = Improvement)) + geom_histogram() + scale_x_log10(), 
             ggplot(new_kent, aes(x = Total)) + geom_histogram() + scale_x_log10(), 
             nrow = 2)
```
```{r}
new_kent <- new_kent %>%
  dt_mutate(value_sum = Land + Improvement) %>%
  dt_mutate(difference = Total - value_sum)
sum(new_kent$difference)
```

Looking at: Record Date

```{r}
check_complete(select(new_kent, 29))
introduce(select(new_kent, 29))
plot_missing(select(new_kent, 29))
```
```{r}
records <- as.POSIXlt(na.omit(new_kent$`Record Date`))
record_years <- records$year + 1900
h <- hist(record_years)
text(h$mids, h$counts, labels = h$counts, adj = c(0.5, -0.1))
```

No information about tax amount, tax year, assessed year, market value as opposed to assessed value

```{r}
nk_sub <- new_kent %>%
  select(Lot, City, GPIN, Unit, Subdivision, Consideration)
```


**completeness of variables**
1) Unit is only 22% complete, consideration (which is similar to value) is 86% complete 
2) remaining variable are between 99% and 100% complete. Lot is 100% complete.

```{r}
#completeness of variables
check_complete(nk_sub) %>% knitr::kable()
```

Basic data intro
```{r}
plot_intro(nk_sub)
```

```{r}
lot <- nk_sub[, c("Consideration", "Subdivision")]
plot_scatterplot(lot, by="Consideration") 
```

consideration historgram
```{r}
grid.arrange(ggplot(nk_sub, aes(x = `Consideration`)) + geom_histogram() + scale_x_log10())
```

unit distribution
```{r}
grid.arrange(ggplot(nk_sub, aes(x = `Unit`)) + geom_bar())
```

```{r}
#building__square_feet, living_square_feet, year_built, effective_year_built, bedrooms, total_baths, full_baths, half_baths

profile <- select(new_kent, c(36,50:52,53,55))

names(profile) <- c('year_built', 'num_beds', 'num_baths', 'num_halfbaths', 'building_area_eff', 'building_area_living')
names(profile)

sapply(profile, class)

profile$num_baths <- as.numeric(profile$num_baths)
profile$num_halfbaths <- as.numeric(profile$num_halfbaths)
profile$num_beds <- as.numeric(profile$num_beds)
profile$year_built <- as.factor(profile$year_built)
profile$numeric <- as.numeric(profile$year_built)
```


```{r}
nrow(profile)
```


```{r}
check_complete(profile) %>% knitr::kable()
```

```{r}
ggplot(data = profile) +
  aes(x = year_built) + 
  geom_histogram(stat = 'count')
```


```{r}
describe(profile$year_built)
```


```{r}
ggplot(profile) +
  aes(num_beds) +
  geom_histogram()
```

```{r}
ggplot(profile) +
  aes(num_baths) +
  geom_histogram()
```


```{r}
ggplot(profile) +
  aes(num_halfbaths) +
  geom_histogram()
```


```{r}
ggplot(profile) +
  aes(building_area_eff) +
  geom_histogram()
```


```{r}
summary(profile$building_area_eff)
describe(profile$building_area_eff)
```


```{r}
ggplot(profile) +
  aes(building_area_living) +
  geom_histogram()
```

```{r}
summary(profile$building_area_living)
describe(profile$building_area_living)
```



