---
title: "ellen_vital_stats"
output: html_document
---

```{r}
library(here)
library(dplyr)
library(ggplot2)
library(lubridate)
library(naniar)

theme_set(theme_minimal())
```

```{r}
vital_stats <- vroom::vroom(here("data", "original", "vital_stats", "Death_dataset_from_2016_to_YTD.csv"),
                            col_types = paste0(rep("c", 46), collapse = "")) %>% 
  mutate(DATE_OF_DEATH = mdy(DATE_OF_DEATH))
```

```{r}
head(vital_stats)
tail(vital_stats)
```

Any mapping work will be difficult, maybe zip codes good enough

```{r}
miss_var_summary(vital_stats)
```

why is HISPANIC_ORIGIN missing 100% of the time??

```{r}
length(unique(vital_stats$RACE))
length(unique(vital_stats$FACILITY_NAME))
length(unique(vital_stats$CAUSE_OF_DEATH))
```

There's a lot of differnt causes of death, all coded, we need to know what those codes mean to do anything real. 

```{r}
table(vital_stats$GENDER)
```

About equal death breakdown, makes sense. 

```{r}
table(vital_stats$EDUC)
```

I wonder what these differnt catagoreies mean?

```{r}
# cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#CC79A7", "#F0E442", "#0072B2", "#D55E00")
cbPalette <- c(rev(alpha(grey.colors(4, end = 0.7), 0.5)), "#d400d0")

vital_stats %>% 
  mutate(year = as.character(year(DATE_OF_DEATH)),
         yday = yday(DATE_OF_DEATH),
         month = month(DATE_OF_DEATH, label = TRUE, abbr = FALSE),
         yday_group = 15 * (yday %/% 15)) %>%
  group_by(month, year) %>% 
  summarize(n = n()) %>% 
  mutate(n = ifelse(year == 2020 & month == "June", n*30/14, n)) %>% 
  ggplot() +
  geom_point(aes(x = month, y = n, color = year)) +
  geom_line(aes(x = month,y = n, color = year, group = year)) +
  labs(x = NULL,
       y = "Monthly Deaths",
       title = "Charlottesville and Albemarle\nMonthly Deaths by Year") +
  theme(plot.title = element_text(hjust = 0.5, size = 22),
        axis.text = element_text(size = 16, 
                                 angle = 90,
                                 hjust = 1),
        axis.title = element_text(size = 20),
        strip.text = element_text(size = 18),
        legend.key = element_blank(),
        legend.text = element_text(size = 16),
        #legend.title = element_text(size = 18),
        legend.title = element_blank()) +
  scale_color_manual(values = cbPalette) +
  lims(y = c(0, 230))




ggsave(here::here("output", "mortality.png"), device = "png", width = 10, height = 6)
```

Not the perfect vis for it, but no obvious increase in deaths this year. 

```{r}
vital_stats %>% 
  mutate(AGE = as.numeric(AGE)) %>% 
  ggplot() +
  geom_histogram(aes(x = AGE), boundary = 0, binwidth = 5) +
  geom_vline(aes(xintercept = median(AGE)), color = "red") +
  xlim(0, 120)
```

Looks like either a lot of babies die while being born, they count miscarriages as death, or something odd is going on with recording people as dying at age 0. 



