---
title: "EMS data EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basic EDA

Looking for anything in the data that we need to ask questions about.

```{r, message=FALSE}
library(naniar)
library(ggplot2)
library(dplyr)
library(lubridate)
```

```{r}
ems_full <- read.csv("../../data/working/ems_full_v1.csv", colClasses = c(rep("character", 4), "POSIXct", rep("character", 78)))

#head(ems_full)
#summary(ems_full)

```

```{r}
str(ems_full)
```

```{r}
missing <- miss_var_summary(ems_full) %>%
  mutate_if(is.numeric, round, digits = 2)
missing

#ggplot(data=missing, aes(x=reorder(variable, -pct_miss), y=pct_miss)) +
#  geom_bar(stat='identity') +
#  theme(axis.text.x = element_text(angle = 60))

```

```{r}
table(ems_full$source)
table(ems_full$incident_type)  # all EMS, 8 that are not are from parsing errors

```

```{r}
table(ems_full$agency_name)

#ems_full[ems_full$agency_name == '118',]  # also giving results where agency_name is NA
#ems_full[which(ems_full$agency_name == '118'), ]
```

```{r}

```

## Counting number of incidents by year and month

**Noted issues in data - There are 19020 duplicated incident numbers with non-matching dates!** Sometimes these are very close dates, but for now I'm just keeping the first listed incident number and corresponding date.

```{r}
#table(ems_full$incident_date)

dates <- data.frame(inc_num = ems_full$response_incident_number,
                    yr = as.factor(year(ems_full$incident_date)), 
                    mo = as.factor(month(ems_full$incident_date)), 
                    dy = as.character(day(ems_full$incident_date)) )

miss_var_summary(dates)  # 106 missing indicent numbers (unique incidents or not?), 17 missing dates.  
                         # For now - leave these out                 

# find only dates for unique incidents

comp_dates <- dates[complete.cases(dates), ] # removes 116 rows with at least one NA
unique_incidents_idx <- !duplicated(comp_dates$inc_num)  
u_dates <- comp_dates[unique_incidents_idx, ]

```

```{r}
yr_mo <- u_dates %>%
            group_by(yr, mo) %>%
            count()
yr_mo  

```

```{r}

#years_count <- table(dates$yr)
#barplot(years_count)

cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#CC79A7", "#F0E442", "#0072B2", "#D55E00")

ggplot(data=u_dates) +
  geom_bar(mapping = aes(x = mo, fill = yr), position = 'dodge') +
  scale_fill_manual(values=cbPalette) +
  labs(title = "Number of Incidents by Month and Year ", 
       x = "Month", y = "Number of Incidents") 
 
  
```

```{r}
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#CC79A7", "#F0E442", "#0072B2", "#D55E00")

ggplot(data=yr_mo) +
  geom_point(mapping = aes(x = mo, y = n, colour = yr)) +
  geom_line(mapping = aes(x = as.numeric(mo), y = n, colour = yr)) +
  labs(title = "Number of Incidents by Month and Year ", 
       x = "Month", y = "Number of Incidents") +
  scale_colour_manual(values=cbPalette) + 
  theme_minimal()
  
```

## Next Step - looking at the incident_complaint_reported_by dispatch and situation_provider_primary_impression_code_and_description

