---
title: "Examining NAs"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

```{r, warning = FALSE}
library(naniar)
library(ggplot2)

source(here::here("src","Profiling","joining_albemarle_charlottesville.R"))
```

## NA Counts

```{r}
miss_var_summary(ems_full)
```

### Core Administrative Variables

Get a list of variables that could reasonably exist on every call:

```{r}
# non patient vars
ems_full %>% 
  select(incident_complaint_reported_by_dispatch, 
         response_incident_number,
         response_ems_unit_call_sign,
         incident_type,
         incident_date,
         scene_gps_latitude,
         scene_gps_longitude,
         agency_name) %>% 
  gg_miss_var()
```

#### NA Co-ocurrence

```{r}
# non patient vars
ems_full %>% 
  select(incident_complaint_reported_by_dispatch,
         response_incident_number,
         response_ems_unit_call_sign,
         incident_type,
         incident_date,
         scene_gps_latitude,
         scene_gps_longitude,
         agency_name) %>%  
  gg_miss_upset()
```

Show for just albemarle and charlottesville respectively

```{r}
# non patient vars
albemarle %>% 
  select(incident_complaint_reported_by_dispatch,
         response_incident_number,
         response_ems_unit_call_sign,
         incident_type,
         incident_date,
         scene_gps_latitude,
         scene_gps_longitude,
         agency_name) %>%  
  gg_miss_upset()

# non patient vars
charlottesville %>% 
  select(incident_complaint_reported_by_dispatch,
         response_incident_number,
         response_ems_unit_call_sign,
         incident_type,
         incident_date,
         scene_gps_latitude,
         scene_gps_longitude,
         agency_name) %>%  
  gg_miss_upset()
```

For albemarle, incident date is most specific ie non dates will turn into NAs, so I bet rows with that messed up are rows with weird data specification. Compare to albemarle, were date is missing for 0 observations. 


### Variables Relevant to Analysis

Get a list of variables that would be useful to have for our analysis:

```{r}
# non patient vars
ems_full %>% 
  select(patient_race_list, 
         patient_age,
         patient_age_range_in_years,
         patient_gender,
         situation_provider_primary_impression_code_and_description,
         total_unit_response_time,
         scene_gps_latitude,
         scene_gps_longitude,
         incident_psap_call_date_time,
         patient_initial_pulse_oximetry,
         patient_last_pulse_oximetry) %>% 
  gg_miss_var()
```

#### NA Co-ocurrence

```{r}
# non patient vars
ems_full %>% 
  select(patient_race_list, 
         patient_age,
         patient_age_range_in_years,
         patient_gender,
         situation_provider_primary_impression_code_and_description,
         total_unit_response_time,
         scene_gps_latitude,
         scene_gps_longitude,
         incident_psap_call_date_time,
         patient_initial_pulse_oximetry,
         patient_last_pulse_oximetry) %>% 
  rename(primary_impression = situation_provider_primary_impression_code_and_description) %>%
  gg_miss_upset()
```

Show for just albemarle and charlottesville respectively

```{r}
# non patient vars
albemarle %>% 
  select(patient_race_list, 
         patient_age,
         patient_age_range_in_years,
         patient_gender,
         situation_provider_primary_impression_code_and_description,
         total_unit_response_time,
         scene_gps_latitude,
         scene_gps_longitude,
         incident_psap_call_date_time,
         patient_initial_pulse_oximetry,
         patient_last_pulse_oximetry) %>% 
  rename(primary_impression = situation_provider_primary_impression_code_and_description) %>%
  gg_miss_upset()

# non patient vars
charlottesville %>% 
  select(patient_race_list, 
         patient_age,
         patient_age_range_in_years,
         patient_gender,
         situation_provider_primary_impression_code_and_description,
         total_unit_response_time,
         scene_gps_latitude,
         scene_gps_longitude,
         incident_psap_call_date_time,
         patient_initial_pulse_oximetry,
         patient_last_pulse_oximetry) %>% 
  rename(primary_impression = situation_provider_primary_impression_code_and_description) %>%
  gg_miss_upset()
```

Most often if we have data on the patient, what we're missing is pulse oximetry data. However, this own occurs in ~4000 cases for charlottesville. Also of note is that we sometimes have demographic data, but no primary impression. Albemarle also tends to have more missing pulse oximetry data. 

## NA Distribution

### Over Index

Do the number of NAs vary with location in the table?

```{r, fig.height= 8}
# note these are printed out of order
map(0:5, function(i) {
  ems_full %>% 
    select(incident_complaint_reported_by_dispatch,
           response_incident_number,
           response_ems_unit_call_sign,
           incident_type,
           incident_date,
           scene_gps_latitude,
           scene_gps_longitude,
           agency_name) %>% 
    slice((1 + 20000*i):min((20000*(i + 1)), nrow(ems_full))) %>% 
    vis_miss() +
    labs(title = paste((i*20000 + 1), min((20000*(i + 1)), nrow(ems_full)), sep = ":"))
})
```

Looks like albemarle has a lot more NAs, possibly due to the shifted values on some rows. 

Now lets look at number of NAs per row over index. 

```{r}
ems_full %>% 
  mutate(total_na = rowSums(is.na(ems_full))) %>%
  ggplot() + 
  geom_point(aes(x = 1:nrow(ems_full), y = total_na), alpha = 0.3) + 
  labs(y = "Number of NA's", x = "index") +
  ylim(0, 90) +
  geom_vline(xintercept = nrow(albemarle), color = "red") + # change between albemarle and charlottesville data 
  geom_hline(yintercept = 74, color = "green") + # min number of variables to contain all admin data
  theme_minimal()

# As density

ems_full %>% 
  mutate(total_na = rowSums(is.na(ems_full))) %>%
  ggplot() + 
  geom_hex(aes(x = 1:nrow(ems_full), y = total_na), alpha = 1) + 
  labs(y = "Number of NA's", x = "index") +
  ylim(0, 90) +
  geom_vline(xintercept = 60918, color = "red") +
  geom_hline(yintercept = 74, color = "green") +
  theme_minimal()
```

### Over Time

Let's look at the density of NAs over time. 

```{r}
ems_full %>% 
  mutate(total_na = rowSums(is.na(ems_full))) %>%
  ggplot() + 
  geom_point(aes(x = incident_date, y = total_na), alpha = 0.3) + 
  labs(y = "Number of NA's", x = "index") +
  ylim(0, 90) +
  geom_hline(yintercept = 74, color = "green") +
  theme_minimal() +
  facet_wrap(~source)

# As density

ems_full %>% 
  mutate(total_na = rowSums(is.na(ems_full))) %>%
  ggplot() + 
  geom_hex(aes(x = incident_date, y = total_na), alpha = 1) + 
  labs(y = "Number of NA's", x = "index") +
  ylim(0, 90) +
  geom_hline(yintercept = 74, color = "green") +
  theme_minimal() +
  facet_wrap(~source)
```

To me this says most of the weird rows in ems_full also don't have a real date (expected)
