---
title: "Untitled"
output: html_document
---

```{r}
library(glue)
library(tidyr)
library(ggridges)
library(ggplot2)
library(stringr)

source(here::here("src","Profiling","joining_albemarle_charlottesville.R"))
```

```{r}
charlottesville %>% 
  mutate(total_na = rowSums(is.na(.))) %>%
  select(total_na, everything()) %>% 
  # filter(response_incident_number == "2017-00002675")
  group_by(response_incident_number) %>%
  count(sort = TRUE) %>% 
  group_by(n) %>% 
  count()

charlottesville %>% 
  group_by(response_incident_number) %>%
  mutate(num_dups = n()) %>% 
  filter(num_dups == 2)
```

Things that appear to split a set:
  - conflicting situation_provider_secondary_impression_description_and_code (could likely turn into a list column)
  - one row is missing data another has
  - conflicting vitals_level_of_responsiveness_avpu
  - One row from albemarle, one from charlottesville
  
Going forward, I'm going to consider a unique observation to be a (response_incident_number, response_ems_unit_call_sign) pair
  
Lets get a more formal look at where the mismatches are:

```{r}
unique_elements <- charlottesville %>% 
  filter(response_incident_number != "0") %>% # an error response incident number
  group_by(response_incident_number, response_ems_unit_call_sign, outcome_external_report_number) %>% 
  mutate(total_dups = n()) %>%
  filter(total_dups > 1) %>% 
  summarise(across(everything(), ~length(unique(.x[!is.na(.x)]))))
```

This figure shows which variables with duplicated rows have more than one unique element in each column. It has been scaled to exagerrate the effect. The variables with peaks at 2 or greater are problem variables, meaning many of them have data that conflict with eachother. Examples of this include vitals_level_of_responsivness_avpu, which can read someone is alert for one duplicate and unresponsive for the other. 
```{r, fig.height = 10}
unique_elements %>% 
  ungroup() %>% 
  pivot_longer(cols = 3:61) %>% 
  mutate(total_dups = 1/total_dups, value = total_dups * value) %>% 
  ggplot(aes(x = value, y = name)) +
  geom_density_ridges(panel_scaling = FALSE, scale = 5) +
  theme_minimal() +
  coord_cartesian(xlim= c(1.8, 13))
```

Biggest problems vars:
- vitals_level_of_responsiveness_avpu
- situation_provider_secondary_impression_description_and_code

Smaller problem vars:
- situation_complaint_duration
- situation_complaint_duration_time_units
- patient_age
- patient_age_range
- situation_provider_primary_impression_code_and_description
- situation_primary_complaint_statement_list
- vitals_cardiac_rhythm_ecg_findings_list


Do some duplicate removal

Start by removing full duplicates and filling NAs within observations with other data from them. 
```{r}
charlottesville <- as_tibble(charlottesville) # coerce into a tibble

fill_step <- charlottesville %>% 
  mutate(situation_provider_secondary_impression_description_and_code = ifelse(situation_provider_secondary_impression_description_and_code %in% c("Not Applicable", "Not Recorded"), NA, situation_provider_secondary_impression_description_and_code)) %>% 
  filter(!duplicated(.)) %>%  # remove 6868 full duplicates sum(dulplicated(select(charlottesvill, -to_drop)))
  filter(response_incident_number != "0") %>%  # remove bogus response incident number
  select(response_incident_number, response_ems_unit_call_sign, everything()) %>% 
  group_by(response_incident_number, response_ems_unit_call_sign, outcome_external_report_number) %>% 
  fill(3:ncol(.), .direction = "updown")

fill_deduped <- fill_step %>% 
  ungroup() %>% 
  filter(!duplicated(.)) # gets rid of an additional 1054 duplicate rows sum(duplicated(fill_step))

# we want there to be only 36692 rows, still have ~6000 too many
# charlottesville %>% 
#   filter(response_incident_number != "0") %>% 
#   count(response_incident_number, response_ems_unit_call_sign) %>% 
#   nrow()
```


Now I'm going to turn situation_provider_secondary_impression_description_and_code into a list column. 

```{r}
combine_secondary <- fill_deduped %>% 
  group_by(response_incident_number, response_ems_unit_call_sign, outcome_external_report_number) %>% 
  mutate(situation_provider_secondary_impression_description_and_code = paste0(unique(situation_provider_secondary_impression_description_and_code[!is.na(situation_provider_secondary_impression_description_and_code)]), collapse = "|")) %>% 
  ungroup() %>% 
  filter(!duplicated(.)) %>% 
  mutate(situation_provider_secondary_impression_description_and_code = ifelse(situation_provider_secondary_impression_description_and_code == "", NA, situation_provider_secondary_impression_description_and_code))

# down to 39023 rows, making progress!
# nrow(combine_secondary)


```

I'm going to try and remove vitals_level_of_responsiveness_avpu to see if that improves things (probably can't use it since its so messed up)

```{r}
removed_responsiveness <- combine_secondary %>% 
  select(-vitals_level_of_responsiveness_avpu) %>% 
  filter(!duplicated(.))

# This removes another 1000 rows, down to 38092
# nrow(removed_responsivness)

removed_responsiveness %>% 
  ungroup() %>% 
  group_by(response_incident_number,
           response_ems_unit_call_sign,
           outcome_external_report_number) %>% 
  count()
```

Time to see whats going on with the age columns. 

```{r}
remaining_duplicates <- removed_responsiveness %>% 
  group_by(response_incident_number, response_ems_unit_call_sign) %>% 
  filter(n() > 1)
```

```{r}
deduped_plot_data <- removed_responsiveness %>% 
  group_by(response_incident_number, response_ems_unit_call_sign) %>% 
  mutate(total_dups = n()) %>%
  filter(total_dups > 1) %>% 
  summarise(across(everything(), ~length(unique(.x[!is.na(.x)]))))
```

```{r, fig.height=12}
deduped_plot_data %>% 
  ungroup() %>% 
  pivot_longer(cols = 3:60) %>% 
  mutate(total_dups = 1/total_dups, value = total_dups * value) %>% 
  ggplot(aes(x = value, y = name)) +
  geom_density_ridges(panel_scaling = FALSE, scale = 3) +
  theme_minimal() +
  coord_cartesian(xlim= c(1.8, 5))
```


#### Attempting to make unit of observation a single person at an incident

To constuct these units of observation, I'll be using response_incident_number, outcome_external_report, address (might not need), and 3 demographic variables (age, sex, race), where if one is different they are assumed to be different people. This will be done using the original dataset, because the analysis steps before had a different unit of observation assumed. To build a identifier for people, I'll use outcome_external_report, and the 3 demographic variables. 

I'll start by looking at addresses to see if they are consistent
```{r}
charlottesville$scene_incident_street_address %>% 
  unique() %>%
  length() # how are there only 8000 unique street addresses?

charlottesville %>% 
  count(scene_incident_street_address)
```

Need to turn all upper case, remove ., change all Street to ST, Avenue to AVE, Court to CT, Drive to DR

```{r}
charlottesville %>% 
  mutate(scene_incident_street_address = str_replace(toupper(scene_incident_street_address), "AVENUE", "AVE") %>% 
                                         str_replace(r"(\.)", "") %>% 
                                         ifelse((. == "-1" | . == "0" | . == "0 <UNKNOWN>" | . == "1"), NA, .) %>% 
                                         str_replace(r"(^0 )", "") %>% 
                                         str_replace("STREET", "ST") %>% 
                                         str_replace(r"(^[0-9]+$)", NA_character_) %>% 
                                         str_replace("CIRCLE", "CIR")) %>%
  count(scene_incident_street_address)
```

Looks like I caught most of them with this. 

Find out how many unique response_incident_number, outcome_external_report, address (might not need), and 3 demographic variables (age, sex, race)
```{r}
# look at unique pairs
charlottesville %>% 
  mutate(scene_incident_street_address = str_replace(toupper(scene_incident_street_address), "AVENUE", "AVE") %>% 
                                         str_replace(r"(\.)", "") %>% 
                                         ifelse((. == "-1" | . == "0" | . == "0 <UNKNOWN>" | . == "1"), NA, .) %>% 
                                         str_replace(r"(^0 )", "") %>% 
                                         str_replace("STREET", "ST") %>% 
                                         str_replace(r"(^[0-9]+$)", NA_character_) %>% 
                                         str_replace("CIRCLE", "CIR")) %>% 
  group_by(response_incident_number, outcome_external_report_number, scene_incident_street_address, patient_age, patient_gender, patient_race_list) %>% 
  count()

# see how many more rows we get with address included (only 9!)
charlottesville %>% 
  mutate(scene_incident_street_address = str_replace(toupper(scene_incident_street_address), "AVENUE", "AVE") %>% 
                                         str_replace(r"(\.)", "") %>% 
                                         ifelse((. == "-1" | . == "0" | . == "0 <UNKNOWN>" | . == "1"), NA, .) %>% 
                                         str_replace(r"(^0 )", "") %>% 
                                         str_replace("STREET", "ST") %>% 
                                         str_replace(r"(^[0-9]+$)", NA_character_) %>% 
                                         str_replace("CIRCLE", "CIR")) %>% 
  group_by(response_incident_number, outcome_external_report_number, patient_age, patient_gender, patient_race_list) %>% 
  count() %>% 
  nrow() - 
  charlottesville %>% 
  mutate(scene_incident_street_address = str_replace(toupper(scene_incident_street_address), "AVENUE", "AVE") %>% 
                                         str_replace(r"(\.)", "") %>% 
                                         ifelse((. == "-1" | . == "0" | . == "0 <UNKNOWN>" | . == "1"), NA, .) %>% 
                                         str_replace(r"(^0 )", "") %>% 
                                         str_replace("STREET", "ST") %>% 
                                         str_replace(r"(^[0-9]+$)", NA_character_) %>% 
                                         str_replace("CIRCLE", "CIR")) %>% 
  group_by(response_incident_number, outcome_external_report_number, scene_incident_street_address, patient_age, patient_gender, patient_race_list) %>% 
  count() %>% 
  nrow()
```

I'm going to try and run the deduplication steps on these new units of observations and see what happens, I won't be including address because with only 9 differences, those are likely due to data entry errors

```{r}
prepared_charlottesville <- charlottesville %>% 
  as_tibble() %>% 
  mutate(situation_provider_secondary_impression_description_and_code = ifelse(situation_provider_secondary_impression_description_and_code %in% c("Not Applicable", "Not Recorded"), NA, situation_provider_secondary_impression_description_and_code)) %>% 
  mutate(scene_incident_street_address = str_replace(toupper(scene_incident_street_address), "AVENUE", "AVE") %>% 
                                         str_replace(r"(\.)", "") %>% 
                                         ifelse((. == "-1" | . == "0" | . == "0 <UNKNOWN>" | . == "1"), NA, .) %>% 
                                         str_replace(r"(^0 )", "") %>% 
                                         str_replace("STREET", "ST") %>% 
                                         str_replace(r"(^[0-9]+$)", NA_character_) %>% 
                                         str_replace("CIRCLE", "CIR")) %>% 
  distinct() %>% 
  filter(response_incident_number != "0") %>%  # remove bogus response incident number
  select(response_incident_number, response_ems_unit_call_sign, everything()) # reorder columns for ease of reading


nrow(prepared_charlottesville)

# fill missing values that have existing values within unit of observation
# because NAs will make their own group, going to do some complicated thing to fill in age, gender, race, address, and report number
# filled_charlottesville <- 
filled_grouping_columns <- prepared_charlottesville %>% 
  group_by(response_incident_number) %>% 
  mutate(across(c(outcome_external_report_number, 
                  scene_incident_street_address, 
                  patient_age,
                  patient_gender,
                  patient_race_list),
                function(column) {
                  if(length(unique(patient_age[!is.na(patient_age)])) <= 1 &
                     length(unique(patient_gender[!is.na(patient_gender)])) <= 1 &
                     length(unique(patient_race_list[!is.na(patient_race_list)])) <= 1 &
                     length(unique(scene_incident_street_address[!is.na(scene_incident_street_address)])) <= 1 &
                     length(unique(outcome_external_report_number[!is.na(outcome_external_report_number)])) <= 1) {
                    
                    column = ifelse(length(unique(column[!is.na(column)])) == 0, NA, unique(column[!is.na(column)])) # catch where all NA
                  } else {
                    column
                  }
                })) %>% 
  ungroup() %>% 
  distinct()

nrow(filled_grouping_columns)

filled_charlottesville <- filled_grouping_columns %>% 
  group_by(response_incident_number, outcome_external_report_number, scene_incident_street_address, patient_age, patient_gender, patient_race_list) %>% 
  fill(3:ncol(.), .direction = "updown") %>% 
  ungroup() %>% 
  distinct()

nrow(filled_charlottesville)
```

Now to combine secondary impressions
```{r}
combined_sec_impress <- filled_charlottesville %>% 
  group_by(response_incident_number, 
           outcome_external_report_number, 
           scene_incident_street_address, 
           patient_age, 
           patient_gender, 
           patient_race_list) %>% 
  mutate(situation_provider_secondary_impression_description_and_code = paste0(unique(situation_provider_secondary_impression_description_and_code[!is.na(situation_provider_secondary_impression_description_and_code)]), collapse = "|")) %>% 
  ungroup() %>% 
  distinct() %>% 
  mutate(situation_provider_secondary_impression_description_and_code = ifelse(situation_provider_secondary_impression_description_and_code == "", NA, situation_provider_secondary_impression_description_and_code))

nrow(combined_sec_impress)
```

Now to combine unit call signs

First find out how many units can show up at most

```{r}
combined_sec_impress %>% 
  group_by(response_incident_number) %>% 
  summarize(unique_call_signs = length(unique(response_ems_unit_call_sign[!is.na(response_ems_unit_call_sign)]))) %>% 
  arrange(desc(unique_call_signs))
```


```{r}
combined_departments <- combined_sec_impress %>% 
  group_by(response_incident_number, 
           outcome_external_report_number, 
           scene_incident_street_address, 
           patient_age, 
           patient_gender, 
           patient_race_list) %>% 
  mutate(across(c("response_ems_unit_call_sign",
                  "incident_dispatch_notified_to_unit_arrived_on_scene_in_minutes",
                  "total_unit_response_time",
                  "scene_incident_location_type",
                  "agency_name",
                  "incident_crew_member_full_name_list",
                  "incident_crew_member_level_list"
                  ), 
                function(column) {
                    comb_col = paste0(unique(column[!is.na(column)]), collapse = "|")
                  ifelse(comb_col == "", NA, comb_col)
                })) %>% 
  ungroup() %>% 
  mutate(response_ems_unit_call_sign = ifelse(response_ems_unit_call_sign == "", NA, response_ems_unit_call_sign)) %>% 
  distinct()
```

```{r}
nrow(combined_departments)

combined_departments %>% 
  group_by(response_incident_number, 
           outcome_external_report_number, 
           scene_incident_street_address, 
           patient_age, 
           patient_gender, 
           patient_race_list) %>% 
  filter(n() > 1)

combined_departments %>% 
  group_by(response_incident_number, 
           outcome_external_report_number, 
           scene_incident_street_address, 
           patient_age, 
           patient_gender, 
           patient_race_list) %>% 
  filter(n() > 1)
```

```{r}
charlottesville %>% filter(response_incident_number == "2016-00006406")
```



```{r}
unique_elements_dept <- combined_departments %>% 
  filter(response_incident_number != "0") %>% # an error response incident number
  group_by(response_incident_number, 
           outcome_external_report_number, 
           scene_incident_street_address, 
           patient_age, 
           patient_gender, d
           patient_race_list) %>% 
  mutate(total_dups = n()) %>%
  filter(total_dups > 1) %>% 
  summarise(across(everything(), ~n_distinct(.x, na.rm = TRUE)))
```

This figure shows which variables with duplicated rows have more than one unique element in each column. It has been scaled to exagerrate the effect. The variables with peaks at 2 or greater are problem variables, meaning many of them have data that conflict with eachother. Examples of this include vitals_level_of_responsivness_avpu, which can read someone is alert for one duplicate and unresponsive for the other. 
```{r, fig.height = 10}
unique_elements_dept %>% 
  ungroup() %>% 
  pivot_longer(cols = -c(response_incident_number, 
           outcome_external_report_number, 
           scene_incident_street_address, 
           patient_age, 
           patient_gender, 
           patient_race_list, 
           total_dups)) %>% 
  mutate(total_dups = 1/total_dups, value = total_dups * value) %>% 
  ggplot(aes(x = value, y = name)) +
  geom_density_ridges(panel_scaling = FALSE, scale = 5) +
  theme_minimal() +
  coord_cartesian(xlim= c(1.8, 5.5))
```

