---
title: "11 Baseline Onet55 Clustering"
author:
  "Crystal Zang"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    highlight: tango
editor_options: 
  chunk_output_type: console
weight: 1
draft: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
for (pkg in c("tidyverse",  "data.table", "R.utils", "maditr", "stringr", "stringi", "dplyr", "ggplot2", "lubridate", "gt", "DataExplorer", "TraMineR", "TraMineRextras", "cluster", "tools")) {
  library(pkg, character.only = TRUE)
}

uva_color_palette <- 
c("#232D4B", #space cadet
  "#2C4F6B", #indigo dye
  "#0E879C", #blue munsell
  "#60999A", #cadet blue
  "#D1E0BF", #tea green
  "#D9E12B", #pear
  "#E6CE3A", #citrine
  "#E6A01D", #marigold
  "#E57200" #princeton orange
)
```

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
#import veteran job data
bg_vet_job <-read_csv("~/git/dspg20career/data/04_bg_vet_job.csv")%>%
  mutate(year_enter_job_market = year(date_enter_job_market))%>%
  select(-noofjobs, -sector, -tenure)
colnames(bg_vet_job)

#veterans and the date they ended their last ONET 55 job
vet_endmilitary <- bg_vet_job%>%
  mutate(date_end_onet55 = if_else(is_onet55==T, enddate, as.Date(NA)))%>%
  filter(!is.na(date_end_onet55))%>%  #exluce people who don't have valid onet55 code
  select(id, date_end_onet55) %>%
  #keep the latest onet55 job
  group_by(id)%>%   
  arrange(desc(date_end_onet55))%>%
  group_by(id)%>%
  distinct(id, .keep_all = TRUE) 

bg_vet_job  <- inner_join(bg_vet_job, vet_endmilitary, by = "id")

## sequence data
bg_vet_job_seq <- bg_vet_job %>%
  select(id, end_year, onet_job_zone, startdate, enddate, job_duration_day, date_end_onet55, year_enter_job_market)%>%
  mutate(year_end_onet55 = year(date_end_onet55))%>%
  select(-year_enter_job_market)%>%
  filter(startdate >= date_end_onet55)%>% #find jobs that came after the date ended onet55 job
  filter(onet_job_zone != 55)  %>%  #filter out 55 jobs that have the same start and end date  
  mutate(start_year = year(startdate))


check <- bg_vet_job_seq%>%
  mutate(year_until_first_job = start_year- year_end_onet55)%>%
  group_by(id)%>%
  summarize(years = min(year_until_first_job))

ggplot(check, aes(x = years)) + 
  geom_histogram(binwidth = 1.4) + 
  labs(title = "", x = "years until first job") 

summary(check$years)

bg_vet_job_seq <- bg_vet_job_seq%>%
  mutate(start_year = start_year - year_end_onet55 + 1)%>%  #transform from calender year to year start sequence analysis
  mutate(end_year = end_year - year_end_onet55 + 1) %>% #transform from calender year to year start sequence analysis
  select(id, start_year, end_year, onet_job_zone)%>%
  group_by(id)%>%
  arrange(desc(onet_job_zone))%>%
  group_by(id)%>%
  distinct(id, start_year, end_year, .keep_all = TRUE)
```


```{r}
#table(bg_vet_job_seq$onet_job_zone)
#length(unique(bg_vet_job_seq$id))  #confirm sample size match

bg_vet_job_seq <- as.matrix(bg_vet_job_seq)
bg_vet_job_seq <- as.data.frame(bg_vet_job_seq)

min(bg_vet_job_seq$start_year)
max(bg_vet_job_seq$start_year)
```

```{r}
# STS format
sts_vet <- seqformat(bg_vet_job_seq, from = "SPELL", to = "STS",
                     id = "id",  begin = "start_year", end = "end_year", 
                     status = "onet_job_zone", process = FALSE)

#obtain jobs that appear after 10 years they exit military 
sts_vet <- sts_vet[, 1:10]

# Here we are renaming columns to be in format "yn" (year in the job market)
names(sts_vet) <- paste0("y", 1:ncol(sts_vet))
```

# Normalize transition matrix
```{r}
vet.seq <- seqdef(sts_vet, left="Unemployment", gaps="Unemployment", right="Unemployment/Exit", cpal = c("#CBBEB5",  uva_color_palette[1], uva_color_palette[2],uva_color_palette[4],  uva_color_palette[6], uva_color_palette[9], "#CBBEB5", "#CBBEB5","#CBBEB5"))

## Computing transition rates
#counts
transition_matrix <- seqtrate(vet.seq, weighted=FALSE, count=TRUE)
round(transition_matrix,2)
#standardize counts

#make the diagnal 0
diag(transition_matrix) = 0
round(transition_matrix,2)

#calculate rowsum
rowsum <- apply(transition_matrix, 1, sum)

#calculate percent 
for(i in 1:nrow(transition_matrix)) {
  for (j in 1:ncol(transition_matrix)){
    transition_matrix[i,j] = transition_matrix[i,j]/rowsum[i]
  }
}

#code NAs as 0
for(j in 1:ncol(transition_matrix)){
      transition_matrix[9,j] = transition_matrix[9,j] = 0
}
round(transition_matrix,2)


#convert transition matrix to cost matrix
cost_matrix <- 2-transition_matrix
round(cost_matrix,2)

#make the diagnal 0
diag(cost_matrix) = 0
round(cost_matrix,2)
```

# Sequence Clustering
Sequence starts at the second year after they finished their last ONET 55 job
```{r, fig.height = 10, fig.width = 8}
# vet.seq.OM <- seqdist(vet.seq, method = "OM", indel = 3, sm = cost_matrix, with.missing = TRUE)
# clusterward <- agnes(vet.seq.OM, diss = TRUE, method = "ward")
# saveRDS(clusterward, file = "data/clusterward_onet55_10yrs_normalized_matrix.rds")

clusterward <- readRDS(file = "~/git/dspg20career/data/clusterward_onet55_10yrs_normalized_matrix.rds")


#dendrogram 
plot(clusterward, which.plots =2)
abline(h=200, col="purple")
abline(h=90, col="red")
```


## Six clusters
```{r, fig.height = 10, fig.width = 8}

cluster6 <- cutree(clusterward, k=6)

cluster6 <- factor(cluster6, labels = c("Type 1",  "Type 2", "Type 3", "Type 4", "Type 5", "Type 6"))
table(cluster6)

#longitudinal plot
seqfplot(vet.seq, group = cluster6, pbarw = T)

#another cluster plot
seqmtplot(vet.seq, group = cluster6)
```


## Eight clusters
```{r, fig.height = 13, fig.width = 8}
cluster8 <- cutree(clusterward, k=8)
cluster8 <- factor(cluster8, labels = c("Type 1",  "Type 2", "Type 3", "Type 4", "Type 5", "Type 6", "Type 7", "Type 8"))
table(cluster8)

#longitudinal plot
seqfplot(vet.seq, group = cluster8, pbarw = T)

#another cluster plot
seqmtplot(vet.seq, group = cluster8)
```


## Twelve clusters
```{r, fig.height = 18, fig.width = 8}
cluster12 <- cutree(clusterward, k=12)
cluster12 <- factor(cluster12, labels = c("Type 1",  "Type 2", "Type 3", "Type 4", "Type 5", "Type 6", "Type 7", "Type 8", "Type 9", "Type 10", "Type 11", "Type 12"))
table(cluster12)

#longitudinal plot
seqfplot(vet.seq, group = cluster12, pbarw = T)

#another cluster plot
seqmtplot(vet.seq, group = cluster12)
```

## Fifteen clusters
```{r, fig.height = 22, fig.width = 9}
cluster15 <- cutree(clusterward, k=15)
cluster15 <- factor(cluster15, labels = c("Type 1",  "Type 2", "Type 3", "Type 4", "Type 5", "Type 6", "Type 7", "Type 8", "Type 9", "Type 10", "Type 11", "Type 12", "Type 13", "Type 14", "Type 15"))
table(cluster15)

#longitudinal plot
seqfplot(vet.seq, group = cluster15, pbarw = T)

#another cluster plot
seqmtplot(vet.seq, group = cluster15)
```


## Twenty clusters
```{r, fig.height = 31, fig.width = 12}
cluster20 <- cutree(clusterward, k=20)
cluster20 <- factor(cluster20, labels = c("Type 1",  "Type 2", "Type 3", "Type 4", "Type 5", "Type 6", "Type 7", "Type 8", "Type 9", "Type 10", "Type 11", "Type 12", "Type 13", "Type 14", "Type 15", "Type 16", "Type 17", "Type 18", "Type 19","Type 20"))
table(cluster20)

#longitudinal plot
seqfplot(vet.seq, group = cluster20, pbarw = T)

#another cluster plot
seqmtplot(vet.seq, group = cluster20)
```

# Q1. Do Veterans Who Had Different Education Level Have Different Career Trajectory?
## Clusters V.S. Degree in 20 clusters
```{r}
vet <- unique(bg_vet_job_seq$id)
#vet <- rownames(sts_vet)

vet_df <- as.data.frame(cbind(vet, cluster20))
colnames(vet_df) <- c("id", "cluster")

class(vet_df$id)

bg_vet_demographic <-read_csv("~/git/dspg20career/data/02_bg_vet_demographic.csv")
vet_df <- left_join(vet_df, bg_vet_demographic, by = "id")

degree <- vet_df %>%
  filter(!is.na(degree_highest))%>%
  group_by(cluster, degree_highest)%>%
  summarize(count=n())%>%
  group_by(degree_highest)%>%
  mutate(sum = sum(count))%>%
  mutate(perc = count/sum)

#degree$count[is.na(degree$count)] <- 0

degree$cluster <- as.factor(degree$cluster)
degree$degree_highest <- factor(degree$degree_highest, levels = c("others","certificate","highschool", "associate", "bachelor", "master", "doctor"))


degree <- as.data.frame(degree)

ggplot(degree, aes(cluster, degree_highest, fill= count)) + 
  geom_tile()+
  scale_fill_gradient(low="white", high="blue")+
  theme_classic()+
  labs(title = "Clustering X Degree --before normalization")

ggplot(degree, aes(cluster, degree_highest, fill= perc)) + 
  geom_tile()+
  scale_fill_gradient(low="white", high="blue")+
  theme_classic()+
  labs(title = "Clustering X Degree --after normalization")

```

## Clusters V.S. Degree in 8 clusters
```{r}
vet <- unique(bg_vet_job_seq$id)
vet_df <- as.data.frame(cbind(vet, cluster8))
colnames(vet_df) <- c("id", "cluster")

class(vet_df$id)

vet_df <- left_join(vet_df, bg_vet_demographic, by = "id")

degree <- vet_df %>%
  filter(!is.na(degree_highest))%>%
  group_by(cluster, degree_highest)%>%
  summarize(count=n())%>%
  group_by(degree_highest)%>%
  mutate(sum = sum(count))%>%
  mutate(perc = count/sum)

#degree$count[is.na(degree$count)] <- 0

degree$cluster <- as.factor(degree$cluster)
degree$degree_highest <- factor(degree$degree_highest, levels = c("others","certificate","highschool", "associate", "bachelor", "master", "doctor"))


degree <- as.data.frame(degree)

ggplot(degree, aes(cluster, degree_highest, fill= count)) + 
  geom_tile()+
  scale_fill_gradient(low="white", high="blue")+
  theme_classic()+
  labs(title = "Clustering X Degree --before normalization")


ggplot(degree, aes(cluster, degree_highest, fill= perc)) + 
  geom_tile()+
  scale_fill_gradient(low="white", high="blue")+
  theme_classic()+
  labs(title = "Clustering X Degree --after normalization")
```

```{r}
# contingency table
vet <- unique(bg_vet_job_seq$id)
vet_df <- as.data.frame(cbind(vet, cluster8))
colnames(vet_df) <- c("id", "cluster")
vet_df$cluster <- as.factor(vet_df$cluster)

vet_df <- vet_df %>%
  left_join(bg_vet_demographic, by = "id")%>%
  filter(!is.na(degree_highest))

table <- table(vet_df$cluster, vet_df$degree_highest)
table
degree_p_matrix <- c()
for (degree in 1: ncol(table)){ 
  degree_p_vector_i <- c()
  for (i in 1:nrow(table)){ #cluster 1-8
    tb_i <- matrix(NA, nrow = 2, ncol = 2)
    tb_i[1,1] <- table[i,1]
    tb_i[2,1] <- sum(table[,1])-table[i,1]
    tb_i[1,2] <- sum(table[1,])-table[i,1]
    tb_i[2,2] <- sum(table)-tb_i[1,1] - tb_i[2,1] -tb_i[1,2] 
    degree_p_vector_i[i] <- fisher.test(tb_i)$p.value
  }
  degree_p_matrix <- cbind(degree_p_matrix, degree_p_vector_i)
}
colnames(degree_p_matrix) <- c("associate","bachelor","certificate", "doctor", "highschool", "master", "others")
degree_p_matrix <- degree_p_matrix[,c("others","certificate","highschool", "associate", "bachelor", "master", "doctor")]

degree_p_matrix
degree_p_matrix <= 0.05
table(degree_p_matrix <= 0.05)
round(degree_p_matrix,2)
```



# Q2. Do Female and Male Veterans have different career trajectory?
## Clusters V.S. Gender in 8 clusters
In our sample, we have 15.9% (N=824) female veterans, 84.1% (N=3666) male veterans.
We conducted $\chi^2$ test. Using 0.05 as the critical value, we conclude that female is statistically different from male in cluster 3, 4, 7, and 8.
```{r}
dev.off()

gender <- vet_df %>%
  filter(!is.na(gender))%>%
  group_by(cluster, gender)%>%
  summarize(count=n())%>%
  group_by(gender)%>%
  mutate(sum = sum(count))%>%
  mutate(perc = count/sum )

#degree$count[is.na(degree$count)] <- 0

gender$cluster <- as.factor(gender$cluster)
gender$gender <- as.factor(gender$gender)


gender <- as.data.frame(gender)

ggplot(gender, aes(cluster, gender, fill= count)) + 
  geom_tile()+
  scale_fill_gradient(low="white", high="blue")+
  theme_classic()+
  labs(title = "Clustering X Gender --before normalization")
  

ggplot(gender, aes(cluster, gender, fill= perc)) + 
  geom_tile()+
  scale_fill_gradient(low="white", high="blue")+
  theme_classic()+
  labs(title = "Clustering X Gender --after normalization")

# contingency table
vet <- unique(bg_vet_job_seq$id)
vet_df <- as.data.frame(cbind(vet, cluster8))
colnames(vet_df) <- c("id", "cluster")
vet_df$cluster <- as.factor(vet_df$cluster)
vet_df <- vet_df %>%
  left_join(bg_vet_demographic, by = "id")%>%
  filter(!is.na(gender))

table <- table(vet_df$cluster, vet_df$gender)
table

gender_p_vector <- c()
for (i in 1: nrow(table)){
  tb_i <- matrix(NA, nrow = 2, ncol = 2)
  tb_i[1,1] <- table[i,1]
  tb_i[2,1] <- sum(table[,1])-table[i,1]
  tb_i[1,2] <- table[i,2]
  tb_i[2,2] <- sum(table[,2])-table[i,2]
  gender_p_vector[i] <- chisq.test(tb_i)$p.value
}
gender_p_vector
gender_p_vector <= 0.05

```

