---
title: "CoreLogic Fairfax Profiling"
author: "Morgan Klutzke"
date: "6/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE, message=FALSE, echo=FALSE}
library(DataExplorer)
library(ggplot2)
library(scales)
library(dplyr)
library(purrr)
library(data.table)
library(maditr)
```

```{r functions}
check_calc <- function(vec) {
  blanks <- 0L
  true_na <- 0L
  written_na <- 0L
  len <- length(x = vec)
  for (elem in vec) {
    if (is.na(x = elem)) {
      true_na <- true_na + 1L
    } else if (elem == "na") {
      written_na <- written_na + 1L
    } else if (elem == "") {
      blanks <- blanks + 1L
    }
  }
  percent_complete <- (len - (blanks + true_na + written_na)) / len
  unique_values <- length(unique(vec))
  tibble(blanks = blanks,
         true_na = true_na,
         written_na = written_na,
         percent_complete = percent_complete,
         unique_values = unique_values)
}
check_complete <- function(df) {
  z <- deparse(substitute(df))
  map_df(.x = df, .f = check_calc) %>%
    mutate(column = colnames(df)) %>%
    mutate(set = print(z))  %>%
    select(set, column, blanks, true_na, written_na, percent_complete, unique_values)
}

draw_histograms <- function(variable) {
  # For plotting data that are heavily skewed right
  # Plots using two histograms so R doesn't freak out
  data <- as.numeric(na.omit(variable))
  print(range(data))
  
  # Lower 98%
  data_sort <- sort(data.matrix(data))
  upper <- length(data_sort) - 2 * (length(data_sort) %/% 100)
  data_lower <- data_sort[0:upper]
  range(data_lower)
  lower_hist <- hist(data_lower)
  #text(lower_hist$mids, lower_hist$counts, labels = lower_hist$counts, adj = c(0.5, -0.1))
  
  # Upper 2%
  options(scipen = 15)
  data_upper <- data_sort[upper:length(data_sort)]
  range(data_upper)
  upper_hist <- hist(data_upper)
  text(upper_hist$mids, upper_hist$counts, labels = upper_hist$counts, adj = c(0.5, -0.1))
}
```

Checking completeness

```{r}
check_complete(select(ffx, 9:15)) %>% knitr::kable()
```

```{r}
introduce(select(ffx, 9:15))
plot_missing(select(ffx, 9:15))
```

Looking at market value: "The Total Market Value of the Parcel's Land & Improvement values as provided by the county or local taxing/assessment authority"

```{r}
draw_histograms(ffx$mkt_total_value)
```

Looking at tax amount: "The Total Tax amount provided by the county or local taxing/assessment authority"

```{r}
draw_histograms(ffx$tax_amount)
```

Looking at tax_year: "The tax year for which the taxes were billed" 
& assessed_year: "The assessment year for which the taxes were billed" 
(orange is tax_year, blue is assessed_year)

```{r}
ffx <- ffx %>%
  dt_mutate(tax_year = as.numeric(tax_year)) %>%
  dt_mutate(assessed_year = as.numeric(assessed_year))
ffx <- dt_mutate(ffx, year_diff = ifelse(test = tax_year & assessed_year, yes = tax_year - assessed_year, no = NA_integer_))
unique(ffx$year_diff)
unique(ffx$tax_year)
unique(ffx$assessed_year)
```

```{r}
tax_year <- as.numeric(na.omit(ffx$tax_year))
tax_year_counts <- as.data.frame(table(tax_year))

assessed_year <- as.numeric(na.omit(ffx$assessed_year))
assessed_year_counts <- as.data.frame(table(assessed_year))

years <- data.frame("year" = tax_year_counts$tax_year, "tax_year" = tax_year_counts$Freq, "assessed_year" = assessed_year_counts$Freq)
 
ggplot(years, aes(x = year)) + geom_line(mapping = aes(y = tax_year, group = 1), color = "#E69F00") + geom_line(mapping = aes(y = assessed_year, group = 1), color = "#56B4E9") + ylab("frequency")
```


Looking at acres: "CoreLogic derived or county provided value that indicates the total acreage of the parcel"

```{r}
draw_histograms(ffx$acres)
```

Looking at land_square_footage: "Total land mass in square feet"

```{r}
draw_histograms(ffx$land_square_footage)
```

Looking at building__square_feet: "The size of the building in square feet. This field is most commonly populated as an aggregate total when a county does not differentiate between living and non-living areas"

```{r}
draw_histograms(ffx$building__square_feet)
```

